---
title: "Biomarkers Analysis Severes (manuscript)"
author: "Alex Gould"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: yes
    toc_depth: 3
    number_sections: yes
    latex_engine: xelatex
  rmarkdown::pdf_document:
    extra_dependencies: float
    toc: yes
html_document:
  toc: yes
toc_float: yes
---

# Setup

First let us set up the R environment:

## Global options

```{r global options, include = TRUE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE,
                      error = FALSE,
                      eval = TRUE,
                      dev = c('pdf', 'png'),
                      width = 60,
                      fig.align = 'center', 
                      fig.height = 10, 
                      fig.width = 10, 
                      fig.dpi = 300,
                      fig.pos = "!H",
                      dpi = 300,
                      out.extra = "",
                      message = FALSE, 
                      cache = FALSE,
                      pdf.options(encoding = "ISOLatin9.enc"),
                      fig.path='figs/rmd/') 
```

## Library 

```{r Library}
## Load the libraries
library(xlsx)
library(rstudioapi)
library(tidyverse)
library(jmv)
library(ggtext)
library(ggpubr)
library(gridtext) 
library(scales)
```

## Parameters

``` {r Parameters}
analysis_dir = "..."
data_dir     = "..."
data_dir2    = "..."
pw_message   = "..."
font         = element_text(family = "sans")
group_cols   = c("Control" = "#0099CC", "EHI" = "#ea4714") 
```

### Figure parameters 

First, create custom theme to use throughout:

```{r figure_theme}
#create custom theme
theme_ag <- function(){

  theme_classic() %+replace%
    
    theme(
      text = font,
      axis.text.y  = element_text(size = 12, margin = margin(r = 3)),
      axis.title.y = element_text(size = 16, angle = 90, vjust = 2),
      axis.text.x  = element_text(size = 16, margin = margin(r = 3)),
      axis.title.x = element_blank(),
      legend.position = "none"
    )
}
```

Now, make custom axis titles to use throughout:

```{r figure_custom_y_axis}
IFABP_axis = expression("Serum [I-FABP]" ~ "(pg·" * mL^-1 * ")")
CLDN3_axis = expression("Plasma [CLDN-3]" ~ "(ng·" * mL^-1 * ")")
Zon_axis   = expression("Serum [Zonulin]" ~ "(ng·" * mL^-1 * ")")
LBP_axis   = expression("Plasma [LBP]" ~ "(µg·" * mL^-1 * ")")
sCD14_axis = expression("Plasma [sCD14]" ~ "(µg·" * mL^-1 * ")")
```

## Load Data
Next we load in the data:

``` {r load_data}
## load data here...
```

### Data Filtering

Check cohort numbers:

```{r cohort_numbers}
## Cohort number
length(unique(data$UIN))

## Number per group (i.e. control/EHI)
length(unique(data$Pair))

## Check modsev numbers
table(data$ModSev)
```

Now remove the three troops that we did not take blood from.

```{r remove_troops}
## data filtering here...
```

Create two seperate data frames from raw data, 1 for serum bloods and 1 for EDTA bloods. 

```{r data_serum}
## data filtering here...
```

```{r data_EDTA}
## data filtering here...
```

# Severes vs Controls

This section will analyse the blood biomarkers of the severe EHI cases vs their pair matched controls. The N is presented in the descriptive section of each biomarker. 

## Intestinal Fatty Acid Binding Protein

### Descriptives

```{r IFABP_descriptives_modsev}
jmv::descriptives(
    formula = IFABP ~ ModSev,
    data = data_serum,
    iqr = TRUE,
    sw = TRUE)
```

### Test of Difference

non-parametric data:

```{r IFABP_non-parametric_modsev}
jmv::ttestIS(
    formula = IFABP ~ ModSev,
    data = data_serum,
    vars = IFABP,
    students = FALSE,
    mann = TRUE,
    norm = TRUE,
    effectSize = TRUE)
```

### Figure

``` {r IFABP_modsev_fig, fig.height = 5, fig.width = 6}
set.seed(123)

ymax <- max(data_serum$IFABP, na.rm = TRUE)
ymin <- min(data_serum$IFABP, na.rm = TRUE)

range <- ymax - ymin

## spacing parameters 
## vertical bracket leg drop = 2% of axis height
drop_prop   <- 0.02
## distance from bracket to p-value text = 3% of axis height
text_offset <- 0.03   

## derived values
leg_drop_ifabp <- range * drop_prop
text_raise <- range * text_offset

pval <- "italic(P) == 0.696"
y_bracket_ifabp <- 19500

IFABP_modsev_p <- ggplot(data_serum, aes(x = ModSev, y = IFABP)) +
  geom_jitter(aes(color = ModSev), 
              width = 0.3,
              height = 0,
              alpha = 0.6,
              size = 3) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  stat_summary(fun = median, 
               geom="crossbar", 
               color="black",
               width = 0.6,
               size = 0.6,
               show.legend = FALSE) +
  stat_summary(fun.data = median_iqr,
               geom="errorbar",
               color="black",
               width = 0.3,
               size = 0.8,
               show.legend = FALSE) +
  labs(x = "Group",
       y = IFABP_axis) +
  scale_x_discrete(labels = c("Non-EHI", "Severe EHI")) +
  scale_y_continuous(breaks = seq(0, 20000, 2000),
                     limits = c(-550, 20200),
                     labels = scales::label_number(big.mark = ",")) +
  scale_color_manual(values = group_cols) +
  geom_segment(aes(x = 1, xend = 2, y = y_bracket_ifabp, yend = y_bracket_ifabp), inherit.aes = FALSE) +
  geom_segment(aes(x = 1, xend = 1, y = y_bracket_ifabp, yend = y_bracket_ifabp - leg_drop_ifabp), inherit.aes = FALSE) +
  geom_segment(aes(x = 2, xend = 2, y = y_bracket_ifabp, yend = y_bracket_ifabp - leg_drop_ifabp), inherit.aes = FALSE) +
  annotate("text",
           x = 1.5,
           y = y_bracket_ifabp + text_raise,
           label = pval,
           parse = TRUE,
           size = 4,
           family = "sans") +
  theme_ag() +
  theme(axis.text.y.right = element_blank(),
        axis.ticks.y.right = element_blank(),
        axis.line.y.right = element_blank())

#view figure
IFABP_modsev_p
```

## Claudin 3

### Descriptives

```{r CLDN3_descriptives_modsev}
jmv::descriptives(
    formula = CLDN3 ~ ModSev,
    data = data_EDTA,
    iqr = TRUE,
    sw = TRUE)
```

### Test of Difference

parametric data:

```{r CLDN3_non-parametric_modsev}
jmv::ttestIS(
    formula = CLDN3 ~ ModSev,
    data = data_EDTA,
    vars = CLDN3,
    norm = TRUE,
    eqv = TRUE,
    effectSize = TRUE)
```

### Figure

``` {r CLDN3_modsev_fig, fig.height = 5, fig.width = 6}
set.seed(123)

ymax <- max(data_EDTA$CLDN3, na.rm = TRUE)
ymin <- min(data_EDTA$CLDN3, na.rm = TRUE)

range <- ymax - ymin

## spacing parameters 
## vertical bracket leg drop = 3% of axis height
drop_prop   <- 0.04
## distance from bracket to p-value text = 4% of axis height
text_offset <- 0.04 

## derived values
leg_drop_cldn3    <- range * drop_prop
text_raise  <- range * text_offset

p_numeric <- 0.6
pval <- paste0('italic(P) == "', formatC(p_numeric, format = "f", digits = 3), '"')
y_bracket_cldn3 <- 6.6

CLDN3_modsev_p <- ggplot(data_EDTA, aes(x = ModSev, y = CLDN3)) +
  geom_jitter(aes(color = ModSev),
              width = 0.3,
              height = 0,
              alpha = 0.6,
              size = 3) +
  stat_summary(fun = mean,
               geom = "point",
               color = "black",
               size = 5,
               show.legend = FALSE) +
  stat_summary(fun.data = mean_sd,
               geom = "errorbar",
               color = "black",
               width = 0.3,
               size = 1.0,
               show.legend = FALSE) +
  labs(x = "Group",
       y = CLDN3_axis) +
  scale_x_discrete(labels = c("Non-EHI", "Severe EHI")) +
  scale_y_continuous(breaks = seq(0, 7, 1),
                     limits = c(0, 7),
                     labels = scales::label_number(accuracy = 1)) +
  scale_color_manual(values = group_cols) +
  geom_segment(aes(x = 1, xend = 2, y = y_bracket_cldn3, yend = y_bracket_cldn3), inherit.aes = FALSE) +
  geom_segment(aes(x = 1, xend = 1, y = y_bracket_cldn3, yend = y_bracket_cldn3 - leg_drop_cldn3), inherit.aes = FALSE) +
  geom_segment(aes(x = 2, xend = 2, y = y_bracket_cldn3, yend = y_bracket_cldn3 - leg_drop_cldn3), inherit.aes = FALSE) +
  annotate("text",
           x = 1.5,
           y = y_bracket_cldn3 + text_raise,
           label = pval,
           parse = TRUE,
           size = 4,
           family = "sans") +
  theme_ag()

#view figure
CLDN3_modsev_p
```

## Zonulin

### Descriptives

```{r Zonulin_descriptives_modsev}
jmv::descriptives(
    formula = Zonulin ~ ModSev,
    data = data_serum,
    iqr = TRUE,
    sw = TRUE)
```

### Test of Difference

non-parametric data:

```{r Zonulin_non-parametric_modsev}
jmv::ttestIS(
    formula = Zonulin ~ ModSev,
    data = data,
    vars = Zonulin,
    norm = TRUE,
    eqv = TRUE,
    effectSize = TRUE)
```

### Figure

``` {r Zonulin_modsev_fig, fig.height = 5, fig.width = 6}
set.seed(123)

ymax <- max(data_serum$Zonulin, na.rm = TRUE)
ymin <- min(data_serum$Zonulin, na.rm = TRUE)

range <- ymax - ymin

## spacing parameters 
## vertical bracket leg drop = 3% of axis height
drop_prop   <- 0.03
## distance from bracket to p-value text = 3% of axis height
text_offset <- 0.03   

## derived values
leg_drop_zon    <- range * drop_prop
text_raise  <- range * text_offset

pval <- "italic(P) == 0.333"
y_bracket_zon <- 58

Zon_modsev_p <- ggplot(data_serum, aes(x = ModSev, y = Zonulin)) +
  geom_jitter(aes(color = ModSev),
              width = 0.3,
              height = 0,
              alpha = 0.6,
              size = 3) +
  stat_summary(fun = mean,
               geom = "point",
               color = "black",
               size = 5,
               show.legend = FALSE) +
  stat_summary(fun.data = mean_sd,
               geom = "errorbar",
               color = "black",
               width = 0.3,
               size = 1.0,
               show.legend = FALSE) +
  labs(x = "Group",
       y = Zon_axis) +
  scale_x_discrete(labels = c("Non-EHI", "Severe EHI")) +
  scale_y_continuous(breaks = seq(0, 60, 10),
                     limits = c(0, 60)) +
  scale_color_manual(values = group_cols) +
  geom_segment(aes(x = 1, xend = 2, y = y_bracket_zon, yend = y_bracket_zon), inherit.aes = FALSE) +
  geom_segment(aes(x = 1, xend = 1, y = y_bracket_zon, yend = y_bracket_zon - leg_drop_zon), inherit.aes = FALSE) +
  geom_segment(aes(x = 2, xend = 2, y = y_bracket_zon, yend = y_bracket_zon - leg_drop_zon), inherit.aes = FALSE) +
  annotate("text",
           x = 1.5,
           y = y_bracket_zon + text_raise,
           label = pval,
           parse = TRUE,
           size = 4,
           family = "sans") +
  theme_ag()

#view figure
Zon_modsev_p
```

## Lipopolysaccharide Binding Protein

### Descriptives

```{r LBP_descriptives_modsev}
jmv::descriptives(
    formula = LBP ~ ModSev,
    data = data_EDTA,
    iqr = TRUE,
    sw = TRUE)
```

### Test of Difference

parametric data:

```{r LBP_parametric_class}
jmv::ttestIS(
    formula = LBP ~ ModSev,
    data = data_EDTA,
    vars = LBP,
    norm = TRUE,
    eqv = TRUE,
    effectSize = TRUE)
```

### Figure

``` {r LBP_modsev_fig, fig.height = 5, fig.width = 6}
set.seed(123)

ymax <- max(data_EDTA$LBP, na.rm = TRUE)
ymin <- min(data_EDTA$LBP, na.rm = TRUE)

range <- ymax - ymin

## spacing parameters 
## vertical bracket leg drop = 3% of axis height
drop_prop  <- 0.03
## distance from bracket to p-value text = 3% of axis height
text_offset <- 0.03   

p_numeric <- 0.93
pval <- paste0('italic(P) == "', formatC(p_numeric, format = "f", digits = 3), '"')
y_bracket_lbp <- 6.3

## derived values
leg_drop_lbp <- range * drop_prop
text_raise <- range * text_offset

LBP_modsev_p <- ggplot(data_EDTA, aes(x = ModSev, y = LBP)) +
  geom_jitter(aes(color = ModSev), 
              width = 0.3,
              height = 0,
              alpha = 0.6,
              size = 3) +
  stat_summary(fun = mean, 
               geom = "point", 
               color = "black",
               size = 5,
               show.legend = FALSE) +
  stat_summary(fun.data = mean_sd, 
               geom = "errorbar", 
               color = "black",
               width = 0.3,
               size = 1.0,
               show.legend = FALSE) +
  labs(x = "Group",
       y = LBP_axis) +
  scale_x_discrete(labels = c("Non-EHI", "Severe EHI")) +
  scale_y_continuous(breaks = seq(0.0, 6.0, 1),
                     limits = c(0.0, 6.5),
                     labels = scales::label_number(accuracy = 1)) +
  scale_color_manual(values = group_cols) +
  geom_segment(aes(x = 1, xend = 2, y = y_bracket_lbp, yend = y_bracket_lbp), inherit.aes = FALSE) +
  geom_segment(aes(x = 1, xend = 1, y = y_bracket_lbp, yend = y_bracket_lbp - leg_drop_lbp), inherit.aes = FALSE) +
  geom_segment(aes(x = 2, xend = 2, y = y_bracket_lbp, yend = y_bracket_lbp - leg_drop_lbp), inherit.aes = FALSE) +
  annotate("text",
           x = 1.5,
           y = y_bracket_lbp + text_raise,
           label = pval,
           parse = TRUE,
           size = 4,
           family = "sans") +
  theme_ag()

#view figure
LBP_modsev_p
```

## Soluble Cluster of Differentiation 14

### Descriptives

```{r sCD14_descriptives_modsev}
jmv::descriptives(
    formula = sCD14 ~ ModSev,
    data = data_EDTA,
    iqr = TRUE,
    sw = TRUE)
```

### Test of Difference

parametric data:

```{r sCD14_parametric_modsev}
jmv::ttestIS(
    formula = sCD14 ~ ModSev,
    data = data_EDTA,
    vars = sCD14,
    norm = TRUE,
    eqv = TRUE,
    effectSize = TRUE)
```

### Figure

``` {r sCD14_modsev_fig, fig.height = 5, fig.width = 6}
set.seed(123)

ymax <- max(data_EDTA$sCD14, na.rm = TRUE)
ymin <- min(data_EDTA$sCD14, na.rm = TRUE)

range <- ymax - ymin

## spacing parameters 
## vertical bracket leg drop = 3% of axis height
drop_prop   <- 0.03
## distance from bracket to p-value text = 4% of axis height
text_offset <- 0.04   

## derived values
leg_drop_scd14    <- range * drop_prop
text_raise  <- range * text_offset

pval <- "italic(P) == 0.916"
y_bracket_scd14 <- 1.75

sCD14_modsev_p <- ggplot(data_EDTA, aes(x = ModSev, y = sCD14)) +
  geom_jitter(aes(color = ModSev), 
              width = 0.3,
              height = 0,
              alpha = 0.6,
              size = 3) +
  stat_summary(fun = mean, 
               geom = "point", 
               color = "black",
               size = 5,
               show.legend = FALSE) +
  stat_summary(fun.data = mean_sd, 
               geom = "errorbar", 
               color = "black",
               width = 0.3,
               size = 1.0,
               show.legend = FALSE) +
  labs(x = "Group",
       y = sCD14_axis) +
  scale_x_discrete(labels = c("Non-EHI", "Severe EHI")) +
  scale_y_continuous(breaks = seq(0, 1.75, .25),
                     limits = c(0, 1.8),
                     labels = scales::label_number(accuracy = 0.01)) +
  scale_color_manual(values = group_cols) +
  geom_segment(aes(x = 1, xend = 2, y = y_bracket_scd14, yend = y_bracket_scd14), inherit.aes = FALSE) +
  geom_segment(aes(x = 1, xend = 1, y = y_bracket_scd14, yend = y_bracket_scd14 - leg_drop_scd14), inherit.aes = FALSE) +
  geom_segment(aes(x = 2, xend = 2, y = y_bracket_scd14, yend = y_bracket_scd14 - leg_drop_scd14), inherit.aes = FALSE) +
  annotate("text",
           x = 1.5,
           y = y_bracket_scd14 + text_raise,
           label = pval,
           parse = TRUE,
           size = 4,
           family = "sans") +
  theme_ag()

#view figure
sCD14_modsev_p
```

# Sensitivity Analysis

## Data

Now create the two data frames for serum and EDTA analysis as done previously. Remove NSAID users and their respective matched-pair UIN.    
Serum:

```{r sen_data_serum}
## Create data frame for serum sensitivity analysis
sen_data_serum <- data_serum

## Create object which contains NSAID users
NSAID_users_serum <- sen_data_serum %>% 
  ## Filter rows where NSAID is "Yes"
  dplyr::filter(NSAID == "Yes")

## Extract unique Pair values
NSAID_pairs_serum <- unique(NSAID_users_serum$Pair)

## Count number of NSAID users
length(unique(NSAID_users_serum$UIN))

## Create serum data frame for sensivitivty analysis
sen_data_serum <- sen_data_serum %>%
  ## Remove NSAID pairs from data
  dplyr::filter(!Pair %in% NSAID_pairs_serum)

## Updated numbers:
## Cohort number
length(unique(sen_data_serum$UIN))

## Number per group (i.e. control/EHI)
length(unique(sen_data_serum$Pair))

## Check modsev numbers
table(sen_data_serum$ModSev)
```

EDTA:

```{r sen_data_EDTA}
## Create data frame for EDTA sensitivity analysis
sen_data_EDTA <- data_EDTA

## Create object which contains NSAID users
NSAID_users_EDTA <- sen_data_EDTA %>% 
  ## Filter rows where NSAID is "Yes"
  dplyr::filter(NSAID == "Yes")

## Extract unique Pair values
NSAID_pairs_EDTA <- unique(NSAID_users_EDTA$Pair)

## Count number of NSAID users
length(unique(NSAID_users_EDTA$UIN))

## Create EDTA data frame for sensivitivty analysis
sen_data_EDTA <- sen_data_EDTA %>%
  ## Remove NSAID pairs from data
  dplyr::filter(!Pair %in% NSAID_pairs_EDTA)

## Updated numbers:
## Cohort number
length(unique(sen_data_EDTA$UIN))

## Number per group (i.e. control/EHI)
length(unique(sen_data_EDTA$Pair))

## Check modsev numbers
table(sen_data_EDTA$ModSev)
```

As part of sensitivity analysis, this section will analyse the blood biomarkers of EHI cases vs their pair matched controls, where NSAID users and their respective pair are excluded. The N is presented in the descriptive section of each biomarker. 

## Intestinal Fatty Acid Binding Protein

### Descriptives

```{r sen_IFABP_descriptives_modsev}
jmv::descriptives(
    formula = IFABP ~ ModSev,
    data = sen_data_serum,
    iqr = TRUE,
    sw = TRUE)
```

### Test of Difference

non-parametric data:

```{r sen_IFABP_non-parametric_modsev}
jmv::ttestIS(
    formula = IFABP ~ ModSev,
    data = sen_data_serum,
    vars = IFABP,
    students = FALSE,
    mann = TRUE,
    norm = TRUE,
    effectSize = TRUE)
```

### Figure

``` {r sen_IFABP_modsev_fig, fig.height = 5, fig.width = 6}
set.seed(123)
sen_IFABP_modsev_p <- ggplot(sen_data_serum, aes(x = ModSev, y = IFABP)) +
  geom_jitter(aes(color = ModSev), 
              width = 0.3,
              height = 0,
              alpha = 0.6,
              size = 3) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  stat_summary(fun = median, 
               geom="crossbar", 
               color="black",
               width = 0.6,
               size = 0.6,
               show.legend = FALSE) +
  stat_summary(fun.data = median_iqr,
               geom="errorbar",
               color="black",
               width = 0.3,
               size = 0.8,
               show.legend = FALSE) +
  labs(x = "Group",
       y = IFABP_axis) +
  scale_x_discrete(labels = c("Non-EHI", "Mod/Sev EHI")) +
  scale_y_continuous(breaks = seq(-2000, 20000, 2000),
                     limits = c(-550, 20000)) +
  scale_color_manual(values = group_cols) +
  stat_compare_means(comparisons = list(c("EHI", "Control")),
                     method = "wilcox.test",
                     label = "p.signif",
                     family = "sans",
                     paired = FALSE,
                     bracket.size = 0.6,
                     vjust = -0.2) +
  theme_ag()

#view figure
sen_IFABP_modsev_p
```

## Claudin 3

### Descriptives

```{r sen_CLDN3_descriptives_modsev}
jmv::descriptives(
    formula = CLDN3 ~ ModSev,
    data = sen_data_EDTA,
    iqr = TRUE,
    sw = TRUE)
```

### Test of Difference

parametric data:

```{r sen_CLDN3_non-parametric_modsev}
jmv::ttestIS(
    formula = CLDN3 ~ ModSev,
    data = sen_data_EDTA,
    vars = CLDN3,
    norm = TRUE,
    eqv = TRUE,
    effectSize = TRUE)
```

### Figure

``` {r sen_CLDN3_modsev_fig, fig.height = 5, fig.width = 6}
set.seed(123)
sen_CLDN3_modsev_p <- ggplot(sen_data_EDTA, aes(x = ModSev, y = CLDN3)) +
  geom_jitter(aes(color = ModSev),
              width = 0.3,
              height = 0,
              alpha = 0.6,
              size = 3) +
  stat_summary(fun = mean,
               geom = "point",
               color = "black",
               size = 5,
               show.legend = FALSE) +
  stat_summary(fun.data = mean_sd,
               geom = "errorbar",
               color = "black",
               width = 0.3,
               size = 1.0,
               show.legend = FALSE) +
  labs(x = "Group",
       y = CLDN3_axis) +
  scale_x_discrete(labels = c("Non-EHI", "Mod/Sev EHI")) +
  scale_y_continuous(breaks = seq(0, 7, 1),
                     limits = c(0, 7),
                     labels = scales::label_number(accuracy = 1)) +
  scale_color_manual(values = group_cols) +
  stat_compare_means(comparisons = list(c("EHI", "Control")),
                     method = "t.test",
                     label = "p.signif",
                     family = "sans",
                     paired = FALSE,
                     bracket.size = 0.6,
                     vjust = -0.2) +
  theme_ag()

#view figure
sen_CLDN3_modsev_p
```

## Zonulin

### Descriptives

```{r sen_Zonulin_descriptives_modsev}
jmv::descriptives(
    formula = Zonulin ~ ModSev,
    data = sen_data_serum,
    iqr = TRUE,
    sw = TRUE)
```

### Test of Difference

non-parametric data:

```{r sen_Zonulin_non-parametric_modsev}
jmv::ttestIS(
    formula = Zonulin ~ ModSev,
    data = sen_data_serum,
    vars = Zonulin,
    norm = TRUE,
    eqv = TRUE,
    effectSize = TRUE)
```

### Figure

``` {r sen_Zonulin_modsev_fig, fig.height = 5, fig.width = 6}
set.seed(123)
sen_Zon_modsev_p <- ggplot(sen_data_serum, aes(x = ModSev, y = Zonulin)) +
  geom_jitter(aes(color = ModSev),
              width = 0.3,
              height = 0,
              alpha = 0.6,
              size = 3) +
  stat_summary(fun = mean,
               geom = "point",
               color = "black",
               size = 5,
               show.legend = FALSE) +
  stat_summary(fun.data = mean_sd,
               geom = "errorbar",
               color = "black",
               width = 0.3,
               size = 1.0,
               show.legend = FALSE) +
  labs(x = "Group",
       y = Zon_axis) +
  scale_x_discrete(labels = c("Non-EHI", "Mod/Sev EHI")) +
  scale_y_continuous(breaks = seq(0, 60, 10),
                     limits = c(0, 60)) +
  scale_color_manual(values = group_cols) +
  stat_compare_means(comparisons = list(c("EHI", "Control")),
                     method = "t.test",
                     label = "p.signif",
                     family = "sans",
                     paired = FALSE,
                     bracket.size = 0.6,
                     vjust = -0.2) +
  theme_ag()

#view figure
sen_Zon_modsev_p
```

## Lipopolysaccharide Binding Protein

### Descriptives

```{r sen_LBP_descriptives_modsev}
jmv::descriptives(
    formula = LBP ~ ModSev,
    data = sen_data_EDTA,
    iqr = TRUE,
    sw = TRUE)
```

### Test of Difference

parametric data:

```{r sen_LBP_parametric_class}
jmv::ttestIS(
    formula = LBP ~ ModSev,
    data = sen_data_EDTA,
    vars = LBP,
    norm = TRUE,
    eqv = TRUE,
    effectSize = TRUE)
```

### Figure

``` {r sen_LBP_modsev_fig, fig.height = 5, fig.width = 6}
set.seed(123)
sen_LBP_modsev_p <- ggplot(sen_data_EDTA, aes(x = ModSev, y = LBP)) +
  geom_jitter(aes(color = ModSev), 
              width = 0.3,
              height = 0,
              alpha = 0.6,
              size = 3) +
  stat_summary(fun = mean, 
               geom = "point", 
               color = "black",
               size = 5,
               show.legend = FALSE) +
  stat_summary(fun.data = mean_sd, 
               geom = "errorbar", 
               color = "black",
               width = 0.3,
               size = 1.0,
               show.legend = FALSE) +
  labs(x = "Group",
       y = LBP_axis) +
  scale_x_discrete(labels = c("Non-EHI", "Mod/Sev EHI")) +
  scale_y_continuous(breaks = seq(0.0, 6.0, 1.0),
                     limits = c(0.0, 6.5),
                     labels = scales::label_number(accuracy = 1)) +
  scale_color_manual(values = group_cols) +
  stat_compare_means(comparisons = list(c("EHI", "Control")),
                     method = "t.test",
                     label = "p.signif",
                     family = "sans",
                     paired = FALSE,
                     bracket.size = 0.6,
                     vjust = -0.2) +
  theme_ag()

#view figure
sen_LBP_modsev_p
```

## Soluble Cluster of Differentiation 14

### Descriptives

```{r sen_sCD14_descriptives_modsev}
jmv::descriptives(
    formula = sCD14 ~ ModSev,
    data = sen_data_EDTA,
    iqr = TRUE,
    sw = TRUE)
```

### Test of Difference

parametric data:

```{r sen_sCD14_parametric_modsev}
jmv::ttestIS(
    formula = sCD14 ~ ModSev,
    data = sen_data_EDTA,
    vars = sCD14,
    norm = TRUE,
    eqv = TRUE,
    effectSize = TRUE)
```

### Figure

``` {r sen_sCD14_modsev_fig, fig.height = 5, fig.width = 6}
set.seed(123)
sen_sCD14_modsev_p <- ggplot(sen_data_EDTA, aes(x = ModSev, y = sCD14)) +
  geom_jitter(aes(color = ModSev), 
              width = 0.3,
              height = 0,
              alpha = 0.6,
              size = 3) +
  stat_summary(fun = mean, 
               geom = "point", 
               color = "black",
               size = 5,
               show.legend = FALSE) +
  stat_summary(fun.data = mean_sd, 
               geom = "errorbar", 
               color = "black",
               width = 0.3,
               size = 1.0,
               show.legend = FALSE) +
  labs(x = "Group",
       y = sCD14_axis) +
  scale_x_discrete(labels = c("Non-EHI", "Mod/Sev EHI")) +
  scale_y_continuous(breaks = seq(0, 1.75, .25),
                     limits = c(0, 1.75),
                     labels = scales::label_number(accuracy = 0.01)) +
  scale_color_manual(values = group_cols) +
  stat_compare_means(comparisons = list(c("EHI", "Control")),
                     method = "t.test",
                     label = "p.signif",
                     family = "sans",
                     paired = FALSE,
                     bracket.size = 0.6,
                     vjust = -0.2) +
  theme_ag()

#view figure
sen_sCD14_modsev_p
```

# Figure Panel

Create a panel of figures.    
  
Note: wrap plots made with `scale_y_break()` from {ggbreak} in "print(PLOT_NAME_HERE)"

```{r biomarkers_panel, fig.height = 9, fig.width = 12}
## arrange by biomarker type
## Enterocyte damage & tight junction integrity
row1 <- ggarrange(IFABP_modsev_p, 
                  Zon_modsev_p, 
                  CLDN3_modsev_p, 
                  ncol = 3, 
                  labels = paste0("(", letters[1:3], ")"),
                  font.label = list(size = 14, 
                                    face = "bold",
                                    family = "sans"))
## Microbial translocation
row2 <- ggarrange(LBP_modsev_p, 
                  sCD14_modsev_p,
                  ncol = 3, 
                  labels = paste0("(", letters[4:5], ")"), 
                  font.label = list(size = 14, 
                                    face = "bold",
                                    family = "sans"))
## Combine rows
biomarkers_modsev_panel <- ggarrange(row1, row2, nrow=2)
## Show panel
biomarkers_modsev_panel
```

Add annotation: 

```{r biomarkers_annotation_panel, fig.height = 9, fig.width = 13}
## Annotate row1
row1ann <- annotate_figure(row1,
                           left = text_grob("Enterocyte damage &\ntight junction integrity\n", 
                                            rot = 90, 
                                            size = 16, 
                                            face = "bold",
                                            family = "sans"))

## Annotate row2
row2ann <- annotate_figure(row2,
                           left = text_grob("\nMicrobial translocation\n", 
                                            rot = 90, 
                                            size = 16, 
                                            face = "bold",
                                            family = "sans"))
## Combine rows
biomarkers_modsev_annotation_panel <- ggarrange(row1ann, row2ann, nrow=2)
## Show panel
biomarkers_modsev_annotation_panel

## Save
ggsave(
  filename = paste0(
    analysis_dir,
    "figs/biomarkers_sev_annotation_panel",
    ".tiff"
  ),
  plot = biomarkers_modsev_annotation_panel,
  height = 9,
  width = 13,
  unit = "in",
  dpi = 600
)

## Save
ggsave(
  filename = paste0(
    analysis_dir,
    "figs/biomarkers_sev_annotation_panel",
    ".png"
  ),
  plot = biomarkers_modsev_annotation_panel,
  height = 9,
  width = 13,
  unit = "in",
  dpi = 600
)

## Save
ggsave(
  filename = paste0(
    analysis_dir,
    "figs/biomarkers_sev_annotation_panel",
    ".pdf"
  ),
  plot = biomarkers_modsev_annotation_panel,
  height = 9,
  width = 13,
  unit = "in",
  dpi = 600
)
```

# Session info

``` {r sessionInfo}
sinfo <- sessionInfo()
sinfo
save.image(paste0(analysis_dir, "/prospective_analysis_manuscript_biomarkers_modsev.Rout"))
```