---
title: "Participant characteristics (manuscript)"
author: "Alex Gould"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: yes
    toc_depth: 3
    number_sections: yes
    latex_engine: xelatex
  rmarkdown::pdf_document:
    extra_dependencies: float
    toc: yes
html_document:
  toc: yes
toc_float: yes
---

# Setup
First let us set up the R environment:

## Global options

```{r global options, include = TRUE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      error = FALSE,
                      eval = TRUE,
                      dev = c('pdf', 'png'),
                      width = 60,
                      fig.align = 'center',
                      fig.height = 10,
                      fig.width = 10,
                      fig.dpi = 300,
                      fig.pos = "!H",
                      dpi = 300,
                      out.extra = "",
                      message = FALSE,
                      cache = FALSE,
                      pdf.options(encoding = "ISOLatin9.enc"),
                      fig.path='figs/rmd/')
```

## Library

```{r setup}
## Load the libraries
library(xlsx)
library(rstudioapi)
library(tidyverse)
library(jmv)
library(ggtext)
library(ggpubr)
library(scales)
```

## Parameters

```{r directory}
data_dir          = '...'
analysis_dir      = '...'
pw_message        = "..."
font              = element_text(family = "serif")
group_col         = c(
  "Control"       = "#1B9E77",
  "Mild"          = "#DDDD77" ,
  "Moderate"      = "#D95F02",
  "Severe"        = "#771122"
  )
```

### Figure Parameters

```{r figure_theme}
## create custom theme
theme_ag <- function(){

  theme_classic() %+replace%
    
    theme(
      text = font,
      axis.text.y  = element_text(size = 12, margin = margin(r = 3), colour = "black"),
      axis.title.y = element_text(size = 16, angle = 90, vjust = 2),
      axis.text.x  = element_text(size = 16, margin = margin(r = 3), colour = "black"),
      axis.title.x = element_blank(),
      legend.title = element_text(size = 16),
      legend.text = element_text(size = 14)
    )
}
```

Now, make custom axis titles to use throughout:

```{r figure_custom_y_axis}
tgi_axis = expression(paste("Peak ", italic("T")*""[gi] ~ "(\u00b0C)"))
qcs_axis = "QCS score"
```

Fix for `stat_compare_means()` function to allow Times New Roman font. Thank you: https://stackoverflow.com/questions/77346644/stat-compare-means-unable-to-change-font-family-when-using-comparisons-argu 

```{r stat_compare_means_fix}
fixed_call <- quote(ggsignif::geom_signif(comparisons = comparisons,
                                          y_position = label.y,
                                          test = method, test.args = method.args,
                                          step_increase = step.increase,
                                          size = bracket.size, textsize = 6, color = color,
                                          map_signif_level = map_signif_level,
                                          tip_length = tip.length, data = data,
                                          vjust = vjust, ...))

scm <- as.list(ggpubr::stat_compare_means)
scm[[26]][[2]][[3]][[13]] <- fixed_call
stat_compare_means <- function(...) {
  ggp <- getNamespace('ggpubr')
  .method_info <- ggp$.method_info
  .add_item <- ggp$.add_item
  .is_p.signif_in_mapping <- ggp$.is_p.signif_in_mapping
  .is_empty <- ggp$.is_empty
  do.call(as.function(scm), list(...), quote = FALSE)
}
```

## Load Data

Next we load in the data:

```{r load_data}
## load data here... 
```

# Matched Cohort Characteristics

## Control vs EHI

### Descriptives

```{r descriptives_case_anthro}
jmv::descriptives(
    formula = Age + Height + Mass + BMI ~ case,
    data = dat_casecon,
    iqr = TRUE,
    sw = TRUE)
```

```{r descriptives_case_bsa}
jmv::descriptives(
    formula = BSA + BSA_M ~ case,
    data = dat_casecon,
    iqr = TRUE,
    sw = TRUE)
```

```{r descriptives_case_fitness_temp}
jmv::descriptives(
    formula = RMFT_s + PT_rank + BFT_s + Tcore ~ case,
    data = dat_casecon,
    iqr = TRUE,
    sw = TRUE)
```

```{r descriptives_qcs}
jmv::descriptives(
    formula = QCS ~ case,
    data = dat_casecon,
    iqr = TRUE,
    sw = TRUE)
```

### Test of Difference

parametric data:

```{r parametric_Case}
jmv::ttestIS(
    formula = Mass + BMI + BSA_M + RMFT_s + BFT_s ~ case,
    data = dat_casecon,
    vars = vars(Mass, BMI, BSA_M, RMFT_s, BFT_s),
    effectSize = TRUE,
    norm = TRUE,
    eqv = TRUE)
```

non-parametric data:

```{r non-parametric_Case}
jmv::ttestIS(
    formula = Age + Height + BSA + PT_rank + Tcore + QCS ~ case,
    data = dat_casecon,
    vars = vars(Age, Height, BSA, PT_rank, Tcore, QCS),
    effectSize = TRUE,
    students = FALSE,
    mann = TRUE)
```

## Moderate and Severes Combined

### Descriptives

```{r descriptives_modsev_anthro}
jmv::descriptives(
    formula = Age + Height + Mass + BMI ~ modsev,
    data = dat_casecon,
    iqr = TRUE,
    sw = TRUE)
```

```{r descriptives_modsev_bsa}
jmv::descriptives(
    formula = BSA + BSA_M ~ modsev,
    data = dat_casecon,
    iqr = TRUE,
    sw = TRUE)
```

```{r descriptives_modsev_fitness_temp}
jmv::descriptives(
    formula = RMFT_s + PT_rank + BFT_s + Tcore ~ modsev,
    data = dat_casecon,
    iqr = TRUE,
    sw = TRUE)
```

```{r descriptives_modsev_qcs}
jmv::descriptives(
    formula = QCS ~ modsev,
    data = dat_casecon,
    iqr = TRUE,
    sw = TRUE)
```

### Test of Difference

parametric data:

```{r parametric_modsev}
jmv::ttestIS(
    formula = Height + Mass + BMI + BSA + BSA_M + BFT_s ~ modsev,
    data = dat_casecon,
    vars = vars(Height, Mass, BMI, BSA, BSA_M, BFT_s),
    effectSize = TRUE,
    norm = TRUE,
    eqv = TRUE)
```

non-parametric data:

```{r non-parametric_modsev}
jmv::ttestIS(
    formula = Age + RMFT_s + PT_rank + Tcore + QCS ~ modsev,
    data = dat_casecon,
    vars = vars(Age, RMFT_s, PT_rank, Tcore, QCS),
    effectSize = TRUE,
    students = FALSE,
    mann = TRUE)
```

### Tgi ModSev vs Mild cases

Below, compare the Tgi of ModSev vs. mild EHI cases.

```{r modsev_vs_mild_cases}
## create df of only mild cases
mild_cases <- dat_casecon %>% 
  dplyr::select(Pair, UIN, Troop, classification2, Tcore) %>% 
  dplyr::filter(classification2 == "Mild") %>%
  ## create new column called Group
  dplyr::mutate(Group = "Mild") %>% 
  dplyr::select(-classification2)

## create df of only modsev cases
modsev_cases <- dat_casecon %>% 
  dplyr::select(Pair, UIN, Troop, modsev, Tcore) %>% 
  dplyr::filter(modsev == "EHI") %>% 
  ## create new column called Group
  dplyr::mutate(Group = "EHI") %>% 
  dplyr::select(-modsev)

## combine the two dfs
mild_modsev <- dplyr::bind_rows(mild_cases, modsev_cases)

## descriptives
jmv::descriptives(
    formula = Tcore ~ Group,
    data = mild_modsev,
    iqr = TRUE,
    skew = TRUE,
    kurt = TRUE,
    sw = TRUE)

## t-test (Mann-Whitney U)
jmv::ttestIS(
    formula = Tcore ~ Group,
    data = mild_modsev,
    vars = Tcore,
    students = FALSE,
    mann = TRUE)
```

# Figure

First, create a data frame to use for figure plotting.

```{r fig_dataframe}
dat <- dat_casecon %>%
  dplyr::select(case, classification2, Tcore, modsev) %>%
  dplyr::mutate(classification2 = case_when(
    classification2 %in% c("Mild Control", 
                           "Moderate Control", 
                           "Severe Control") ~ "Control",
    TRUE ~ classification2)) %>%
  dplyr::rename(Classification = classification2) %>%
  dplyr::mutate(Classification = factor(Classification, levels = c("Control", 
                                                                   "Mild",
                                                                   "Moderate",
                                                                   "Severe"))) 
```

How many controls have a Tcore > 40°C? 0

```{r}
dat %>%
  dplyr::select(case, Tcore) %>%
  dplyr::filter(case == "Control") %>%
  dplyr::filter(Tcore > 40)
```

How many mild EHIs have a Tcore > 40°C? 5

```{r}
dat %>%
  dplyr::select(Classification, Tcore) %>%
  dplyr::filter(Classification == "Mild") %>%
  dplyr::filter(Tcore > 40)
```

How many severe EHIs have a Tcore < 40°C? 1

```{r}
dat %>%
  dplyr::select(Classification, Tcore) %>%
  dplyr::filter(Classification == "Severe") %>%
  dplyr::filter(Tcore < 40)
```

## Peak Gastrointestinal Temperature

```{r fig_peak_gi_temp, fig.height = 5, fig.width = 7}
set.seed(123)
Tcore_case_p <- 
  dat %>%
  ggplot(aes(x = case, y = Tcore)) +
  geom_boxplot(outlier.colour = NA) +
  geom_jitter(aes(color = Classification), 
              width = 0.3,
              alpha = 0.6,
              size = 2
              ) +
  labs(x = "Group",
       y = tgi_axis) +
  scale_x_discrete(labels = c("Non-EHI", "EHI")) +
  scale_y_continuous(breaks = seq(37, 43, 1), 
                     limits = c(37, 43),
                     labels = scales::label_number(accuracy = 0.1)) +
  scale_color_manual(values = group_col,
                     labels = c("Non-EHI", "Mild", "Moderate", "Severe")) +
  stat_compare_means(comparisons = list(c("EHI", "Control")),
                     method = "wilcox.test",
                     label = "p.signif",
                     family = "serif",
                     paired = FALSE,
                     bracket.size = 0.6,
                     vjust = -0.2) +
  theme_ag()

## view
Tcore_case_p

ggsave(
  filename = paste0(analysis_dir, "figs/ggsave_peak_tgi_figure", ".png"),
  plot = Tcore_case_p,
  height = 5,
  width = 7,
  unit = "in",
  dpi = 600
)

ggsave(
  filename = paste0(analysis_dir, "figs/ggsave_peak_tgi_figure", ".jpeg"),
  plot = Tcore_case_p,
  height = 5,
  width = 7,
  unit = "in",
  dpi = 600
)
```

# Session info

``` {r sessionInfo}
sinfo <- sessionInfo()
sinfo
save.image(paste0(analysis_dir, "/prospective_analysis_manuscript_characteristics.Rout"))
```
