---
title: "16S rRNA Analysis - Case/Control (manuscript)"
author: "Alex Gould"
date: "`r date()`"
output:
  pdf_document:
    toc: yes
    toc_depth: 4
    number_sections: no
    latex_engine: xelatex
  rmarkdown::pdf_document:
    extra_dependencies: float
    toc: yes
html_document:
  toc: yes
toc_float: yes
---

# Section 0: Setup

## Global options

```{r global options, include = TRUE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE,
                      error = FALSE,
                      eval = TRUE,
                      dev = c('pdf', 'png'),
                      fig.path='figs/rmd/',
                      width = 60,
                      fig.align = 'center', 
                      fig.height = 10, 
                      fig.width = 10, 
                      fig.dpi = 300,
                      fig.pos = "!H",
                      dpi = 300,
                      out.extra = "",
                      message = FALSE, 
                      cache = FALSE,
                      pdf.options(encoding = "ISOLatin9.enc")) 
```

## Parameters

``` {r SET_parameters, eval = TRUE}
analysis_dir     = "..."
qiime_dir        = "..."
data_dir         = "..."
pw_message       = "...."
experiment_type  = 'illumina'               
load_type        = 'input'
experiment_code  = '...'
process_num      = '1'
indir            = "..."
min_otus         = 20000
nclass           = 15 
group_cols       = c('Control' = '#0099CC', 'EHI' = '#FF9900')
custom_col16     = c(
  "#771155",
  "#AA4488",
  "#CC99BB",
  "#114477",
  "#4477AA",
  "#117777",
  "#77CCCC",
  "#117744",
  "#88CCAA",
  "#777711",
  "#AAAA44",
  "#DDDD77",
  "#774411",
  "#AA7744",
  "#DDAA77",
  "#771122"
)
test_type        = "non-parametric"
control          = "Control"
otu_or_asv       = "ASV"
read_type        = "filter"
rep_shapes       = rep(c(15, 16, 17, 18, 3, 7, 8, 9, 10, 11, 12, 13), 7)
```

## Environment setup
First check that the input parameters are suitable:

``` {r check_parameters}
if (!file.exists(analysis_dir)) stop("Error: Output directory ", analysis_dir, " not found. Please ensure that the output directory exists before running.\n")
if (!experiment_type %in% c("illumina", "nanopore")) stop ("Error: experiment_type must be one of 'illumina' or 'nanopore'.\n")
if (!load_type %in% c("pipeline", "input")) stop ("Error: load_type must be one of 'pipeline' or 'input'.\n")
if(load_type == "pipeline" && (is.na(experiment_code) || is.na(process_num))) stop("Error: Parameters 'experiment_code' and 'process_num' are required if load_type 'pipeline' used.\n")
if(load_type == "input" && (is.na(indir))) stop("Error: Parameters 'experiment_code' and 'process_num' are required if load_type 'pipeline' used.\n")
if(load_type == "input" && (!file.exists(indir))) stop("Error: Input directory ", indir, " not found. Please ensure that the input directory containing the required input files is correct before running.\n")
if (nclass < 5 || nclass > 20) stop("Error: nclass must be between 5 and 20.\n")
if (min_otus < 5000) stop("Error: min_otus must be greater than 5000.\n")
if (!test_type %in% c("parametric", "non-parametric")) stop ("Error: test_type must be one of 'parametric' or 'non-parametric'.\n")
if (!read_type %in% c("raw", "trim", "filter")) stop ("Error: read_type must be one of 'raw', 'trim' or 'filter'.\n")
###!!!! Check group_cols and control
if(!file.exists(paste0(analysis_dir, "/tabs"))) dir.create(paste0(analysis_dir, "/tabs"))
```

Then set up the R environment:

```{r setup}
## Load the libraries
library("DBI")
library("RMySQL")
library("phyloseq")
library("tidyverse")
library("xlsx")
library("rstudioapi")
library("qiime2R")
library("jmv")
library("ggpubr")
library("ggtext")
library("vegan")
library("plyr")
library("reshape2")
library("DESeq2")
library("broom")
library("scales")
library("corrplot")
library("ggrepel")
library("ggvenn")
library("rlang")
library("ALDEx2")
library("ANCOMBC")
library("openxlsx")

## Set seed for reproducibility
set.seed(0)

## Set font for figures
font = element_text(family = "serif")
```

## stat_compare_means() fix

Fix for `stat_compare_means()` function to allow Times New Roman font. Thank you https://stackoverflow.com/questions/77346644/stat-compare-means-unable-to-change-font-family-when-using-comparisons-argu 

```{r stat_compare_means_fix}
fixed_call <- quote(ggsignif::geom_signif(comparisons = comparisons, 
                                          y_position = label.y, 
                                          test = method, test.args = method.args, 
                                          step_increase = step.increase, 
                                          size = bracket.size, textsize = 6, color = color, 
                                          map_signif_level = map_signif_level, 
                                          tip_length = tip.length, data = data,
                                          vjust = vjust, ...))

scm <- as.list(ggpubr::stat_compare_means)
scm[[26]][[2]][[3]][[13]] <- fixed_call
stat_compare_means <- function(...) {
  ggp <- getNamespace('ggpubr')
  .method_info <- ggp$.method_info
  .add_item <- ggp$.add_item
  .is_p.signif_in_mapping <- ggp$.is_p.signif_in_mapping
  .is_empty <- ggp$.is_empty
  do.call(as.function(scm), list(...), quote = FALSE)
}
```

## Load data

### Sample annotation

``` {r load_sample_ann}
## load sample annotation here...
```

### Set sample annotation

``` {r SET_sample_ann}
## set sample annotation here...
```

### Phyloseq

``` {r load_phyloseq_data}
## load phyloseq object here...
```

# Section 1: Quality Control - all data

## Rarefaction curve
We can check our coverage to be sure that we have ample coverage to allow us to detect all species present in these samples. We estimate this by using rarefaction curves, which show the diversity when looking at random subsets of the data. The idea here is that if we have saturated the sample, we should see the same diversity if we only look at a subset of the sample. By looking at the curve of successive sub-populations, we can see at what coverage we see saturation. This will appear as the point at which the curve becomes assymptotic:

``` {r Section1_QC_alldata_rarefaction_curve, cache = TRUE}
## Calculate a rarefaction curve (code taken from https://github.com/joey711/phyloseq/issues/143)
calculate_rarefaction_curves <- function(psdata, measures, depths) {

  estimate_rarified_richness <- function(psdata, measures, depth) {
    if(max(sample_sums(psdata)) < depth) return()
    psdata <- prune_samples(sample_sums(psdata) >= depth, psdata)

    rarified_psdata <- rarefy_even_depth(psdata, depth, verbose = FALSE)

    alpha_diversity <- estimate_richness(rarified_psdata, measures = measures)

    ## as.matrix forces the use of melt.array, which includes the Sample names (rownames)
    molten_alpha_diversity <- melt(as.matrix(alpha_diversity), varnames = c('Sample', 'Measure'), value.name = 'Alpha_diversity')

    molten_alpha_diversity
  }

  names(depths) <- depths ## this enables automatic addition of the Depth to the output by ldply
  rarefaction_curve_data <- ldply(depths, estimate_rarified_richness, psdata = psdata, measures = measures, .id = 'Depth', .progress = ifelse(interactive(), 'text', 'none'))

  ## convert Depth from factor to numeric
  rarefaction_curve_data$Depth <- as.numeric(levels(rarefaction_curve_data$Depth))[rarefaction_curve_data$Depth]

  rarefaction_curve_data
}

## Calculate the rarefaction curve
set.seed(42)
rarefaction_curve_data <- calculate_rarefaction_curves(phyloseq_dat,
                                                       c('Observed', "Chao1", 'Shannon', "Simpson"),
                                                       rep(c(1, 10, 100, 1000, 1:100 * 1000), each = 10))

## Summarize the alpha diversity
rarefaction_curve_data_summary <- ddply(rarefaction_curve_data, c('Depth', 'Sample', 'Measure'),
                                        summarise,
                                        Alpha_diversity_mean = mean(Alpha_diversity, na.rm = TRUE),
                                        Alpha_diversity_sd = sd(Alpha_diversity, na.rm = TRUE))
rarefaction_curve_data_summary [["Sample"]] <- gsub("\\.", " ", rarefaction_curve_data_summary [["Sample"]])
rarefaction_curve_data_summary_verbose <- merge(rarefaction_curve_data_summary, data.frame(sample_data(phyloseq_dat)), by.x = 'Sample', by.y = 'row.names')
rarefaction_curve_data_summary_verbose[["Depth"]] <- rarefaction_curve_data_summary_verbose[["Depth"]]/1000
```

### Plot 

``` {r Section1_QC_alldata_rarefaction_curve_plot, cache = TRUE, fig.height = 10, fig.width = 15}
## Plot the rarefaction curve
rarefaction_curve <-
  ggplot(data = subset(rarefaction_curve_data_summary_verbose, Measure %in% c('Observed', "Chao1", 'Shannon', "Simpson")),
                mapping = aes(
                x = Depth,
                y = Alpha_diversity_mean,
                ymin = Alpha_diversity_mean - Alpha_diversity_sd,
                ymax = Alpha_diversity_mean + Alpha_diversity_sd,
                colour = Group,
                #shape  = Rep,
                #group  = Rep,
                alpha = 0.8)) +
  geom_line() +
  geom_pointrange(size = 1) +
  xlab("Read Depth (1,000s)") +
  ylab("Alpha Diversity") +
  scale_fill_manual(values = group_cols) +
  scale_color_manual(values = group_cols) +
  scale_shape_manual(values = rep_shapes) +
  facet_grid(Measure ~ ., scales="free_y") +
  theme_bw() +
  theme(
    text = font,
    axis.title = element_text(size = 24),
    axis.text = element_text(size = 18),
    legend.text = element_text(size = 18),
    legend.title = element_text(size=18),
    strip.text.x = element_text(size=24, face = "bold"),
    strip.text.y = element_text(size=24, face = "bold"))

rarefaction_curve

## reset seed to 0
set.seed(0)
```

## Taxa count per sample - All data

The following plot shows the total number of observed taxa (OTUs or ASVs) per sample, with the minimum number of taxa specified above (min_otus) highlighted with a dashed line. Samples lower than this will be removed below in Section 2. 

``` {r Section1_QC_alldata_OTU_count, fig.height = 12, fig.width = 10}
data.frame(sample_data(phyloseq_dat)) %>%
  dplyr::mutate(count = sample_sums(phyloseq_dat)) %>%
  ggplot(aes(x = UIN, y = count/1000, fill = Group)) +
  geom_bar(stat = "identity", position = "dodge", color = "black") +
  xlab("") + ylab(paste0("Total Number ", otu_or_asv, "s (x1000)")) +
  facet_wrap(.~Group, scales = "free_x") +
  geom_hline(yintercept = min_otus/1000, linetype = 2, linewidth = 1.2) +
  scale_fill_manual(values = group_cols) +
  scale_color_manual(values = group_cols) +
  scale_shape_manual(values = rep_shapes) +
  theme_bw() +
  theme(
    text = font,
    axis.title   = element_text(size = 18),
    axis.text.y  = element_text(size = 12),
    axis.text.x  = element_text(size = 14, angle = 315, face = "bold", hjust = 0),
    legend.text  = element_text(size = 14),
    legend.title = element_text(size = 20, face = "bold"),
    strip.text.x = element_text(size=20, face = "bold"),
    strip.text.y = element_text(size=16, face = "bold")
  )
```

# Section 2: Quality Control - Rarefied and filtered data

## Sample filtering and rarefaction

In this section, the data will be filtered to remove all samples with fewer than `r min_otus` OTUs in the data set. The remaining data will be rarefied to reduce all samples to the same depth, which will be the lowest depth from the remaining samples following filtering.

``` {r Section2_QC_filterdata_filter_and_rarefy_samples}
## Check sample sums
sample_sums_values <- sample_sums(phyloseq_dat)

## Find samples with sums below 20,000
samples_below_min_otus <- names(sample_sums_values[sample_sums_values <= min_otus])

## Print
print(samples_below_min_otus)

## Now identify the associated pair to also remove
filter_pairs_to_remove <- sample_data(phyloseq_dat)$Pair[sample_data(phyloseq_dat)$sample_code %in% samples_below_min_otus]

## Print
print(filter_pairs_to_remove)

## Filter to remove samples with fewer than min_otus
phyloseq_dat_filter <- prune_samples(!(sample_data(phyloseq_dat)$Pair %in% filter_pairs_to_remove), phyloseq_dat)

## Create a non rarefied phyloseq object with filtered samples for DESeq2 analysis
phyloseq_dat_deseq <- prune_samples(!(sample_data(phyloseq_dat)$Pair %in% filter_pairs_to_remove), phyloseq_dat)

## Rarefy the data to the minimum otu depth
rarefy_otu_depth <- min(sample_sums(phyloseq_dat_filter))
cat("Rarefication of data to", rarefy_otu_depth, "OTUs:\n")
phyloseq_dat_filter <- rarefy_even_depth(phyloseq_dat_filter, rarefy_otu_depth, verbose = FALSE)
```

### Summary N for Analysis

Below, a summary of the n taken forward for analysis is provided: 

``` {r Section2_n_summary}
## N by group
sample_ann_summary %>%
  dplyr::select(UIN, sample_code, Pair, Classification, Group, NSAID) %>%  
  ## Remove Pos & Neg controls using reg expression
  dplyr::filter(!grepl(" Control", Group)) %>%
  ## Remove pairs which have ASV count lower than min_otus (< 20,000)
  dplyr::filter(!Pair %in% filter_pairs_to_remove) %>%
  dplyr::group_by(Group) %>%
  dplyr::summarise(n = n())

## Total N removed by having ASV count lower than min_outs (< 20,000)
length(filter_pairs_to_remove)

## By classification 
sample_ann_summary %>%
  dplyr::select(UIN, sample_code, Pair, Classification, Group, NSAID) %>%
  ## Remove Pos & Neg controls using reg expression
  dplyr::filter(!grepl(" Control", Group)) %>%
  ## Remove pairs which have ASV count lower than min_otus (< 20,000)
  dplyr::filter(!Pair %in% filter_pairs_to_remove) %>%
  dplyr::group_by(Classification) %>%
  dplyr::summarise(n = n())
```

## Taxa count per sample - Rarefied and filtered data

The following plot shows the total number of observed taxa (OTUs or ASVs) per sample after rarefaction and filtering of samples to remove low coverage samples, with the minimum number of taxa specified above (min_otus) highlighted with a dashed line. Samples lower than this will be removed below in Section 2. All samples should have the same depth. If this is not the case, then please debug the code.

``` {r Section2_QC_filterdata_OTU_count, fig.height = 8, fig.width = 10}
data.frame(sample_data(phyloseq_dat_filter)) %>%
  dplyr::mutate(count = sample_sums(phyloseq_dat_filter)) %>%
  ggplot(aes(x = UIN, y = count/1000, fill = Group)) +
  geom_bar(stat = "identity", position = "dodge", color = "black") +
  xlab("") + ylab(paste0("Total Number ", otu_or_asv, "s (x1000)")) +
  facet_wrap(.~Group, scales = "free_x") +
  geom_hline(yintercept = min_otus/1000, linetype = 2, linewidth = 1.2) +
  scale_fill_manual(values = group_cols) +
  scale_color_manual(values = group_cols) +
  scale_shape_manual(values = rep_shapes) +
  theme_bw() +
  theme(
    text = font,
    axis.title   = element_text(size = 18),
    axis.text.y  = element_text(size = 12),
    axis.text.x  = element_text(size = 14, angle = 315, face = "bold", hjust = 0),
    legend.text  = element_text(size = 14),
    legend.title = element_text(size = 20, face = "bold"),
    strip.text.x = element_text(size=20, face = "bold"),
    strip.text.y = element_text(size=16, face = "bold")
  )
```


## Alpha diversity - Rarefied and filtered data

The following will create a .csv file containing the observed, Chao1 and Shannon Alpha diversity metrics for each sample. Below is the median across each Group level.

```{r Section2_QC_filterdata_alpha_diversity_summary}
## Create alpha diversity data frame
adiv <- data.frame(
  "Pair" = phyloseq::sample_data(phyloseq_dat_filter)$Pair,
  "UIN" = phyloseq::sample_data(phyloseq_dat_filter)$UIN,
  "Group" = phyloseq::sample_data(phyloseq_dat_filter)$Group,
  "Troop" = phyloseq::sample_data(phyloseq_dat_filter)$Troop,
  "Observed" = phyloseq::estimate_richness(phyloseq_dat_filter, measures = "Observed"),
  "Chao1" = phyloseq::estimate_richness(phyloseq_dat_filter, measures = "Chao1"),
  "Shannon" = phyloseq::estimate_richness(phyloseq_dat_filter, measures = "Shannon"),
  "Simpson" = phyloseq::estimate_richness(phyloseq_dat_filter, measures = "Simpson"))

adiv <- adiv %>%
  ## Remove Chao1 SE column
  dplyr::select(-Chao1.se.chao1) %>%
  ## Rename Chao1 column
  dplyr::rename(Chao1 = Chao1.Chao1) %>%
  dplyr::arrange(Pair, Group) %>%
  tibble::remove_rownames()

head(adiv)

## write as csv
write.table(
  adiv,
  paste0(analysis_dir, "/tabs/alpha_diversity_table.csv"),
  row.names = FALSE,
  col.names = TRUE,
  quote = FALSE,
  sep = ","
)
``` 

### Normality

The following analysis will test for normality of the distribution of Alpha Diversity metrics, which will determine whether analyses should be performed using parametric or non-parametric tests. This table gives summary statistics for a variaty of tests related to the distribution of the data to helkp make this decision. However, generally speaking, if the p-value for the Shapiro-Wilk test is greater than 0.05, then one can assume that the assumption of normality will hold (i.e. there is not sufficient evidence to reject the null hypothesis).

```{r Section2_QC_filterdata_alpha_diversity_normality}
jmv::descriptives(
    formula = Observed + Chao1 + Shannon + Simpson ~ Group,
    data = adiv,
    iqr = TRUE,
    sw = TRUE)
```

### t-test

```{r Section2_QC_filterdata_alpha_diversity_comparison_stat}
jmv::ttestIS(
    formula = Shannon ~ Group,
    data = adiv,
    vars = vars(Shannon),
    eqv = TRUE,
    effectSize = TRUE)
```

For Shannon, Levene's test is violated (p < 0.05) so we will now use a Welch t-test: 

```{r Section2_QC_filterdata_alpha_diversity_comparison_stat_welch}
jmv::ttestIS(
    formula = Shannon ~ Group,
    data = adiv,
    vars = vars(Shannon),
    welchs = TRUE,
    students = FALSE,
    eqv = TRUE,
    effectSize = TRUE)
```


```{r Section2_QC_filterdata_alpha_diversity_comparison_stat_non_parametric}
jmv::ttestIS(
    formula = Observed + Chao1 + Simpson ~ Group,
    data = adiv,
    vars = vars(Observed, Chao1, Simpson),
    students = FALSE,
    mann = TRUE,
    effectSize = TRUE)
```

### Plot

The following plot shows a boxplot comparing the alpha diversity between the Group levels (p-values are based in a `r ifelse(test_type == "parametric", "t.test", "wilcox.test")` compared to the '`r control`' level, with ns: p > 0.05; \*: p <= 0.05; \*\*: p <= 0.01; \*\*\*: p <= 0.001; \*\*\*\*: p <= 0.0001). 

```{r Section2_QC_filterdata_alpha_diversity_comparison_plot, fig.height = 8, fig.width = 10}
## TRUE is normal, FALSE is non-normal distribution 
metrics_normality <-
  c(
    "Observed" = FALSE,
    "Chao1" = FALSE,
    "Shannon" = TRUE,
    "Simpson" = FALSE
  )

## Add normality information and tidy data frame
adiv_long <- adiv %>%
  tidyr::gather(key = "metric",
                value = "value",
                c("Observed", "Chao1", "Shannon", "Simpson")) %>%
  dplyr::mutate(metric = factor(metric, levels = c("Observed", "Chao1", "Shannon", "Simpson")),
                Normal = metrics_normality[as.character(metric)])

## Function for Normal data: mean SD
mean_sd <- function(x) {
  mean_data <- mean(x)
  sd_data <- sd(x)
  ymin <- mean_data - sd_data
  ymax <- mean_data + sd_data
  return(data.frame(y = mean_data, ymin = ymin, ymax = ymax))
}

## Function for Non-Normal data: median IQR
median_iqr <- function(x) {
  med_data <- median(x)
  iqr_data <- IQR(x)
  lower <- med_data - iqr_data / 2
  upper <- med_data + iqr_data / 2
  return(data.frame(y = med_data, ymin = lower, ymax = upper))
}

## Plot
adiv_group_plot <- adiv_long %>%
  ggplot(aes(x = Group, y = value)) +
  geom_jitter(
    aes(color = Group), 
    size = 2, 
    alpha = 0.6,
    width = 0.25
    ) +
  facet_wrap( ~ metric, scales = "free", ncol = 4) +
  stat_compare_means(
    comparisons = list(c("EHI", "Control")),
    method = "wilcox.test",
    label = "p.signif",
    paired = FALSE,
    family = "serif",
    bracket.size = 0.6,
    vjust = -0.2
  ) +
  scale_fill_manual(values = group_cols) +
  scale_color_manual(values = group_cols) +
  scale_x_discrete(labels = c("Non-EHI", "EHI")) +
  labs(x = "",
       y = "Alpha Diversity") +
  theme_classic() +
  theme(
    text = font,
    axis.title = element_text(size = 24),
    axis.text.y = element_text(size = 16),
    axis.text.x = element_text(size = 16),
    strip.text = element_text(size = 26, face = "bold"),
    legend.position = "none"
  ) +
  expand_limits(y = 0) +
  stat_summary(
    data = filter(adiv_long, Normal),
    fun = mean,
    geom = "point",
    color = "black",
    size = 3,
    show.legend = FALSE
  ) +
  stat_summary(
    data = filter(adiv_long, Normal),
    fun.data = mean_sd,
    geom = "errorbar",
    color = "black",
    width = 0.3,
    linewidth = 0.8,
    show.legend = FALSE
  ) +
  stat_summary(
    data = filter(adiv_long,!Normal),
    fun = median,
    geom = "crossbar",
    color = "black",
    width = 0.6,
    linewidth = 0.4,
    show.legend = FALSE
  ) +
  stat_summary(
    data = filter(adiv_long,!Normal),
    fun.data = median_iqr,
    geom = "errorbar",
    color = "black",
    width = 0.3,
    linewidth = 0.6,
    show.legend = FALSE
  )

adiv_group_plot
``` 

## Beta diversity - Rarefied and filtered data

The following principal coordinate analysis (PCoA) plot shows the beta diversity following rarefaction and filtering of samples, based on the Bray Curtis distance between samples. 

``` {r Section2_QC_filterdata_beta_diversity, fig.height = 7, fig.width = 9.5}
## beta diversity - PCoA
bdiv_group_plot <- plot_ordination(phyloseq_dat_filter, 
                                   ordinate(phyloseq_dat_filter, method = "PCoA", distance = "bray", na.rm = TRUE), 
                                   type = "samples", 
                                   color = "Group"
                                   ) + 
  stat_ellipse(aes(group = Group), 
               linetype = 2,
               linewidth = 1) +
  geom_point(size = 5,
             alpha = 0.8) +
  ggtitle("") + 
  scale_fill_manual(values = group_cols) +
  scale_color_manual(values = group_cols,
                     labels = c("Non-EHI", "EHI")) +
  theme_classic() +
  theme(
    text = font,
    axis.title = element_text(size = 20),
    axis.text = element_text(size = 16),
    legend.text = element_text(size = 18),
    legend.title = element_text(size=22))

## view
bdiv_group_plot
```

## PERMANOVA

This section will use permutational MANOVA (PERMANOVA) to look for significant effects of the Group variable on the abundance. Dissimilarity is again based on the Bray Curtis difference. 

``` {r Section2_PERMANOVA_analysis}
## Set seed for reproducibility 
set.seed(0)

## Convert the counts across all samples into a community data matrix
plot_dat <- psmelt(phyloseq_dat_filter)
community_mat <- plot_dat %>%
  dplyr::select(Sample, OTU, Abundance) %>%
  spread(OTU, Abundance)
row.names(community_mat) <- community_mat[["Sample"]]
community_mat <- as.matrix(community_mat[, -1]) ## Remove the Sample column

## Transform the data using square root to minimize influence of most abundant groups
community_mat <- sqrt(community_mat)

## Create a dissimilarity matrix
OTU_dist <- vegdist(community_mat, method = "bray")

## Get the matching sample data
ind_dat <- plot_dat %>%
  dplyr::select(Sample, Group) %>%
  distinct()
ind_dat <- ind_dat[match(rownames(community_mat), ind_dat[["Sample"]]), ]

## Run perMANOVA
perm <- adonis2(OTU_dist ~ Group, data = ind_dat, permutations = 9999, method = "bray")
perm
```

# Section 3: Microbial Abundance

## Normalise

Here we normalise the taxa counts to represent the percentage of all taxas.

``` {r Section3_normalise}
phyloseq_dat_filter_norm <- transform_sample_counts(phyloseq_dat_filter, function (x) 100* ((x+1) / sum(x+1)))
phyloseq_dat_filter_norm_melt <- psmelt(phyloseq_dat_filter_norm)

## melted object of rarefied data
phyloseq_dat_filter_melt <- psmelt(phyloseq_dat_filter)

## melted phyloseq object on non-rarefied data for DESeq2
phyloseq_dat_norm_melt <- psmelt(phyloseq_dat_deseq)
```

## Summerise 

``` {r Section3_summarise_taxa}
## rarefied data: print the number and each unique taxas. Each number total will contain one NA. 
length(unique(phyloseq_dat_filter_melt$Phylum))
unique(phyloseq_dat_filter_melt$Phylum)
length(unique(phyloseq_dat_filter_melt$Class))
unique(phyloseq_dat_filter_melt$Class)
length(unique(phyloseq_dat_filter_melt$Order))
unique(phyloseq_dat_filter_melt$Order)
length(unique(phyloseq_dat_filter_melt$Family))
unique(phyloseq_dat_filter_melt$Family)
length(unique(phyloseq_dat_filter_melt$Genus))
unique(phyloseq_dat_filter_melt$Genus)
```

```{r Section3_summarise_taxa_new}
## Function to calculate distinct counts of taxa for each UIN within Group
summarise_taxa <- function(data) {
  data %>%
    ## Filter by a count >= 1
    dplyr::filter(Abundance >= 1) %>%
    ## Group by UIN and Group
    dplyr::group_by(UIN, Group) %>%
    # Summarise unique counts of taxa at each taxonomic level
    dplyr::summarise(
      Phylum = n_distinct(Phylum, na.rm = TRUE),
      Class = n_distinct(Class, na.rm = TRUE),
      Order = n_distinct(Order, na.rm = TRUE),
      Family = n_distinct(Family, na.rm = TRUE),
      Genus = n_distinct(Genus, na.rm = TRUE),
      Species = n_distinct(Species, na.rm = TRUE),
      ASV = n_distinct(OTU, na.rm = TRUE),
      ## drop groupings
      .groups = 'drop'
    )
}

## Return results for unique taxa
summary_taxa_result <- summarise_taxa(phyloseq_dat_filter_melt)

## View the results
head(summary_taxa_result)

## descriptives with S-W test
jmv::descriptives(
    formula = . ~ Group,
    data = summary_taxa_result,
    iqr = TRUE,
    sw = TRUE)

round_half_up <- function(x) {
  floor(x + 0.5)
}

## summarise average & variance of taxa for each taxa level, by group
summary_by_group <- summary_taxa_result %>%
  dplyr::group_by(Group) %>%
  dplyr::summarise(
    ## phyla: median (IQR)
    median_phyla = median(Phylum, na.rm = FALSE),
    iqr_phyla = IQR(Phylum, na.rm = FALSE),
    ## classes: median (IQR)
    median_classes = median(Class, na.rm = FALSE),
    iqr_classes = IQR(Class, na.rm = FALSE),
    ## orders: mean SD
    mean_orders = mean(Order, na.rm = FALSE),
    sd_orders = sd(Order, na.rm = FALSE),
    ## families: median (IQR)
    median_families = median(Family, na.rm = FALSE),
    iqr_families = IQR(Family, na.rm = FALSE),
    ## genera: median (IQR)
    median_genera = median(Genus, na.rm = FALSE),
    iqr_genera = IQR(Genus, na.rm = FALSE),
    ## ASV: median (IQR)
    median_ASV = median(ASV, na.rm = FALSE),
    iqr_ASV = IQR(ASV, na.rm = FALSE)
  ) %>%
  ## round numbers to 0 decimal places, rounding .5 up
  dplyr::mutate(across(where(is.numeric), round_half_up))

summary_by_group

## Write this into a file
write.table(
  summary_by_group,
  paste0(analysis_dir, "/tabs/taxa_summary_table_by_group.csv"),
  row.names = FALSE,
  col.names = TRUE,
  sep = ",",
  quote = FALSE
)
```

## Barplot Function

### Barplot function: Normal text

The following function will plot a barplot showing the normalised abundance across all taxonomic levels with normal text for taxa. 

``` {r Section3_plot_barplot_normal_text}
## Function for non-italicised taxa names in the legend
plot_barplot <-
  function (taxa, plot_dat = phyloseq_dat_filter_norm_melt) {
    # taxa = "Genus"
    # plot_dat = phyloseq_dat_filter_norm_melt
    
    ## Get the total abundance across all samples for each OTU
    tmp <- plot_dat %>%
      dplyr::group_by(OTU) %>%
      dplyr::summarise(sum = sum(Abundance))
    tsums <- tmp[["sum"]]
    names(tsums) <- tmp[["OTU"]]
    tsums <- tsums[order(names(tsums))]
    
    ## Get the OTU taxa
    tmp <- subset(plot_dat, Sample == plot_dat[1, "Sample"])
    taxlist <- tmp[, taxa]
    names(taxlist) <- tmp[, "OTU"]
    
    ## Check the data match
    if (length(tsums) != length(taxlist) ||
        !all(names(tsums) %in% names(taxlist)) ||
        !all(names(taxlist) %in% names(tsums))) {
      stop("Error: tsums and taxlist do not match.\n")
    }
    taxlist <- taxlist[names(tsums)]
    
    ## Keep only the top taxa and add all other taxa into an "other" class
    taxasum <-
      sort(tapply(tsums, taxlist, sum, na.rm = TRUE), decreasing = TRUE)
    taxasum <- taxasum[!names(taxasum) == ""]
    plot_dat[["TaxName"]] <- as.character(plot_dat[[taxa]])
    plot_dat[is.na(plot_dat[["TaxName"]]) |
               plot_dat[["TaxName"]] == "", "TaxName"] <- "Other"
    plot_dat[["TaxName"]] <-
      ifelse(plot_dat[["TaxName"]] %in% names(taxasum)[1:nclass],
             as.character(plot_dat[[taxa]]),
             "Other")
    plot_dat[["TaxName"]] <-
      factor(plot_dat[["TaxName"]], levels = c(names(taxasum)[1:nclass], "Other"))

    ## Combine counts at the taxa level
    plot_dat <- plot_dat %>%
      dplyr::group_by(Sample, Group, UIN, TaxName) %>%
      dplyr::summarise(Abundance = sum(Abundance)) %>%
      ## remove "[]", "-", "_" from taxa names
      dplyr::mutate(TaxName = gsub("\\[|\\]|-|_", " ", TaxName)) %>%
      ## remove leading and trialing spaces!
      dplyr::mutate(TaxName = trimws(TaxName)) %>%
      ## remove instances of doubles spaces
      dplyr::mutate(TaxName = gsub("  ", " ", TaxName))
    
    ## Extract unique and edited TaxName values
    unique_taxa <- unique(plot_dat$TaxName)
    
    ## reset factor levels again
    plot_dat <- plot_dat %>%
      dplyr::mutate(TaxName = factor(TaxName, levels = unique_taxa))
    
    ## Write this into a file
    write.table(
      plot_dat,
      paste0(
        analysis_dir,
        "/tabs/individual_abundance_table_",
        taxa,
        ".csv"
      ),
      row.names = FALSE,
      col.names = TRUE,
      sep = ",",
      quote = FALSE
    )
    
    ## Create a colour pallet for the barplot
    set.seed(8)
    cols <-
      sample(custom_col16, length(unique(plot_dat[["TaxName"]])))
    names(cols) <- unique(plot_dat[["TaxName"]])
    cols["Other"] <- "#D3D3D3" ## set other as grey
    
    custom_labels <- c(Control = "Non-EHI", EHI = "EHI")
    
    ## Create the plot with the custom legend formatting
    p <- plot_dat %>%
      ggplot(aes(x = UIN, y = Abundance, fill = TaxName)) +
      geom_bar(stat = "identity", width = 1) +
      labs(x = "",
           y = "Relative Abundance") +
      scale_y_continuous(labels = scales::label_percent(scale = 1),
                         expand = expansion(add = c(0, 1))) +
      scale_fill_manual(values = cols) +
      facet_wrap(. ~ Group,
                 scales = "free_x",
                 labeller = as_labeller(custom_labels)) +
      theme_classic() +
      theme(
        text = font,
        axis.title = element_text(size = 18),
        axis.text.y = element_text(size = 16),
        axis.text.x = element_blank(),
        legend.text = element_markdown(size = 14),
        legend.title = element_blank(),
        strip.text.x = element_text(size = 16, face = "bold"),
        strip.text.y = element_text(size = 16, face = "bold")
      )
    
    p
    
    ## Return the figure and the data set
    return(list(p, plot_dat))
}
```

### Barplot function: Italicised Text

The following function will plot a barplot showing the normalised abundance across all taxonomic levels with italicised text for taxa. 

```{r Section3_plot_barplot_italised_text}
## Function for italicised taxa names in the legend
plot_barplot2 <- function (taxa,
                           plot_dat = phyloseq_dat_filter_norm_melt,
                           specific_name = NULL) {
  # taxa = "Genus"
  # plot_dat = phyloseq_dat_NFW_filter_norm_melt
  # specific_name = "negative_control"
  
  ## Construct the csv filename based on whether a specific name is provided
  if (is.null(specific_name)) {
    filename <- paste0("individual_abundance_table_", taxa, ".csv")
  } else {
    filename <-
      paste0("individual_abundance_table_",
             specific_name,
             "_",
             taxa,
             ".csv")
  }
  
  ## Include the analysis directory in the file path
  filepath <-
    file.path(analysis_dir, "tabs", filename) 
  
  ## Get the total abundance across all samples for each OTU
  tmp <- plot_dat %>%
    dplyr::group_by(OTU) %>%
    dplyr::summarise(sum = sum(Abundance))
  tsums <- tmp[["sum"]]
  names(tsums) <- tmp[["OTU"]]
  tsums <- tsums[order(names(tsums))]
  
  ## Get the OTU taxa
  tmp <- subset(plot_dat, Sample == plot_dat[1, "Sample"])
  taxlist <- tmp[, taxa]
  names(taxlist) <- tmp[, "OTU"]
  
  ## Check the data match
  if (length(tsums) != length(taxlist) ||
      !all(names(tsums) %in% names(taxlist)) ||
      !all(names(taxlist) %in% names(tsums))) {
    stop("Error: tsums and taxlist do not match.\n")
  }
  taxlist <- taxlist[names(tsums)]
  
  ## Keep only the top taxa and add all other taxa into an "other" class
  taxasum <-
    sort(tapply(tsums, taxlist, sum, na.rm = TRUE), decreasing = TRUE)
  taxasum <- taxasum[!names(taxasum) == ""]
  plot_dat[["TaxName"]] <- as.character(plot_dat[[taxa]])
  plot_dat[is.na(plot_dat[["TaxName"]]) |
             plot_dat[["TaxName"]] == "", "TaxName"] <- "Other"
  plot_dat[["TaxName"]] <-
    ifelse(plot_dat[["TaxName"]] %in% names(taxasum)[1:nclass],
           as.character(plot_dat[[taxa]]),
           "Other")
  plot_dat[["TaxName"]] <-
    factor(plot_dat[["TaxName"]], levels = c(names(taxasum)[1:nclass], "Other"))
  
  ## Combine counts at the taxa level
  plot_dat <- plot_dat %>%
    dplyr::group_by(Sample, Group, UIN, TaxName) %>%
    dplyr::summarise(Abundance = sum(Abundance)) %>%
    dplyr::mutate(## remove "[]", "-", "_" from taxa names
      TaxName = gsub("\\[|\\]|-|_", " ", TaxName),
      ## remove leading and trialing spaces!
      TaxName = trimws(TaxName)) %>%
    ## remove instances of doubles spaces
    dplyr::mutate(TaxName = gsub("  ", " ", TaxName)) %>%
    ## Italicising taxa names
    dplyr::mutate(FormattedTaxName = ifelse(
      ## keep uncultured as normal text
      grepl("^uncultured", TaxName),
      gsub("^(uncultured) (.+)$", "\\1 *\\2*", TaxName),
      ## Keep the stated names (e.g. "Other") as normal text (i.e. not italicised) and wrap taxa names with "*"
      ifelse(
        TaxName %in% c(
          "Other",
          "metagenome",
          "gut metagenome",
          "human gut",
          "unidentified"
        ),
        TaxName,
        paste("*", TaxName, "*", sep = "")
      )
    )) %>%
    ## insert line break in long taxa names
    dplyr::mutate(
      FormattedTaxName = ifelse(
        FormattedTaxName == "*Eubacterium hallii group*",
        "*Eubacterium<br>hallii group*",
        FormattedTaxName
      )
    )

  ## Extract unique and edited TaxName values
  unique_taxa <- unique(plot_dat$TaxName)
  unique_taxa2 <- unique(plot_dat$FormattedTaxName)
  
  ## reset factor levels again
  plot_dat <- plot_dat %>%
    dplyr::mutate(
      TaxName = factor(TaxName, levels = unique_taxa),
      FormattedTaxName = factor(FormattedTaxName, levels = unique_taxa2)
    )
  
  ## Write this into a file
  write.table(
    plot_dat,
    file = filepath,
    row.names = FALSE,
    col.names = TRUE,
    sep = ",",
    quote = FALSE
  )
  
  ## Create a colour pallet for the barplot
  set.seed(8)
  cols <-
    sample(custom_col16, length(unique(plot_dat[["FormattedTaxName"]])))
  names(cols) <- unique(plot_dat[["FormattedTaxName"]])
  cols["Other"] <- "#D3D3D3" ## set other as grey
  
  custom_labels <- c(Control = "Non-EHI", EHI = "EHI")
  
  ## Create the plot with the custom legend formatting
  p <- plot_dat %>%
    ggplot(aes(x = UIN, y = Abundance, fill = FormattedTaxName)) +
    geom_bar(stat = "identity", width = 1) +
    labs(x = "",
         y = "Relative Abundance") +
    scale_y_continuous(labels = scales::label_percent(scale = 1),
                       expand = expansion(add = c(0, 1))) +
    scale_fill_manual(values = cols) +
    facet_wrap(. ~ Group,
               scales = "free_x",
               labeller = as_labeller(custom_labels)) +
    theme_classic() +
    theme(
      text = font,
      axis.title = element_text(size = 18),
      axis.text.y = element_text(size = 16),
      axis.text.x = element_blank(),
      legend.text = element_markdown(size = 14),
      ## Format legend text as markdown so names are italicised
      legend.title = element_blank(),
      strip.text.x = element_text(size = 16, face = "bold"),
      strip.text.y = element_text(size = 16, face = "bold")
    )
  
  p
  
  ## Return the figure and the data set
  return(list(p, plot_dat))
}
```

### Phylum
``` {r Section3_barplot_top_phylum_normalised_abundance_stacked, fig.height = 8, fig.width = 8}
phylum_dat <- plot_barplot("Phylum")
phylum_p <- phylum_dat[[1]]
phylum_p
```

### Class
``` {r Section3_barplot_top_class_normalised_abundance_stacked, fig.height = 8, fig.width = 8}
class_dat <- plot_barplot("Class")
class_dat[[1]]
```

### Order
``` {r Section3_barplot_top_order_normalised_abundance_stacked, fig.height = 8, fig.width = 8}
order_dat <- plot_barplot("Order")
order_dat[[1]]
```

### Family
``` {r Section3_barplot_top_family_normalised_abundance_stacked, fig.height = 8, fig.width = 8}
family_dat <- plot_barplot2("Family")
family_dat[[1]]
```

### Genus
``` {r Section3_barplot_top_genus_normalised_abundance_stacked, fig.height = 8, fig.width = 8}
genus_dat <- plot_barplot2("Genus")
genus_p <- genus_dat[[1]]
genus_p
```

## Firmicute/Bacteroidota ratio

This section will calculate the Firmicutes/Bacteroidota (FB) ratio, which is often used as a measure of gut microbiome dysregulation. 

```{r Section3_calculate_FBratio, echo = TRUE}
fbratio <- phylum_dat[[2]] %>%
  dplyr::group_by(UIN) %>%
  tidyr::spread(TaxName, Abundance) %>%
  dplyr::mutate(FBratio = Firmicutes/Bacteroidota) %>%
  dplyr::select(UIN, Group, Firmicutes, Bacteroidota, FBratio)

head(fbratio)

## Write as csv
write.table(
  fbratio,
  paste0(analysis_dir, "/tabs/FBratio_table.csv"),
  row.names = FALSE,
  col.names = TRUE,
  sep = ",",
  quote = FALSE
)
```

### Normality

```{r FBratio_normality}
jmv::descriptives(
    formula = FBratio ~ Group,
    data = fbratio,
    iqr = TRUE,
    sw = TRUE)
```

### t-test

```{r FBratio_comparison_stat}
jmv::ttestIS(
    formula = FBratio ~ Group,
    data = fbratio,
    vars = FBratio,
    students = FALSE,
    mann = TRUE,
    effectSize = TRUE)
```

### Plot

Below is a boxplot showing how the FB Ratio differs between groups.

``` {r Section3_FBratio_plot, echo = TRUE, fig.height = 6, fig.width = 6}
## plot figure
fbratio_plot <- fbratio %>% 
  ggplot(aes(x = Group, y = FBratio)) +
  geom_boxplot(outlier.color = NA,
               size = 1) +
  geom_jitter(aes(color = Group),
              size = 2,
              alpha = 0.6) +
  labs(x = "", 
       y = "F:B Ratio") +
  stat_compare_means(comparisons = list(c("EHI", "Control")),
                     method = "wilcox.test",
                     label = "p.signif",
                     paired = FALSE,
                     family = "serif",
                     bracket.size = 0.6,
                     vjust = -0.2) +
  scale_x_discrete(labels = c("Non-EHI", "EHI")) +
  scale_y_continuous(breaks = seq(0, 16, 2), 
                     limits = c(0, 16)) +
  scale_color_manual(values = group_cols) +
  theme_classic() +
  theme(
    text = font,
    axis.title   = element_text(size = 24),
    axis.text.y  = element_text(size = 16),
    axis.text.x  = element_text(size = 22),
    legend.position = "none"
    ) 

## view
fbratio_plot
```

## Abundance Comparison All Taxa Function

Below, we will compare all taxa, at each taxonomic level between groups. 

``` {r Section3_abundance_comparison_function}
compare_abundance_comparison <- function (taxa,
                                          plot_dat = phyloseq_dat_filter_norm_melt,
                                          p_type = "individual") {
  # taxa = "Phylum"
  # plot_dat = phyloseq_dat_filter_norm_melt
  # p_type = "individual"
  
  ## Get the total abundance across all samples for each OTU for the given taxa
  tmp <- plot_dat %>%
    dplyr::group_by(OTU) %>%
    dplyr::summarise(sum = sum(Abundance))
  tsums <- tmp[["sum"]]
  names(tsums) <- tmp[["OTU"]]
  tsums <- tsums[order(names(tsums))]
  
  ## Get the OTU taxa
  tmp <- subset(plot_dat, Sample == plot_dat[1, "Sample"])
  taxlist <- tmp[, taxa]
  names(taxlist) <- tmp[, "OTU"]
  
  ## Check the data match
  if (length(tsums) != length(taxlist) ||
      !all(names(tsums) %in% names(taxlist)) ||
      !all(names(taxlist) %in% names(tsums))) {
    stop("Error: tsums and taxlist do not match.\n")
  }
  taxlist <- taxlist[names(tsums)]
  
  taxasum <-
    sort(tapply(tsums, taxlist, sum, na.rm = TRUE), decreasing = TRUE)
  taxasum <- taxasum[!names(taxasum) == ""]
  plot_dat[["TaxName"]] <- as.character(plot_dat[[taxa]])
  plot_dat[["TaxName"]] <- factor(plot_dat[[taxa]])
  
  ## Combine counts at the taxa level
  plot_dat <- plot_dat %>%
    dplyr::group_by(Pair, UIN, Group, Troop, TaxName) %>%
    dplyr::summarise(Abundance = sum(Abundance), .groups = "drop") %>%
    dplyr::filter(TaxName != "Other")
  
  ## create separate wide data frame for all taxa of each individual UIN
  UIN_taxa_wide <- plot_dat
  UIN_taxa_wide <- UIN_taxa_wide %>%
    tidyr::pivot_wider(names_from = TaxName,
                       values_from = Abundance) %>%
  ## arrange by Pair and Group  
  dplyr::arrange(Pair, Group)
    
  ## save as .csv file
  write.table(
    UIN_taxa_wide,
    paste0(
      analysis_dir,
      "/tabs/individual_relative_abundance_",
      taxa,
      ".csv"
    ),
    sep = ",",
    row.names = FALSE,
    col.names = TRUE,
    quote = FALSE
  )
  
  ## Combine at the Group level
  plot_dat_SE <- plot_dat %>%
    dplyr::group_by(Group, TaxName) %>%
    dplyr::summarise(
      average = ifelse(
        test_type == "parametric",
        mean(Abundance),
        median(Abundance)
      ),
      variability = ifelse(
        test_type == "parametric",
        sd(Abundance) / sqrt(length(Abundance)),
        IQR(Abundance)
      )
    )
  
  ## Calculate p-value across all groups
  pval <-
    data.frame(TaxName       = levels(plot_dat_SE[["TaxName"]]),
               pval          = NA)
  for (t in pval[["TaxName"]]) {
    if (test_type == "parametric") {
      pval[pval[["TaxName"]] == t, "pval"] <-
        anova(lm(Abundance ~ Group, data = subset(plot_dat, TaxName == t)))[1, "Pr(>F)"]
    } else if (test_type == "non-parametric") {
      pval[pval[["TaxName"]] == t, "pval"] <-
        kruskal.test(Abundance ~ Group, data = subset(plot_dat, TaxName == t))[["p.value"]]
    }
  }
  
  ## Calculate individual p-values vs control
  for (t in pval[["TaxName"]]) {
    for (g in unique(plot_dat[["Group"]])) {
      if (g == "Control") {
        pval[pval[["TaxName"]] == t, paste0("pval.", g)] <- 1
      } else if (test_type == "parametric") {
        pval[pval[["TaxName"]] == t, paste0("pval.", g)] <-
          t.test(
            subset(plot_dat, TaxName == t &
                     Group == g)[["Abundance"]],
            subset(plot_dat, TaxName == t &
                     Group == control)[["Abundance"]]
          )[["p.value"]]
      } else if (test_type == "non-parametric") {
        pval[pval[["TaxName"]] == t, paste0("pval.", g)] <-
          wilcox.test(
            subset(plot_dat, TaxName == t &
                     Group == g)[["Abundance"]],
            subset(plot_dat, TaxName == t &
                     Group == control)[["Abundance"]]
          )[["p.value"]]
      }
    }
  }
  
  ## Add to data frame
  plot_dat_SE[["pval"]] <- NA
  for (t in pval[["TaxName"]]) {
    for (g in unique(plot_dat[["Group"]])) {
      if (p_type == "combine") {
        if (g == control) {
          plot_dat_SE[plot_dat_SE[["TaxName"]] == t &
                        plot_dat_SE[["Group"]] == g, "pval"] <-
            pval[pval[["TaxName"]] == t, "pval"]
        } else {
          plot_dat_SE[plot_dat_SE[["TaxName"]] == t &
                        plot_dat_SE[["Group"]] == g, "pval"] <- NA
        }
      } else if (p_type == "individual") {
        if (g == control) {
          plot_dat_SE[plot_dat_SE[["TaxName"]] == t &
                        plot_dat_SE[["Group"]] == g, "pval"] <- NA
        } else {
          plot_dat_SE[plot_dat_SE[["TaxName"]] == t &
                        plot_dat_SE[["Group"]] == g, "pval"] <-
            pval[pval[["TaxName"]] == t, paste0("pval.", g)]
        }
      }
    }
  }
  
  ## Correct for multiple testing
  which_nonna <- which(!is.na(plot_dat_SE[["pval"]]))
  plot_dat_SE[["padj"]] <- NA
  plot_dat_SE[which_nonna, "padj"] <-
    p.adjust(as.numeric(plot_dat_SE[which_nonna,][["pval"]]),
             method = "fdr",
             n = nrow(plot_dat_SE[which_nonna,]))
  
  ## Convert to star format
  plot_dat_SE[["padj_star"]] <-
    case_when(
      plot_dat_SE[["padj"]] <= 0.001 ~ "***",
      plot_dat_SE[["padj"]] <= 0.01  ~ "**",
      plot_dat_SE[["padj"]] <= 0.05  ~ "*",
      TRUE ~ ""
    )
  
  ## Save the values
  write.table(
    plot_dat_SE,
    paste0(
      analysis_dir,
      "/tabs/abundance_comparison_table_",
      taxa,
      ".csv"
    ),
    sep = ",",
    row.names = FALSE,
    col.names = TRUE,
    quote = FALSE
  )
  
  ## format table
  plot_dat_format <- plot_dat_SE %>%
    tidyr::pivot_wider(
      names_from = Group,
      values_from = c(average, variability, pval, padj, padj_star)
    ) %>%
    dplyr::mutate(
      average_Control = sprintf("%.1f", average_Control),
      variability_Control = sprintf("%.1f", variability_Control),
      average_EHI = sprintf("%.1f", average_EHI),
      variability_EHI = sprintf("%.1f", variability_EHI),
      Control = paste0(average_Control, " (", variability_Control, ")"),
      EHI = paste0(average_EHI, " (", variability_EHI, ")")
    ) %>%
    dplyr::rename(pval = pval_EHI,
                  padj = padj_EHI,
                  Taxa = TaxName) %>%
    dplyr::mutate(pval = sprintf("%.3f", as.numeric(pval)),
                  padj = sprintf("%.3f", as.numeric(padj))) %>%
    dplyr::select(Taxa, Control, EHI, pval, padj)
  
  ## save formatted table
  write.table(
    plot_dat_format,
    paste0(
      analysis_dir,
      "/tabs/abundance_comparison_table_formatted_",
      taxa,
      ".csv"
    ),
    sep = ",",
    row.names = FALSE,
    col.names = TRUE,
    quote = FALSE
  )
  
  return(plot_dat_format)
}
```

### Phylum

```{r Section3_abundance_comparison_Phylum}
phylum_comparison <- compare_abundance_comparison("Phylum")
phylum_comparison

## filter by pval
phylum_comparison %>%
  dplyr::filter(pval < 0.05)

## filter by padj
phylum_comparison %>%
  dplyr::filter(padj < 0.05)
```

### Class

```{r Section3_abundance_comparison_Class}
class_comparison <- compare_abundance_comparison("Class")
class_comparison

## filter by pval
class_comparison %>%
  dplyr::filter(pval < 0.05)

## filter by padj
class_comparison %>%
  dplyr::filter(padj < 0.05)
```

### Order

```{r Section3_abundance_comparison_Order}
order_comparison <- compare_abundance_comparison("Order")
order_comparison

## filter by pval
order_comparison %>%
  dplyr::filter(pval < 0.05)

## filter by padj
order_comparison %>%
  dplyr::filter(padj < 0.05)
```

### Family

```{r Section3_abundance_comparison_Family}
family_comparison <- compare_abundance_comparison("Family")
family_comparison

## filter by pval
family_comparison %>%
  dplyr::filter(pval < 0.05)

## filter by padj
family_comparison %>%
  dplyr::filter(padj < 0.05)
```

### Genus

```{r Section3_abundance_comparison_Genus}
genus_comparison <- compare_abundance_comparison("Genus")
genus_comparison

## filter by pval
genus_comparison %>%
  dplyr::filter(pval < 0.05)

## filter by padj
genus_comparison %>%
  dplyr::filter(padj < 0.05)
```


## Plot Abundance Comparison Function:

### Abundance Comparison Function: Normal Text

The following function will plot the difference between the groups at different taxonomic levels with normal text for taxa:

``` {r Section3_plot_abundance_comparison_normal_text}
## Function for non-italicised taxa names
plot_abundance_comparison <- function (taxa,
                                       plot_dat = phyloseq_dat_filter_norm_melt,
                                       p_type = "individual",
                                       nclass = 15) {
  # taxa = "Phylum"
  # plot_dat = phyloseq_dat_filter_norm_melt
  # p_type = "individual"
  
  ## Get the total abundance across all samples for each OTU for the given taxa
  tmp <- plot_dat %>%
    dplyr::group_by(OTU) %>%
    dplyr::summarise(sum = sum(Abundance))
  tsums <- tmp[["sum"]]
  names(tsums) <- tmp[["OTU"]]
  tsums <- tsums[order(names(tsums))]
  
  ## Get the OTU taxa
  tmp <- subset(plot_dat, Sample == plot_dat[1, "Sample"])
  taxlist <- tmp[, taxa]
  names(taxlist) <- tmp[, "OTU"]
  
  ## Check the data match
  if (length(tsums) != length(taxlist) ||
      !all(names(tsums) %in% names(taxlist)) ||
      !all(names(taxlist) %in% names(tsums))) {
    stop("Error: tsums and taxlist do not match.\n")
  }
  taxlist <- taxlist[names(tsums)]
  
  ## Keep only the top taxa and add all other taxa into an "other" class
  taxasum <-
    sort(tapply(tsums, taxlist, sum, na.rm = TRUE), decreasing = TRUE)
  taxasum <- taxasum[!names(taxasum) == ""]
  plot_dat[["TaxName"]] <- as.character(plot_dat[[taxa]])
  plot_dat[is.na(plot_dat[["TaxName"]]) |
             plot_dat[["TaxName"]] == "", "TaxName"] <- "Other"
  plot_dat[["TaxName"]] <-
    ifelse(plot_dat[["TaxName"]] %in% names(taxasum)[1:nclass],
           as.character(plot_dat[[taxa]]),
           "Other")
  plot_dat[["TaxName"]] <-
    factor(plot_dat[["TaxName"]], levels = c(names(taxasum)[1:nclass], "Other"))
  
  ## Combine counts at the taxa level
  plot_dat <- plot_dat %>%
    dplyr::group_by(UIN, Group, TaxName) %>%
    dplyr::summarise(Abundance = sum(Abundance), .groups = "drop") %>%
    dplyr::filter(TaxName != "Other")

  ## Combine at the Group level
  plot_dat_SE <- plot_dat %>%
    dplyr::group_by(Group, TaxName) %>%
    dplyr::summarise(
      average = ifelse(
        test_type == "parametric",
        mean(Abundance),
        median(Abundance)
      ),
      variability = ifelse(
        test_type == "parametric",
        sd(Abundance) / sqrt(length(Abundance)),
        IQR(Abundance)
      )
    )

  ## Calculate p-value across all groups
  pval <-
    data.frame(TaxName       = head(levels(plot_dat_SE[["TaxName"]]),-1),
               pval          = NA)
  for (t in pval[["TaxName"]]) {
    if (test_type == "parametric") {
      pval[pval[["TaxName"]] == t, "pval"] <-
        anova(lm(Abundance ~ Group, data = subset(plot_dat, TaxName == t)))[1, "Pr(>F)"]
    } else if (test_type == "non-parametric") {
      pval[pval[["TaxName"]] == t, "pval"] <-
        kruskal.test(Abundance ~ Group, data = subset(plot_dat, TaxName == t))[["p.value"]]
    }
  }
  
  ## Calculate individual p-values vs control
  for (t in pval[["TaxName"]]) {
    for (g in unique(plot_dat[["Group"]])) {
      if (g == "Control") {
        pval[pval[["TaxName"]] == t, paste0("pval.", g)] <- 1
      } else if (test_type == "parametric") {
        pval[pval[["TaxName"]] == t, paste0("pval.", g)] <-
          t.test(
            subset(plot_dat, TaxName == t &
                     Group == g)[["Abundance"]],
            subset(plot_dat, TaxName == t &
                     Group == control)[["Abundance"]]
          )[["p.value"]]
      } else if (test_type == "non-parametric") {
        pval[pval[["TaxName"]] == t, paste0("pval.", g)] <-
          wilcox.test(
            subset(plot_dat, TaxName == t &
                     Group == g)[["Abundance"]],
            subset(plot_dat, TaxName == t &
                     Group == control)[["Abundance"]]
          )[["p.value"]]
      }
    }
  }
  
  ## Add to data frame
  plot_dat_SE[["pval"]] <- NA
  for (t in pval[["TaxName"]]) {
    for (g in unique(plot_dat[["Group"]])) {
      if (p_type == "combine") {
        if (g == control) {
          plot_dat_SE[plot_dat_SE[["TaxName"]] == t &
                        plot_dat_SE[["Group"]] == g, "pval"] <-
            pval[pval[["TaxName"]] == t, "pval"]
        } else {
          plot_dat_SE[plot_dat_SE[["TaxName"]] == t &
                        plot_dat_SE[["Group"]] == g, "pval"] <- NA
        }
      } else if (p_type == "individual") {
        if (g == control) {
          plot_dat_SE[plot_dat_SE[["TaxName"]] == t &
                        plot_dat_SE[["Group"]] == g, "pval"] <- NA
        } else {
          plot_dat_SE[plot_dat_SE[["TaxName"]] == t &
                        plot_dat_SE[["Group"]] == g, "pval"] <-
            pval[pval[["TaxName"]] == t, paste0("pval.", g)]
        }
      }
    }
  }
  
  ## Correct for multiple testing
  which_nonna <- which(!is.na(plot_dat_SE[["pval"]]))
  plot_dat_SE[["pval_adj"]] <- NA
  plot_dat_SE[which_nonna, "pval_adj"] <-
    p.adjust(as.numeric(plot_dat_SE[which_nonna,][["pval"]]),
             method = "fdr",
             n = nrow(plot_dat_SE[which_nonna,]))
  
  ## Convert to star format
  plot_dat_SE[["pval_adj_star"]] <-
    case_when(
      plot_dat_SE[["pval_adj"]] <= 0.001 ~ "***",
      plot_dat_SE[["pval_adj"]] <= 0.01  ~ "**",
      plot_dat_SE[["pval_adj"]] <= 0.05  ~ "*",
      TRUE ~ ""
    )
  
  ## Save the values
  write.table(
    plot_dat_SE,
    paste0(
      analysis_dir,
      "/tabs/plot_abundance_comparison_",
      taxa,
      ".csv"
    ),
    sep = ",",
    row.names = FALSE,
    col.names = TRUE,
    quote = FALSE
  )
  
  ## Plot the data

  ## Determine x-axis limits
  x_max <-
    max(plot_dat$Abundance, na.rm = TRUE) * 1.02 ## Extend by 2%
  
  ## box plots
  p2 <- plot_dat %>%
    ## remove "[]", "-", "_" from taxa names
    dplyr::mutate(TaxName = gsub("\\[|\\]|-|_", " ", TaxName)) %>%
    ## remove leading and trialing spaces!
    dplyr::mutate(TaxName = trimws(TaxName)) %>%
    ## remove instances of double spaces
    dplyr::mutate(TaxName = gsub("  ", " ", TaxName)) %>%
    ggplot(aes(
      x = Abundance,
      y = reorder(TaxName, Abundance),
      ## flip group order
      fill = factor(Group, levels = c("EHI", "Control"))
    )) +
    geom_boxplot(
      width = 0.6,
      position = position_dodge(0.8),
      outlier.size = 1,
      outlier.alpha = 0.8
    ) +
    labs(x = "Relative Abundance",
         y = "",
         ## legend label
         fill = "Group") +
    scale_y_discrete(expand = c(0, 0.5)) +
    scale_x_continuous(
      labels = scales::label_percent(scale = 1),
      limits = c(0, x_max),
      expand = c(0, 0)
    ) +
    ## group colour for boxplot (flip order)
    scale_fill_manual(values = group_cols,
                      labels = c("EHI", "Non-EHI")) +
    ## group colour for jitter (flip order)
    scale_color_manual(values = group_cols,
                       labels = c("EHI", "Non-EHI")) +
    ## flip legend order so control above EHI
    guides(fill = guide_legend(reverse = TRUE)) +
    theme_classic() +
    theme(
      text = font,
      axis.title.x = element_text(size = 18),
      axis.text.x = element_text(size = 16),
      axis.text.y = element_markdown(size = 16, colour = "black"),
      legend.title = element_text(size = 22),
      legend.text  = element_text(size = 18),
      panel.grid.major.x = element_line(colour = "darkgrey", linewidth = 0.25)
    )
  
  ## boxplot with individual data points
  p3 <- plot_dat %>%
    ## remove "[]", "-", "_" from taxa names
    dplyr::mutate(TaxName = gsub("\\[|\\]|-|_", " ", TaxName)) %>%
    ## remove leading and trialing spaces!
    dplyr::mutate(TaxName = trimws(TaxName)) %>%
    ## remove instances of double spaces
    dplyr::mutate(TaxName = gsub("  ", " ", TaxName)) %>%
    ggplot(aes(
      x = Abundance,
      y = reorder(TaxName, Abundance),
      ## flip group order
      fill = factor(Group, levels = c("EHI", "Control"))
    )) +
    geom_jitter(
      ## colour jitter by group but flip factors to match fill global aesthetic
      aes(colour = factor(Group, levels = c("EHI", "Control"))),
      ## set jitter width and keep dodge consistent
      position = position_jitterdodge(jitter.width = 0.15, dodge.width = 0.8),
      size = 1.2,
      alpha = 0.6,
      show.legend = FALSE,
    ) +
    geom_boxplot(width = 0.6,
                 position = position_dodge(0.8),
                 ## remove outlier dots as we have jitter points
                 outlier.shape = NA) +
    labs(x = "Relative Abundance",
         y = "",
         ## legend label
         fill = "Group") +
    scale_y_discrete(expand = c(0, 0.5)) +
    scale_x_continuous(
      labels = scales::label_percent(scale = 1),
      limits = c(0, x_max),
      expand = c(0, 0)
    ) +
    ## group colour for boxplot (flip order)
    scale_fill_manual(values = group_cols,
                      labels = c("EHI", "Non-EHI")) +
    ## group colour for jitter (flip order)
    scale_color_manual(values = group_cols,
                       labels = c("EHI", "Non-EHI")) +
    ## flip legend order so control above EHI
    guides(fill = guide_legend(reverse = TRUE)) +
    theme_classic() +
    theme(
      text = font,
      axis.title.x = element_text(size = 18),
      axis.text.x = element_text(size = 16),
      axis.text.y = element_markdown(size = 16, colour = "black"),
      legend.title = element_text(size = 22),
      legend.text  = element_text(size = 18),
      panel.grid.major.x = element_line(colour = "darkgrey", linewidth = 0.25)
    )
  
  list(
    BoxPlot = p2,
    JitterBoxPlot = p3
    )

} 
```

### Abundance Comparison Function: Italicised Text

The following function will plot the difference between the groups at different taxonomic levels with italicised text for taxa:

```{r Section3_plot_abundance_comparison_italised_text}
## Function for italicised taxa names
plot_abundance_comparison2 <- function (taxa,
                                        plot_dat = phyloseq_dat_filter_norm_melt,
                                        p_type = "individual",
                                        nclass = 15) {
  # taxa = "Genus"
  # plot_dat = phyloseq_dat_filter_norm_melt
  # p_type = "individual"
  # nclass = 15
  
  ## Get the total abundance across all samples for each OTU for the given taxa
  tmp <- plot_dat %>%
    dplyr::group_by(OTU) %>%
    dplyr::summarise(sum = sum(Abundance))
  tsums <- tmp[["sum"]]
  names(tsums) <- tmp[["OTU"]]
  tsums <- tsums[order(names(tsums))]
  
  ## Get the OTU taxa
  tmp <- subset(plot_dat, Sample == plot_dat[1, "Sample"])
  taxlist <- tmp[, taxa]
  names(taxlist) <- tmp[, "OTU"]
  
  ## Check the data match
  if (length(tsums) != length(taxlist) ||
      !all(names(tsums) %in% names(taxlist)) ||
      !all(names(taxlist) %in% names(tsums))) {
    stop("Error: tsums and taxlist do not match.\n")
  }
  taxlist <- taxlist[names(tsums)]
  
  ## Keep only the top taxa and add all other taxa into an "other" class
  taxasum <-
    sort(tapply(tsums, taxlist, sum, na.rm = TRUE), decreasing = TRUE)
  taxasum <- taxasum[!names(taxasum) == ""]
  plot_dat[["TaxName"]] <- as.character(plot_dat[[taxa]])
  plot_dat[is.na(plot_dat[["TaxName"]]) |
             plot_dat[["TaxName"]] == "", "TaxName"] <- "Other"
  plot_dat[["TaxName"]] <-
    ifelse(plot_dat[["TaxName"]] %in% names(taxasum)[1:nclass],
           as.character(plot_dat[[taxa]]),
           "Other")
  plot_dat[["TaxName"]] <-
    factor(plot_dat[["TaxName"]], levels = c(names(taxasum)[1:nclass], "Other"))
  
  ## Combine counts at the taxa level
  plot_dat <- plot_dat %>%
    dplyr::group_by(UIN, Group, TaxName) %>%
    dplyr::summarise(Abundance = sum(Abundance), .groups = "drop") %>%
    dplyr::filter(TaxName != "Other")

  ## Combine at the Group level
  plot_dat_SE <- plot_dat %>%
    dplyr::group_by(Group, TaxName) %>%
    dplyr::summarise(
      average = ifelse(
        test_type == "parametric",
        mean(Abundance),
        median(Abundance)
      ),
      variability = ifelse(
        test_type == "parametric",
        sd(Abundance) / sqrt(length(Abundance)),
        IQR(Abundance)
      )
    )

  ## Calculate p-value across all groups
  pval <-
    data.frame(TaxName       = head(levels(plot_dat_SE[["TaxName"]]),-1),
               pval          = NA)
  for (t in pval[["TaxName"]]) {
    if (test_type == "parametric") {
      pval[pval[["TaxName"]] == t, "pval"] <-
        anova(lm(Abundance ~ Group, data = subset(plot_dat, TaxName == t)))[1, "Pr(>F)"]
    } else if (test_type == "non-parametric") {
      pval[pval[["TaxName"]] == t, "pval"] <-
        kruskal.test(Abundance ~ Group, data = subset(plot_dat, TaxName == t))[["p.value"]]
    }
  }
  
  ## Calculate individual p-values vs control
  for (t in pval[["TaxName"]]) {
    for (g in unique(plot_dat[["Group"]])) {
      if (g == "Control") {
        pval[pval[["TaxName"]] == t, paste0("pval.", g)] <- 1
      } else if (test_type == "parametric") {
        pval[pval[["TaxName"]] == t, paste0("pval.", g)] <-
          t.test(
            subset(plot_dat, TaxName == t &
                     Group == g)[["Abundance"]],
            subset(plot_dat, TaxName == t &
                     Group == control)[["Abundance"]]
          )[["p.value"]]
      } else if (test_type == "non-parametric") {
        pval[pval[["TaxName"]] == t, paste0("pval.", g)] <-
          wilcox.test(
            subset(plot_dat, TaxName == t &
                     Group == g)[["Abundance"]],
            subset(plot_dat, TaxName == t &
                     Group == control)[["Abundance"]]
          )[["p.value"]]
      }
    }
  }
  
  ## Add to data frame
  plot_dat_SE[["pval"]] <- NA
  for (t in pval[["TaxName"]]) {
    for (g in unique(plot_dat[["Group"]])) {
      if (p_type == "combine") {
        if (g == control) {
          plot_dat_SE[plot_dat_SE[["TaxName"]] == t &
                        plot_dat_SE[["Group"]] == g, "pval"] <-
            pval[pval[["TaxName"]] == t, "pval"]
        } else {
          plot_dat_SE[plot_dat_SE[["TaxName"]] == t &
                        plot_dat_SE[["Group"]] == g, "pval"] <- NA
        }
      } else if (p_type == "individual") {
        if (g == control) {
          plot_dat_SE[plot_dat_SE[["TaxName"]] == t &
                        plot_dat_SE[["Group"]] == g, "pval"] <- NA
        } else {
          plot_dat_SE[plot_dat_SE[["TaxName"]] == t &
                        plot_dat_SE[["Group"]] == g, "pval"] <-
            pval[pval[["TaxName"]] == t, paste0("pval.", g)]
        }
      }
    }
  }
  
  ## Correct for multiple testing
  which_nonna <- which(!is.na(plot_dat_SE[["pval"]]))
  plot_dat_SE[["pval_adj"]] <- NA
  plot_dat_SE[which_nonna, "pval_adj"] <-
    p.adjust(as.numeric(plot_dat_SE[which_nonna,][["pval"]]),
             method = "fdr",
             n = nrow(plot_dat_SE[which_nonna,]))
  
  ## Convert to star format
  plot_dat_SE[["pval_adj_star"]] <-
    case_when(
      plot_dat_SE[["pval_adj"]] <= 0.001 ~ "***",
      plot_dat_SE[["pval_adj"]] <= 0.01  ~ "**",
      plot_dat_SE[["pval_adj"]] <= 0.05  ~ "*",
      TRUE ~ ""
    )

  ## Save the values
  write.table(
    plot_dat_SE,
    paste0(
      analysis_dir,
      "/tabs/plot_abundance_comparison_",
      taxa,
      ".csv"
    ),
    sep = ",",
    row.names = FALSE,
    col.names = TRUE,
    quote = FALSE
  )
  
  ## Plot the data

  ## Determine x-axis limits
  x_max <-
    max(plot_dat$Abundance, na.rm = TRUE) * 1.02 ## Extend by 2%
  
  ## box plots
  p2 <- plot_dat %>%
    ## remove "[]", "-", "_" from taxa names
    dplyr::mutate(TaxName = gsub("\\[|\\]|-|_", " ", TaxName)) %>%
    ## remove leading and trialing spaces!
    dplyr::mutate(TaxName = trimws(TaxName)) %>%
    ## remove instances of double spaces
    dplyr::mutate(TaxName = gsub("  ", " ", TaxName)) %>%
    dplyr::mutate(
      FormattedTaxName = ifelse(TaxName != "Other", paste("*", TaxName, "*", sep = ""), TaxName),
      FormattedTaxName = ifelse(
        FormattedTaxName == "*Eubacterium hallii group*",
        "*Eubacterium<br>hallii group*",
        FormattedTaxName
      )
    ) %>%
    ggplot(aes(
      x = Abundance,
      y = reorder(FormattedTaxName, Abundance),
      ## flip group order
      fill = factor(Group, levels = c("EHI", "Control"))
    )) +
    geom_boxplot(
      width = 0.6,
      position = position_dodge(0.8),
      outlier.size = 1,
      outlier.alpha = 0.8
    ) +
    labs(x = "Relative Abundance",
         y = "",
         ## legend label
         fill = "Group") +
    scale_y_discrete(expand = c(0, 0.5)) +
    scale_x_continuous(
      labels = scales::label_percent(scale = 1),
      limits = c(0, x_max),
      expand = c(0, 0)
    ) +
    ## group colour for boxplot (flip order)
    scale_fill_manual(values = group_cols,
                      labels = c("EHI", "Non-EHI")) +
    ## group colour for jitter (flip order)
    scale_color_manual(values = group_cols,
                       labels = c("EHI", "Non-EHI")) +
    ## flip legend order so control above EHI
    guides(fill = guide_legend(reverse = TRUE)) +
    theme_classic() +
    theme(
      text = font,
      axis.title.x = element_text(size = 18),
      axis.text.x = element_text(size = 16),
      axis.text.y = element_markdown(size = 16, colour = "black"),
      legend.title = element_text(size = 22),
      legend.text  = element_text(size = 18),
      panel.grid.major.x = element_line(colour = "darkgrey", linewidth = 0.25)
    )
  
  ## boxplot with individual data points
  p3 <- plot_dat %>%
    ## remove "[]", "-", "_" from taxa names
    dplyr::mutate(TaxName = gsub("\\[|\\]|-|_", " ", TaxName)) %>%
    ## remove leading and trialing spaces!
    dplyr::mutate(TaxName = trimws(TaxName)) %>%
    ## remove instances of double spaces
    dplyr::mutate(TaxName = gsub("  ", " ", TaxName)) %>%
    dplyr::mutate(
      FormattedTaxName = ifelse(TaxName != "Other", paste("*", TaxName, "*", sep = ""), TaxName),
      FormattedTaxName = ifelse(
        FormattedTaxName == "*Eubacterium hallii group*",
        "*Eubacterium<br>hallii group*",
        FormattedTaxName
      )
    ) %>%
    ggplot(aes(
      x = Abundance,
      y = reorder(FormattedTaxName, Abundance),
      ## flip group order
      fill = factor(Group, levels = c("EHI", "Control"))
    )) +
    geom_jitter(
      ## colour jitter by group but flip factors to match fill global aesthetic
      aes(colour = factor(Group, levels = c("EHI", "Control"))),
      ## set jitter width and keep dodge consistent
      position = position_jitterdodge(jitter.width = 0.15, dodge.width = 0.8),
      size = 1.2,
      alpha = 0.6,
      show.legend = FALSE,
    ) +
    geom_boxplot(width = 0.6,
                 position = position_dodge(0.8),
                 ## remove outlier dots as we have jitter points
                 outlier.shape = NA) +
    labs(x = "Relative Abundance",
         y = "",
         ## legend label
         fill = "Group") +
    scale_y_discrete(expand = c(0, 0.5)) +
    scale_x_continuous(
      labels = scales::label_percent(scale = 1),
      limits = c(0, x_max),
      expand = c(0, 0)
    ) +
    ## group colour for boxplot (flip order)
    scale_fill_manual(values = group_cols,
                      labels = c("EHI", "Non-EHI")) +
    ## group colour for jitter (flip order)
    scale_color_manual(values = group_cols,
                       labels = c("EHI", "Non-EHI")) +
    ## flip legend order so control above EHI
    guides(fill = guide_legend(reverse = TRUE)) +
    theme_classic() +
    theme(
      text = font,
      axis.title.x = element_text(size = 18),
      axis.text.x = element_text(size = 16),
      axis.text.y = element_markdown(size = 16, colour = "black"),
      legend.title = element_text(size = 22),
      legend.text  = element_text(size = 18),
      panel.grid.major.x = element_line(colour = "darkgrey", linewidth = 0.25)
    )
  
  list(
    BoxPlot = p2,
    JitterBoxPlot = p3
    )

}  
```

### Phylum
``` {r Section3_plot_abundance_comparison_phylum, fig.height = 5, fig.width = 7}
plot_abundance_comparison("Phylum", nclass = 6)

phylum_abund_group <- plot_abundance_comparison("Phylum", nclass = 6)
phylum_abund_group_p <- phylum_abund_group$JitterBoxPlot
```

### Class
``` {r Section3_plot_abundance_comparison_class, fig.height = 5, fig.width = 7}
plot_abundance_comparison("Class")
```

### Order
``` {r Section3_plot_abundance_comparison_order, fig.height = 5, fig.width = 7}
plot_abundance_comparison("Order")
```

### Family
``` {r Section3_plot_abundance_comparison_family, fig.height = 5, fig.width = 7}
plot_abundance_comparison2("Family")
```

### Genus
``` {r Section3_plot_abundance_comparison_genus, fig.height = 5, fig.width = 7}
plot_abundance_comparison2("Genus")

genus_abund_group <- plot_abundance_comparison2("Genus")
genus_abund_group_p <- genus_abund_group$JitterBoxPlot
```

# Section 4: Differential Abundance Analysis

## DESeq2

In this section, we will use the [DESeq2](https://bioconductor.org/packages/release/bioc/html/DESeq2.html) package in R to identify differentially abundant bacterial species at the OTU level between the conditions. 

### Setup

``` {r Section4_DESeq2_analysis_setup}
## Perform the differential expression analysis
## Run DESeq2 on non-rarefied feature table
deseq_dat <- phyloseq_dat_deseq
## plot the abundance non-rarefied feature table
deseq_plot <- phyloseq_dat_norm_melt 
```

### Group model

``` {r Section4_DESeq2_model_1, cache = TRUE}
## create DESeq2 object from phyloseq object
dds_model1 <- phyloseq_to_deseq2(deseq_dat, design = ~ Group) 

## Pre-filtering
smallestGroupSize <- 3
keep <- rowSums(counts(dds_model1) >= 10) >= smallestGroupSize
dds_model1 <- dds_model1[keep,]

## estimation of size factors was set to use ‘poscounts’, which calculates a modified relative log expression that helps account for features missing in at least one sample
dds_model1 <- DESeq(dds_model1, test = "Wald", fitType = "parametric", sfType = "poscounts")

## Plot deseq2 object
plotDispEsts(dds_model1)
## coefficients DESeq estimated:
resultsNames(dds_model1)

## sum all deviance values and na.rm set as TRUE to ignore NAs
dds_model1_dev <- sum(mcols(dds_model1)$deviance, na.rm = TRUE)
dds_model1_dev

## results
res_model1 <- results(dds_model1, 
                      contrast = c("Group", "EHI", "Control"), 
                      tidy = TRUE, 
                      format = "DataFrame")

rownames(res_model1) <- res_model1$row

res_model1 <- res_model1[,-1]

## summary 
summary(res_model1)

length(res_model1$padj)
```

### Functions

Below are a number of functions that will generate results and plots for the various Group comparisons that will be made.

```{r Section4_DESeq2_analysis_functions}
## Function to take the results from a specific contrast from DESeq2 and return a significance table
signif_OTU <- function (diff_OTU) {
  diff_OTU <-
    subset(diff_OTU,!is.na(padj) &
             padj <= 0.05 & abs(log2FoldChange) >= 1)
  if (nrow(diff_OTU) == 0)
    return(NULL)
  diff_OTU <-
    cbind(as(diff_OTU, "data.frame"), as(tax_table(deseq_dat)[rownames(diff_OTU),], "matrix"))
  diff_OTU <-
    diff_OTU[order(diff_OTU[["padj"]], decreasing = FALSE), ]
  diff_OTU[["OTU"]] <- rownames(diff_OTU)
  diff_OTU[["OTU_num"]] <- 1:nrow(diff_OTU)
  return(diff_OTU)
}

## Function to plot a specific OTU
plot_OTU <- function (otu) {
  ggplot(subset(deseq_plot, OTU == otu),
         aes(x = Group, y = Abundance, fill = Group)) +
    geom_boxplot() +
    ggtitle(paste(c(otu, as.character(unlist(
      subset(deseq_plot, OTU == otu)[1, c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus")]
    ))), collapse = "; ")) +
    labs(x = "", y = "Relative Abundance") +
    scale_fill_manual(values = group_cols) +
    scale_y_continuous(labels = scales::label_percent(scale = 1)) +
    theme(
      text = font,
      plot.title = element_text(size = 8),
      axis.title   = element_text(size = 18),
      axis.text.y  = element_text(size = 16),
      axis.text.x  = element_text(
        angle = 90,
        hjust = 1,
        size = 24
      ),
      legend.text  = element_text(size = 22),
      legend.title = element_text(size = 22, face = "bold")
    )
}

## Function to process a specific comparison
diff_out <- function (i, j, plot = TRUE) {
  ## Get the significant OTUs
  if (i == j) {
    cat ("Error: ", i, " and ", j, " must be different\n")
    return (NULL)
  }
  ## dds_model1 used for ~Group design
  all_OTU <- results(dds_model1, contrast = c("Group", i, j))
  diff_OTU <- signif_OTU(all_OTU)
  if (is.null(diff_OTU)) {
    cat ("Error: No significant differences identified for ",
         i,
         " vs ",
         j,
         "\n")
    return (list(
      "All"    = all_OTU,
      "Signif" = NULL,
      "Up"     = NULL,
      "Down"   = NULL
    ))
  }
  
  ## Save the results
  write.table(
    diff_OTU,
    paste0(
      analysis_dir,
      "/tabs/DESeq2_SIGNIF_OTUs_",
      i,
      "_vs_",
      j,
      "_log2foldchange.csv"
    ),
    row.names = FALSE,
    col.names = TRUE,
    sep = ",",
    quote = FALSE
  )
  
  ## Plot the data for these OTUs
  if (plot) {
    outdir <-
      paste0(analysis_dir, "/figs/SIGNIF_OTUs2_", i, "_vs_", j, "/")
    if (file.exists(outdir)) {
      unlink(outdir, recursive = TRUE)
      dir.create(outdir)
    }
    for (k in diff_OTU[["OTU_num"]]) {
      otu <- rownames(diff_OTU)[k]
      ggsave(
        filename = paste0(outdir, "OTU_", k, "_", otu, ".png"),
        plot = plot_OTU(otu),
        height = 6,
        width = 8,
        unit = "in",
        dpi = 300
      )
    }
  }
  
  ## Return the results
  return (list(
    "All"    = all_OTU,
    "Signif" = diff_OTU,
    "Up"     = subset(diff_OTU, log2FoldChange > 0),
    "Down"   = subset(diff_OTU, log2FoldChange < 0)
  ))
}

## Function to plot a volcano plot for the results using ggplot2
plot_volcano <- function(x) {
  if (length(x) == 0)
    return(NULL)
  
  # Create a data frame for plotting
  plot_dat <- x[["All"]]
  plot_col <- rep("black", nrow(plot_dat))
  which_down <- which(rownames(plot_dat) %in% rownames(x[["Down"]]))
  which_up <- which(rownames(plot_dat) %in% rownames(x[["Up"]]))
  plot_col[which_down] <- "blue"
  plot_col[which_up] <- "red"
  plot_dat_df <- data.frame(
    log2FoldChange = plot_dat$log2FoldChange,
    padj = plot_dat$padj,
    plot_col = factor(plot_col)
  )
  
  # Create the volcano plot using ggplot2
  p <-
    ggplot(data = plot_dat_df, aes(
      x = log2FoldChange,
      y = -log10(padj),
      color = plot_col
    )) +
    geom_point(size = 3,
               #shape = 1,
               alpha = 0.3) +
    geom_hline(
      yintercept = -log10(0.05),
      linetype = "dashed",
      color = "black"
    ) +
    geom_vline(
      xintercept = log2(2),
      linetype = "dashed",
      color = "black"
    ) +
    geom_vline(
      xintercept = -log2(2),
      linetype = "dashed",
      color = "black"
    ) +
    scale_color_manual(values = c("black", "blue", "red")) +
    labs(x = "log2(Fold Change)", y = "-log10(P Value)") +
    # scale_x_continuous(breaks = seq(-6, 6, 2)) +
    theme_classic() +
    theme(
      text = font,
      legend.position = "none",
      plot.title = element_text(size = 12),
      axis.title = element_text(size = 14),
      axis.text = element_text(size = 16),
      legend.text = element_text(size = 12),
      legend.title = element_text(size = 12, face = "bold")
    )
  
  
  return(p)
}

#### modified below:
plot_volcano2 <- function(x) {
  if (length(x) == 0)
    return(NULL)
  
  ## Create a data frame for plotting
  plot_dat <- x[["All"]]
  which_down <- which(rownames(plot_dat) %in% rownames(x[["Down"]]))
  which_up <- which(rownames(plot_dat) %in% rownames(x[["Up"]]))
  plot_col <- rep("black", nrow(plot_dat))
  plot_col[which_down] <- "blue"
  plot_col[which_up] <- "red"
  
  plot_dat_df <- data.frame(
    log2FoldChange = plot_dat$log2FoldChange,
    padj = plot_dat$padj,
    plot_col = factor(plot_col),
    label = rownames(plot_dat)  ## Use rownames for labels to annotate plot
  )
  
  ## Create the volcano plot using ggplot2
  p <-
    ggplot(data = plot_dat_df, aes(
      x = log2FoldChange,
      y = -log10(padj),
      color = plot_col
    )) +
    geom_point(size = 3, alpha = 0.3) +
    geom_hline(
      yintercept = -log10(0.05),
      linetype = "dashed",
      color = "black"
    ) +
    geom_vline(
      xintercept = log2(2),
      linetype = "dashed",
      color = "black"
    ) +
    geom_vline(
      xintercept = -log2(2),
      linetype = "dashed",
      color = "black"
    ) +
    scale_color_manual(values = c("black", "blue", "red")) +
    labs(x = "log2(Fold Change)", y = "-log10(P Value)") +
    theme_classic() +
    theme(
      text = font,
      legend.position = "none",
      plot.title = element_text(size = 12),
      axis.title = element_text(size = 14),
      axis.text = element_text(size = 16),
      legend.text = element_text(size = 12),
      legend.title = element_text(size = 12, face = "bold")
    ) +
    ## Use ggrepel for text repelling, only for OTUs with -log10(padj) > 0.05
    geom_text_repel(
      data = plot_dat_df[which_down | which_up,],
      aes(label = ifelse(
        -log10(padj) > 0.05, as.character(label), ""
      )),
      box.padding = 0.5,
      point.padding = 0.1,
      segment.size = 0.2,
      vjust = 1.5,
      hjust = 0,
      size = 3
    )
  
  return(p)
}

## Function to get the top hits from the DESeq2 analysis
get_top_hits <- function(x, n = 10) {
  if (length(x) == 0)
    return()
  
  # Get the top up and downregulated hits
  positive_df <-
    head(x[["Up"]][order(x[["Up"]][["log2FoldChange"]], decreasing = TRUE),], min(n, nrow(x[["Up"]])))
  negative_df <-
    tail(x[["Down"]][order(x[["Down"]][["log2FoldChange"]], decreasing = TRUE),], min(n, nrow(x[["Down"]])))
  df <- rbind.data.frame(positive_df, negative_df)
  df[["Direction"]] <-
    c(rep("Increase", min(n, nrow(x[["Up"]]))), rep("Decrease", min(n, nrow(x[["Down"]]))))
  
  # Set the sample names from the "Genus" column
  df[["Taxa"]] <-
    factor(paste(otu_or_asv, df[["OTU_num"]], ": ", df[["Genus"]], sep = ""),
           levels = paste(otu_or_asv, df[["OTU_num"]], ": ", df[["Genus"]], sep = ""))
  
  # Create a bar plot using the log2FoldChange values and updated sample names, with colored bars
  df %>%
    ggplot(aes(x = Taxa, y = log2FoldChange, fill = Direction)) +
    geom_bar(stat = 'identity') +
    coord_flip() +
    scale_fill_manual(values = c(
      'Increase' = "#117744",
      'Decrease' = "#DC3220"
    )) +
    ylab("log2 (Fold Change)") + xlab("") +
    # scale_y_continuous(breaks = seq(-6, 6, 2)) +
    theme_classic() +
    theme(
      text = font,
      axis.title   = element_text(size = 14),
      axis.text.y  = element_text(size = 16, face = "bold"),
      axis.text.x  = element_text(size = 14),
      legend.position = "none",
      panel.grid.major.x = element_line(colour = "lightgrey", linewidth = 0.25)
    )
}
```

### Results
The following results show the differentially abundant taxa between the Patient and Control samples. 

``` {r Section4_DESeq2_analysis_cases_vs_control}
DE_EHIvC <- diff_out("EHI", "Control", plot = FALSE)
cat("Total number of differentially abundant OTUs: ", nrow(DE_EHIvC[["Signif"]]))
cat("Upregulated number of differentially abundant OTUs: ", nrow(DE_EHIvC[["Up"]]))
cat("Downregulated number of differentially abundant OTUs: ", nrow(DE_EHIvC[["Down"]]))
```

#### Volcano

Below is a volcano plot showing the overall differential abundance results for this comparison.

``` {r Section4_DESeq2_analysis_cases_vs_control_volcano, fig.height = 5, fig.width = 7}
volcano_p <- plot_volcano(DE_EHIvC)
volcano_p
```

volcano plot with sig ASVs highlighted

``` {r Section4_DESeq2_analysis_cases_vs_control_volcano2, fig.height = 5, fig.width = 7}
volcano_p <- plot_volcano2(DE_EHIvC)
volcano_p
```

## ALDEx2

### Setup

```{r Section4_ALDEx2_setup}
## create ASV_table from phyloseq object
ASV_table = otu_table(phyloseq_dat_deseq)

## create groupings variable from phyloseq object
groupings = phyloseq_dat_deseq %>%
  sample_data() %>%
  .$Group %>%
  ## make groupings a character vector! 
  as.character()

ncol(ASV_table)
length(groupings)
```

### Model

```{r Section4_ALDEx2_model, cache = TRUE}
results <- aldex(reads = ASV_table, 
                 conditions = groupings, 
                 mc.samples = 128, 
                 test="t", 
                 effect=TRUE,
                 include.sample.summary = TRUE, 
                 CI = TRUE,
                 verbose = TRUE, 
                 denom="all")
```

### Results

```{r Section4_ALDEx2_results}
## Expected p-value of Wilcoxon rank test
results %>%
  dplyr::filter(we.eBH < 0.05)

## Expected Benjamini-Hochberg corrected p-value of Wilcoxon test
results %>%
  dplyr::filter(wi.eBH < 0.05)
```

## ANCOMBC

### Setup

https://bioconductor.org/packages/devel/bioc/vignettes/ANCOMBC/inst/doc/ANCOM.html

```{r Section4_ANCOMBC_setup}
## object to use
phyloseq_dat_deseq 
## make a copy for ancom analysis
ancom_phyloseq = phyloseq_dat_deseq
## Set Group variable as character string 
sample_data(ancom_phyloseq)$Group <- as.character(sample_data(ancom_phyloseq)$Group)
## Check that levels are now NULL
levels(sample_data(ancom_phyloseq)$Group)
## Make sure names are still "Control" and "EHI"
unique(sample_data(ancom_phyloseq)$Group)
```

### Model

```{r Section4_ANCOMBC_model}
## use ANCOMBC with phyloseq object
out = ancombc(data = NULL, 
              assay_name = NULL,
              tax_level = NULL, 
              phyloseq = phyloseq_dat_deseq,
              formula = "Group",
              p_adj_method = "BH", 
              prv_cut = 0.10, 
              lib_cut = min_otus,
              group = "Group", 
              struc_zero = TRUE, 
              neg_lb = TRUE,
              tol = 1e-5, 
              max_iter = 100, 
              conserve = FALSE,
              alpha = 0.05, 
              global = TRUE, 
              n_cl = 1, 
              verbose = TRUE)
```

### Results

```{r Section4_ANCOMBC_results}
## Result from the ANCOM-BC log-linear model to determine taxa that are differentially abundant according to the covariate of interest. It contains: 1) log fold changes; 2) standard errors; 3) test statistics; 4) p-values; 5) adjusted p-values; 6) indicators whether the taxon is differentially abundant (TRUE) or not (FALSE).
res = out$res

## Set Colnames
col_names = c("Taxon", "Intercept", "Group")

## 1) log fold changes
res_lfc = res$lfc
colnames(res_lfc) = col_names
#res_lfc
## 2) standard errors
res_se = res$se
colnames(res_se) = col_names
#res_se
## 3) test statistics
res_w = res$W
colnames(res_w) = col_names
#res_w
## 4) p-values
res_p = res$p_val
colnames(res_p) = col_names
#res_p
## 5) adjusted p-values
res_padj = res$q_val
colnames(res_padj) = col_names
#res_padj
## 6) indicators whether the taxon is differentially abundant (TRUE) or not (FALSE).
res_da = res$diff_abn
colnames(res_da) = col_names
#res_da
## check if we have TRUE
unique(res_da$Group)
```

```{r Section4_ANCOM_save_results}
library(openxlsx)
## Create list of ANCOM results
ancom_res_list <- list(
  "Log_Fold_Change" = res_lfc,
  "Standard_Errors" = res_se,
  "Test_Statistics" = res_w,
  "p_value" = res_p,
  "BH_Adjusted_p_value" = res_padj,
  "Dif_Abun_T_or_F" = res_da
)

## Save ANCOM results into a single workbook
openxlsx::write.xlsx(ancom_res_list,
                     file = paste0(analysis_dir, "/tabs/ANCOM_results.xlsx"))
```

## Compare Methods

```{r Section4_compare}
## DESeq2 results to use
deseq_res = DE_EHIvC$Signif
## set rownames as a new colmun called Taxon
deseq_res = rownames_to_column(deseq_res, var = "Taxon")
## count number of sig ASVs using nrow()
nrow(deseq_res)
## Select only Taxon column
deseq_res = deseq_res %>%
  dplyr::select(Taxon)

## ALDEx2 results to use
aldex2_res = results
## set rownames as a new colmun called Taxon
aldex2_res = rownames_to_column(aldex2_res, var = "Taxon")
## Filter for only p adjust < 0.05
aldex2_res = results %>%
  dplyr::filter(wi.eBH < 0.05)
## count number of sig ASVs using nrow()
nrow(aldex2_res)
## As we have 0 hits, we will create an empty vector for venn diagram
aldex2_hits = character(nrow(aldex2_res))

## ANCOMBC results, filter for only p adjust < 0.05
ancombc_res = res_padj %>%
  dplyr::filter(Group < 0.05)
## count number of sig ASVs using nrow()
nrow(ancombc_res)
## Select only Taxon column
ancombc_res = ancombc_res %>%
  dplyr::select(Taxon)
## As we have 0 hits, we will create an empty vector for venn diagram
ancombc_hits = character(nrow(ancombc_res))

## Prepare the a list for venn diagram plotting.
da_list <- list(
  DESeq2 = deseq_res$Taxon,
  ALDEx2 = aldex2_hits,  
  "ANCOM-BC" = ancombc_hits
)
```

### Venn Diagram

```{r Section4_venn_diagram, fig.height = 8, fig.width = 8}
## Change font, code from https://stackoverflow.com/questions/6699950/r-geom-venn-how-to-change-font
venn_font <- function(p, font)
{

  grep_grob <- function(gt, lab){
    which(sapply(gt, function(x) grepl(lab, x$name)))
  }
  
  p2 <- ggplot_gtable(ggplot_build(p))
  mygrobs <- p2$grobs
  panel_grob <- mygrobs[[grep_grob(mygrobs, "panel")]]
  venn_grob <- panel_grob$children[[grep_grob(panel_grob$children, "venn")]]
  text_grob <- venn_grob$children[grep_grob(venn_grob$children, "text")]
  text_grob <- do.call(grid::gList, 
                       lapply(text_grob, 
                              function(x) {x$gp$fontfamily <- font; 
                                           x}))
  venn_grob$children[grep_grob(venn_grob$children, "text")] <- text_grob
  panel_grob$children[[grep_grob(panel_grob$children, "venn")]] <- venn_grob
  mygrobs[[grep_grob(mygrobs, "panel")]] <- panel_grob
  p2$grobs <- mygrobs
  grid::grid.newpage()
  grid::grid.draw(p2)
}

## Combine all taxa into a single vector and make it unique
venn_all_taxa <- unique(unlist(da_list))

## Create a data frame where each row is a unique taxa
venn_df <- data.frame(Taxon = venn_all_taxa)

## For each method, create a column in taxa_df indicating presence or absence of each taxon
for(method in names(da_list)) {
  venn_df[[method]] <- venn_df$Taxon %in% da_list[[method]]
}

## view
venn_df

## create venn diagram
venn_p <- ggplot(venn_df) +
  geom_venn(
    aes(A = DESeq2, B = `ALDEx2`, C = `ANCOM-BC`),
    fill_color = c("#E69F00", "#56B4E9", "green"),
    text_size = 5
  ) +
  coord_fixed() +
  theme_void()

venn_p

## use custom function to set font as serif
venn_font(venn_p, font = "serif")
```

## Microbiome Analysis Panel

```{r Section4_microbiome_panel, fig.height = 22, fig.width = 17}
## Arrange the figures
arrange_row1 <- ggarrange(adiv_group_plot, 
                          bdiv_group_plot,
                          ncol = 2, 
                          labels = LETTERS[1:2], 
                          font.label = list(size = 18, 
                                            face = "bold", 
                                            family = "serif"))


arrange_row2 <- ggarrange( 
                          phylum_abund_group_p, 
                          genus_abund_group_p,
                          ncol = 2, 
                          labels = LETTERS[3:4], 
                          font.label = list(size = 18, 
                                            face = "bold", 
                                            family = "serif"),
                          common.legend = TRUE,
                          legend = "bottom")

## edit bottom margin to allow more space for common legend
arrange_row2 <- arrange_row2 +
  theme(plot.margin = margin(
    t = 0.2,
    r = 0.2,
    b = 0.4,
    l = 0.2,
    unit = "inches"
  ))

## Create an empty plot to use as a spacer
empty_plot <- ggplot() + theme_void()

arrange_row3 <- ggarrange(empty_plot, 
                          fbratio_plot,
                          empty_plot,
                          venn_p,
                          ncol = 4, 
                          labels = c("", LETTERS[5], "", LETTERS[6]), 
                          font.label = list(size = 18, 
                                            face = "bold", 
                                            family = "serif"),
                          widths = c(0.2, 0.5, 0.2, 1))

microbiome_panel <- ggarrange(arrange_row1,
                              arrange_row2,
                              arrange_row3,
                              nrow = 3,
                              heights = c(0.8,1,0.9))

## view 
microbiome_panel

## Save
ggsave(
  filename = paste0(
    analysis_dir,
    "figs/GRANNUS_microbiome_case_control_panel",
    ".png"
  ),
  plot = microbiome_panel,
  height = 22,
  width = 17,
  unit = "in",
  dpi = 600
)

## Save
ggsave(
  filename = paste0(
    analysis_dir,
    "figs/GRANNUS_microbiome_case_control_panel",
    ".jpeg"
  ),
  plot = microbiome_panel,
  height = 22,
  width = 17,
  unit = "in",
  dpi = 600
)
```

# Section 6: Positive Control Analysis
This data set contained samples generated from the [Zymobiomics Microbial Community Standard](https://zymoresearch.eu/collections/zymobiomics-microbial-community-standards/products/zymobiomics-microbial-community-standard). This is used to assess bias and errors in the extraction methods and identify potential contamination across the data set. 

## Theoretical distribution
These samples have a theoretical Composition Based on Genomic DNA which should be represented in the output distributions following data processing alongside the other samples in the data set. This distribution is as follows:

Listeria monocytogenes - 12%, Pseudomonas aeruginosa - 12%, Bacillus subtilis - 12%, Escherichia coli - 12%, Salmonella enterica - 12%, Lactobacillus fermentum - 12%, Enterococcus faecalis - 12%, Staphylococcus aureus - 12%, Saccharomyces cerevisiae - 2%, and Cryptococcus neoformans - 2%

Note that only the bacterial species (all excluding Saccharomyces cerevisiae and Cryptococcus neoformans) will be identified through 16S rRNA amplicon based sequencing. The theoretical distribution for only the bacterial species is shown below:

``` {r Section6_PC_barplot_theoretical_abundance_stacked, fig.height = 10, fig.width = 6.5}
theo <- data.frame(species = c("*Listeria monocytogenes*", "*Pseudomonas aeruginosa*", 
                               "*Bacillus subtilis*", "*Escherichia coli*", 
                               "*Salmonella enterica*", "*Lactobacillus fermentum*", 
                               "*Enterococcus faecalis*", "*Staphylococcus aureus*"),
                   percent = rep(12.5, 8),
                   rep     = "1",
                   group = rep("Positive Control<br>Theoretical Distribution", 8))

theo_cols = c("*Bacillus subtilis*" = "#AA4488", 
              "*Escherichia coli*" = "#771122",
              "*Lactobacillus fermentum*" = "#114477",
              "*Staphylococcus aureus*" = "#77CCCC", 
              "*Salmonella enterica*" = "#DDDD77",
              "*Listeria monocytogenes*" = "#AA7744",
              "*Enterococcus faecalis*" = "#777711",
              "*Pseudomonas aeruginosa*" = "#AAAA44"
              )


theo_plot <- ggplot(theo, aes(x = rep, 
                 y = percent, 
                 fill = species)) + 
  geom_bar(stat = "identity") +
  scale_y_continuous(labels = scales::label_percent(scale = 1),
                     expand = expansion(add = c(0, 1))) +
  scale_fill_manual(values = theo_cols) +
  labs(x = "", 
       y = "Relative Abundance") + 
  facet_wrap(~group) +
  theme_bw() +
  theme(
    text = font,
    axis.title   = element_text(size = 18),
    axis.text.y  = element_text(size = 16),
    axis.text.x  = element_blank(),
    legend.text  = element_markdown(size = 16),
    legend.title = element_blank(),
    strip.text.x = element_markdown(size = 16, face = "bold")) +
  guides(fill = guide_legend(ncol = 1))

theo_plot
```

The total number of taxa (in thousands) for the PC samples is shown below.

``` {r Section6_QC_OTU_count_PC, fig.height = 6, fig.width = 10}
## print counts
data.frame(sample_data(phyloseq_dat_PC)) %>%
  dplyr::mutate(count = sample_sums(phyloseq_dat_PC)) %>%
  #dplyr::mutate(Group = str_replace(Group, "PositiveControl", "Positive Control")) %>%
  dplyr::select(sample_code, Rep, count) 

## plot counts
data.frame(sample_data(phyloseq_dat_PC)) %>%
  dplyr::mutate(count = sample_sums(phyloseq_dat_PC)) %>%
  #dplyr::mutate(Group = str_replace(Group, "PositiveControl", "Positive Control")) %>%
  ggplot(aes(x = Rep, y = count / 1000, fill = Rep)) +
  geom_bar(stat = "identity",
           position = "dodge",
           color = "black") +
  xlab("") + 
  ylab(paste0("Total Number ", otu_or_asv, "s (x1000)")) +
  facet_wrap(. ~ Group, scales = "free_x") +
  geom_hline(yintercept = min_otus / 1000,
             linetype = 2,
             linewidth = 1.2) +
  scale_fill_manual(values = custom_col16) +
  theme_bw() +
  theme(
    text = font,
    axis.title   = element_text(size = 18),
    axis.text.y  = element_text(size = 12),
    axis.text.x  = element_text(
      size = 14,
      angle = 315,
      face = "bold",
      hjust = 0
    ),
    legend.position = "none",
    strip.text.x = element_text(size = 20, face = "bold"),
    strip.text.y = element_text(size = 16, face = "bold")
  )
```

Then we filter to remove the samples with too few taxa, rarefy to the depth of the lowest sample, and normalise to percentage abundance as we did above.

``` {r Section6_PC_analysis_setup}
## Filter to remove samples with fewer than min_otus
phyloseq_dat_PC_filter <- prune_samples(sample_sums(phyloseq_dat_PC) >= min_otus, phyloseq_dat_PC)

## Rarefy the data to the minimum otu depth
set.seed(0)
rarefy_otu_depth <- min(sample_sums(phyloseq_dat_PC_filter))
cat("Rarefication of data to", rarefy_otu_depth, "OTUs:\n")
phyloseq_dat_PC_filter <- rarefy_even_depth(phyloseq_dat_PC_filter, rarefy_otu_depth, verbose = FALSE)

## Normalise the data
phyloseq_dat_PC_filter_norm <- transform_sample_counts(phyloseq_dat_PC_filter, function (x) 100* ((x+1) / sum(x+1)))

## Create data for barplots
phyloseq_dat_PC_filter_norm_melt <- psmelt(phyloseq_dat_PC_filter_norm) %>%
  dplyr::mutate(Genus = ifelse (Genus == "Limosilactobacillus", "Lactobacillus", Genus)) %>% 
  dplyr::mutate(Genus = ifelse (Genus == "Escherichia-Shigella", "Escherichia", Genus)) 
```

## Chi-squared

Below is a chi-squared based comparison between the PC replicates against the expected distribution:

``` {r Section_6_PC_comparison_with_expected}
## Get the theoretical distribution
PC_comp <- theo %>%
  ## remove "*" from names
  dplyr::mutate(species = gsub("\\*", "", species))
PC_comp[["Genus"]] <- gsub(" .*$", "", PC_comp[["species"]])
PC_comp <- PC_comp[, c("Genus", "percent")]
names(PC_comp)[2] <- "Zymo"

## Get the abundance for all of the species in the positive control
PC_abun <- phyloseq_dat_PC_filter_norm_melt %>%
  ## Change where NA is present for Genus column to Salmonella
  dplyr::mutate(Genus = ifelse(!is.na(Family) & Family == "Enterobacteriaceae" & is.na(Genus), "Salmonella", Genus)) %>%
  ## Make uncultured NA
  dplyr::mutate(Genus = ifelse(is.na(Genus) | Genus == "uncultured", NA, Genus)) %>%
  ## Set levels
  dplyr::mutate(Genus = factor(Genus, levels = unique(Genus))) %>%
  dplyr::group_by(Genus, Sample) %>%
  dplyr::summarise(Abundance = sum(Abundance), .groups = "drop") %>%
  dplyr::filter(Genus %in% PC_comp[["Genus"]]) %>%
  tidyr::spread(key = Sample, value = Abundance)

PC_abun

## Combine and save
PC_abun_merge <- merge(PC_comp, PC_abun, by = "Genus")
write.table(
  PC_abun_merge,
  paste0(analysis_dir, "/tabs/PC_abundance_comparison.csv"),
  sep = ",",
  row.names = FALSE,
  col.names = TRUE,
  quote = FALSE
)

## Chi-squared test: all data including theo
csqtest_theo <- chisq.test(PC_abun_merge[, -1])
csqtest_theo

## Just compare replicates to each other
csqtest <- chisq.test(PC_abun_merge[, -c(1:2)])
csqtest
```


## Plot

The distribution of taxa at the Genus level is shown below for the PC samples:

``` {r Section6_PC_barplot_top_genus_normalised_abundance_stacked, fig.height = 8, fig.width = 6.5}
## create new dataframe to use for positive controls
pc_df <- phyloseq_dat_PC_filter_norm_melt %>%
  # Change where NA is present for Genus column to Salmonella
  dplyr::mutate(Genus = ifelse(!is.na(Family) & Family == "Enterobacteriaceae" & is.na(Genus), "Salmonella", Genus)) %>%
  # Make uncultured NA
  dplyr::mutate(Genus = ifelse(is.na(Genus) | Genus == "uncultured", NA, Genus)) %>%
  # Set levels for Genus
  dplyr::mutate(Genus = factor(Genus, levels = unique(Genus)))

## now plot barplot
PC_plot_df <- plot_barplot2("Genus", pc_df, "Positive_Control")
## show plot
PC_plot_df[[1]]  
# PC_plot_df[[2]] 
## Save plot for additional plotting
PC_plot <- PC_plot_df[[1]] 

## chi-squared stats
stat <- round(csqtest$statistic,3)
df <- csqtest$parameter
p_value <- round(csqtest$p.value, 3)

## title
title <- bquote("Chi-Squared: " ~ {X^2}["(" * .(df) * ")"] ~ "=" ~ .(stat) * ", " ~ italic(p) ~ "=" ~ .(p_value))
caption <- "Chi-Squared test comparing all replicates."

PC_plot <- PC_plot +
  scale_x_discrete(labels = c("Rep 1", "Rep 2", "Rep 3", "Rep 4", "Rep 5",
                              "Rep 6", "Rep 7", "Rep 8", "Rep 9", "Rep 10",
                              "Rep 11")) +
  theme_classic() +
  labs(title = title,
       caption = caption) +
  facet_wrap(~Group) +
  theme(
    text = font,
    title = element_text(size = 16),
    plot.caption = element_text(size = 12),
    plot.caption.position = "plot",
    axis.title   = element_text(size = 24),
    axis.text.y  = element_text(size = 16),
    axis.text.x  = element_text(size = 22, angle = 90, vjust = 0.5),
    legend.text  = element_markdown(size = 14),
    legend.title = element_blank(),
    strip.text.x = element_text(size=26, face = "bold")) 

## view
PC_plot

## PC plot with theo, and chi-squared test stats including theo 
## chi-squared stats
stat_theo <- round(csqtest_theo$statistic,3)
df_theo <- csqtest_theo$parameter
p_value_theo <- round(csqtest_theo$p.value, 3)

## title
title <- bquote("Chi-Squared: " ~ {X^2}["(" * .(df_theo) * ")"] ~ "=" ~ .(stat_theo) * ", " ~ italic(p) ~ "=" ~ .(p_value_theo))
caption <- "Chi-Squared test comparing all replicates and theoretical distribution."

PC_plot_theo <- theo_plot +
  theme_classic() +
  labs(title = title,
       caption = caption) +
  theme(
    text = font,
    title = element_text(size = 16),
    plot.caption = element_text(size = 12),
    plot.caption.position = "plot",
    axis.title   = element_text(size = 24),
    axis.text.y  = element_text(size = 16),
    axis.text.x  = element_blank(),
    legend.text  = element_markdown(size = 14),
    legend.title = element_blank(),
    strip.text.x = element_markdown(size=16, face = "bold")) 

PC_plot_theo
```

# Section 7: Negative Control Analysis

## Taxa count
The total number of taxa for the NFW samples is shown below. This should be as low as possible for all replicates 

``` {r Section7_QC_OTU_count_NFW, fig.height = 8, fig.width = 6}
NC_plot <- data.frame(sample_data(phyloseq_dat_NFW)) %>%
  dplyr::mutate(count = sample_sums(phyloseq_dat_NFW)) %>%
  ggplot(aes(x = Rep, y = count, fill = Rep)) +
  geom_bar(stat = "identity", position = "dodge", color = "black") +
  xlab("") + 
  #ylab(paste0("Total Number ", otu_or_asv, "s")) +
  ylab("Read Count") +
  facet_wrap(.~Group, scales = "free_x") +
  #scale_x_discrete(expand=c(0,0)) +
  scale_y_continuous(expand = expansion(add = c(0, 1))) +
  theme_classic() +
    theme(
      text = font,
      axis.title   = element_text(size = 24),
      axis.text.y  = element_text(size = 16),
      axis.text.x  = element_text(size = 22, angle = 90, vjust = 0.5),
      legend.position = "none",
      strip.text.x = element_text(size=26, face = "bold")
      )

## view
NC_plot

## print rownames (otu id) with counts greater than 1
otu_table(phyloseq_dat_NFW)[apply(otu_table(phyloseq_dat_NFW), 1, max) > 1, ]

## check tax table 
## rep 8 is 0005
tax_table(phyloseq_dat_NFW)["0754f8f2affc62076685af91be1cf6f2",]
tax_table(phyloseq_dat_NFW)["d011d3cefa7bde7276d26b61a1c6b0f2",]
## rep 3 is 0011
tax_table(phyloseq_dat_NFW)["81baf44dadb0e7cd124a8420aeae15e8",]

caption <- "Rep 3 contains one unique ASV \"Bacteria\" that is NA.<br>Rep 8 contains two unique ASVs that are \"Unassigned\"."  

## plot
NC_plot_caption <- NC_plot +
  labs(caption = caption) +
  theme(plot.caption = element_markdown(size = 12))

NC_plot_caption
```

## Control Panel

```{r Section7_pos_neg_control_panel, echo = TRUE, fig.height = 10, fig.width = 18.5}
#plot_widths <- c(1,0.8)
## create positive and negative control panel
P_N_con_panel <- ggarrange(
  PC_plot,
  PC_plot_theo,
  NC_plot_caption,
  labels = LETTERS[1:3],
  font.label = list(
    size = 18,
    face = "bold",
    family = "serif"
  ),
  #widths = plot_widths,
  vjust = 1.0,
  ncol = 3,
  nrow = 1
)

## view 
P_N_con_panel

ggsave(
  filename = paste0(
    analysis_dir,
    "figs/ggsave_Section7_pos_neg_control_panel",
    ".png"
  ),
  plot = P_N_con_panel,
  height = 10,
  width = 18.5,
  unit = "in",
  dpi = 300
)
```

Then we filter normalise to percentage abundance as we did above (note that we are not filtering or rarefying these sample, as we want to observe the negative influences on the data).

``` {r Section7_NFW_analysis_setup}
## Normalise the data
phyloseq_dat_NFW_filter_norm <- transform_sample_counts(phyloseq_dat_NFW, function (x) 100* ((x+1) / sum(x+1)))

## Create data for barplots
phyloseq_dat_NFW_filter_norm_melt <- psmelt(phyloseq_dat_NFW_filter_norm)
```

The distribution of taxa at the Genus level is shown below for the NFW samples (note that these data are NOT filtered, as this would remove all samples - so this is based on very low levels of reads). Note that this plot may be influenced by the fact that all taxa had one added to them when normalising to avoid divide by zero errors, so if the read count is low then this plot can be ignored.

``` {r Section7_NFW_barplot_top_genus_normalised_abundance_stacked, fig.height = 8, fig.width = 7}
genus_dat_NFW <- plot_barplot2("Genus",
                               phyloseq_dat_NFW_filter_norm_melt,
                               "negative_control")
genus_dat_NFW[[1]]
```

# Section 8: Sensitivity Analysis

In our primary analysis, we included individuals who had self reported to have taken NSAIDS within the last week prior to the 6MSM. We also included individuals who reported gastrointestinal (GI) obstructive disease or GI surgery. NSAIDS are reported to have an effect on the composition of the gut microbiota [paper link here](https://www.sciencedirect.com/science/article/pii/S1198743X15009027?via%3Dihub). GI obstructive disease or surgery are often used as exclusion criteria within human microbiome studies. However, they be putative risk factors given the highlighted role of the GI system  in exertional heat illness aetiology.  As such, we performed sensitivity analysis. We removed individuals who reported NSAID use (n=13) within the last week and individuals who reported GI obstructive disease or GI surgery (n=1) from our cohort. We also excluded the respective pair matched case or control. 13 pairs were removed. The UIN who reported GI disease or GI surgery, also reported NSAID use. Below is a summary of UINs that reported NSAID use and/or GI disease or GI surgery:

``` {r Section8_sen_data}
## Cohort N prior to removing GI disease/surgery and NSAID users
sample_ann_summary %>%
  dplyr::select(UIN, sample_code, Pair, Group) %>%
  ## Remove Pos & Neg controls using reg expression
  dplyr::filter(!grepl(" Control", Group)) %>%
  dplyr::group_by(Group) %>%
  dplyr::summarise(n = n())

## Create data frame which contains NSAID users
NSAID_users <- sample_ann_summary %>%
  ## Remove Pos & Neg controls using reg expression
  dplyr::filter(!grepl(" Control", Group)) %>%
  ## Filter rows where NSAID is "Yes"
  dplyr::filter(NSAID == "Yes")
## Show NSAID users
NSAID_users

## Create data frame which contains individuals with reported GI disease or surgery
GI_disease <- sample_ann_summary %>%
  ## Remove Pos & Neg controls using reg expression
  dplyr::filter(!grepl(" Control", Group)) %>%
  ## Filter rows where NSAID is "Yes"
  dplyr::filter(GI_disease_surgery == "Yes")
## Show individuals with reported GI disease or surgery
GI_disease

## Identify NSAID pairs to be removed
nsaid_pairs_to_remove <- unique(NSAID_users$Pair)
## Show pairs to be removed
nsaid_pairs_to_remove
## Number of NSAID users
length(nsaid_pairs_to_remove)

## Identify GI disease/surgery pairs to be removed
gi_pairs_to_remove <- unique(GI_disease$Pair)
## Show pairs to be removed
gi_pairs_to_remove
## Number of individuals with reported GI disease or surgery
length(gi_pairs_to_remove)

## Now check for common pairs i.e., are there any pairs which report GI disease/surgery & NSAID use. 
intersect(nsaid_pairs_to_remove, gi_pairs_to_remove)
## Now combine the two vectors together with union to return unique values. 
pairs_to_remove <- union(nsaid_pairs_to_remove, gi_pairs_to_remove)
## Show pairs to be removed
pairs_to_remove
## Number of pairs to be removed
length(pairs_to_remove)
```

## Create Phyloseq object

``` {r Section8_sen_create_phyloseq_object}
## Create a new phyloseq object which removes the 13 pairs (26 UINs) where GI disease or surgery and NSAID use is "Yes"
## use the filtered phyloseq object from primary analysis which has already been rarefied
phyloseq_dat_sen_filter <- prune_samples(!(sample_data(phyloseq_dat_filter)$Pair %in% pairs_to_remove), phyloseq_dat_filter)
## Check that GI_disease_surgery column contains only "No" and positive and negative controls are absent
unique(sample_data(phyloseq_dat_sen_filter)$GI_disease_surgery)
## Check that NSAID column contains only "No" and positive and negative controls are absent
unique(sample_data(phyloseq_dat_sen_filter)$NSAID)

## Create a non-rarefied object for sensitivity DA analysis
## First, create a new phyloseq object which removes the 13 pairs (26 UINs) where GI disease or surgery and NSAID use is "Yes"
phyloseq_dat_sen <- prune_samples(!(sample_data(phyloseq_dat)$Pair %in% pairs_to_remove), phyloseq_dat)
## Now remove any samples with less the 20,000 reads:
## Check sample sums
sample_sums_values <- sample_sums(phyloseq_dat_sen)
## Find samples with sums below 20,000
samples_below_min_otus <- names(sample_sums_values[sample_sums_values <= min_otus])
## Print
print(samples_below_min_otus)
## Now identify the associated pair to also remove
filter_pairs_to_remove_sen <- sample_data(phyloseq_dat)$Pair[sample_data(phyloseq_dat)$sample_code %in% samples_below_min_otus]
## Print
print(filter_pairs_to_remove_sen)
## Create a non rarefied phyloseq object with filtered samples (i.e., remove identified samples with < 20,000 reads) for DESeq2 analysis
phyloseq_dat_sen_deseq <- prune_samples(!(sample_data(phyloseq_dat_sen)$Pair %in% filter_pairs_to_remove), phyloseq_dat_sen)
```

## Quality Control 

### Summary N for Analysis

Below, a summary of the n taken forward for analysis is provided: 

``` {r Section8_sen_n_summary}
## N by group for sensitivity analysis
sample_ann_summary %>%
  dplyr::select(UIN, sample_code, Pair, Classification, Group, NSAID) %>%  
  ## Remove Pos & Neg controls using reg expression
  dplyr::filter(!grepl(" Control", Group)) %>%
  ## Remove NSAID users
  dplyr::filter(!Pair %in% pairs_to_remove) %>%
  ## Remove pairs which have ASV count lower than min_otus (< 20,000)
  # dplyr::filter(!Pair %in% filter_pairs_to_remove_sen) %>%
  dplyr::filter(!Pair %in% filter_pairs_to_remove) %>%
  dplyr::group_by(Group) %>%
  dplyr::summarise(n = n())

## Total N removed by NSAID use
length(pairs_to_remove)
## Total N removed by having ASV count lower than min_outs (< 20,000)
length(filter_pairs_to_remove)

## By classification 
sample_ann_summary %>%
  dplyr::select(UIN, sample_code, Pair, Classification, Group, NSAID) %>%
  ## Remove Pos & Neg controls using reg expression
  dplyr::filter(!grepl(" Control", Group)) %>%
  ## Remove NSAID users
  dplyr::filter(!Pair %in% pairs_to_remove) %>%
  ## Remove pairs which have ASV count lower than min_otus (< 20,000)
  # dplyr::filter(!Pair %in% filter_pairs_to_remove_sen) %>%
  dplyr::filter(!Pair %in% filter_pairs_to_remove) %>%
  dplyr::group_by(Classification) %>%
  dplyr::summarise(n = n())
```

### Taxa count per sample - Rarefied and filtered data

The following plot shows the total number of observed taxa (OTUs or ASVs) per sample after rarefaction and filtering of samples to remove low coverage samples, with the minimum number of taxa specified above (min_otus) highlighted with a dashed line. Samples lower than this will be removed below in Section 2. All samples should have the same depth. If this is not the case, then please debug the code.

``` {r Section8_sen_QC_filterdata_OTU_count, fig.height = 12, fig.width = 10}
data.frame(sample_data(phyloseq_dat_sen_filter)) %>%
  dplyr::mutate(count = sample_sums(phyloseq_dat_sen_filter)) %>%
  ggplot(aes(x = UIN, y = count/1000, fill = Group)) +
  geom_bar(stat = "identity", position = "dodge", color = "black") +
  xlab("") + ylab(paste0("Total Number ", otu_or_asv, "s (x1000)")) +
  facet_wrap(.~Group, scales = "free_x") +
  geom_hline(yintercept = min_otus/1000, linetype = 2, linewidth = 1.2) +
  scale_fill_manual(values = group_cols) +
  scale_color_manual(values = group_cols) +
  theme_bw() +
  theme(
    text = font,
    axis.title   = element_text(size = 18),
    axis.text.y  = element_text(size = 12),
    axis.text.x  = element_text(size = 14, angle = 315, face = "bold", hjust = 0),
    legend.text  = element_text(size = 14),
    legend.title = element_text(size = 20, face = "bold"),
    strip.text.x = element_text(size=20, face = "bold"),
    strip.text.y = element_text(size=16, face = "bold")
  )
```

## Diversity 

### Alpha diversity - Rarefied and filtered data

The following will create a .csv file containing the observed, Chao1 and Shannon Alpha diversity metrics for each sample. Below is the median across each Group level.

```{r Section8_sen_QC_filterdata_alpha_diversity_summary}
## Create alpha diversity data frame
adiv_sen <- data.frame(
  "Pair" = phyloseq::sample_data(phyloseq_dat_sen_filter)$Pair,
  "UIN" = phyloseq::sample_data(phyloseq_dat_sen_filter)$UIN,
  "Group" = phyloseq::sample_data(phyloseq_dat_sen_filter)$Group,
  "Troop" = phyloseq::sample_data(phyloseq_dat_sen_filter)$Troop,
  "Observed" = phyloseq::estimate_richness(phyloseq_dat_sen_filter, measures = "Observed"),
  "Chao1" = phyloseq::estimate_richness(phyloseq_dat_sen_filter, measures = "Chao1"),
  "Shannon" = phyloseq::estimate_richness(phyloseq_dat_sen_filter, measures = "Shannon"),
  "Simpson" = phyloseq::estimate_richness(phyloseq_dat_sen_filter, measures = "Simpson"))


adiv_sen <- adiv_sen %>%
  ## Remove Chao1 SE column
  dplyr::select(-Chao1.se.chao1) %>%
  ## Rename Chao1 column
  dplyr::rename(Chao1 = Chao1.Chao1) %>%
  dplyr::arrange(Pair, Group) %>%
  tibble::remove_rownames()

head(adiv_sen)

## write as csv
write.table(
  adiv_sen,
  paste0(analysis_dir, "/tabs/sensitivity_alpha_diversity_table.csv"),
  row.names = FALSE,
  col.names = TRUE,
  quote = FALSE,
  sep = ","
)
``` 

#### Normality

The following analysis will test for normality of the distribution of Alpha Diversity metrics, which will determine whether analyses should be performed using parametric or non-parametric tests. This table gives summary statistics for a variaty of tests related to the distribution of the data to helkp make this decision. However, generally speaking, if the p-value for the Shapiro-Wilk test is greater than 0.05, then one can assume that the assumption of normality will hold (i.e. there is not sufficient evidence to reject the null hypothesis).

```{r Section8_sen_QC_filterdata_alpha_diversity_normality, echo = TRUE}
jmv::descriptives(
    formula = Observed + Chao1 + Shannon + Simpson ~ Group,
    data = adiv_sen,
    iqr = TRUE,
    sw = TRUE)
```

#### t-test

```{r Section8_sen_QC_filterdata_alpha_diversity_comparison_stat, echo = TRUE}
jmv::ttestIS(
    formula = Shannon ~ Group,
    data = adiv_sen,
    vars = vars(Shannon),
    eqv = TRUE,
    effectSize = TRUE)
```

non-parametric:

```{r Section8_sen_QC_filterdata_alpha_diversity_comparison_stat_non_parametric, echo = TRUE}
jmv::ttestIS(
    formula = Observed + Chao1 + Simpson ~ Group,
    data = adiv_sen,
    vars = vars(Observed, Chao1, Simpson),
    students = FALSE,
    mann = TRUE,
    effectSize = TRUE)
```

#### Plot

The following plot shows a boxplot comparing the alpha diversity between the Group levels (p-values are based in a `r ifelse(test_type == "parametric", "t.test", "wilcox.test")` compared to the '`r control`' level, with ns: p > 0.05; \*: p <= 0.05; \*\*: p <= 0.01; \*\*\*: p <= 0.001; \*\*\*\*: p <= 0.0001). 

```{r Section8_sen_QC_filterdata_alpha_diversity_comparison_plot, fig.height = 8, fig.width = 10}
## TRUE is normal, FALSE is non-normal distribution 
metrics_normality_sen <-
  c(
    "Observed" = FALSE,
    "Chao1" = FALSE,
    "Shannon" = TRUE,
    "Simpson" = FALSE
  )

## Add normality information and tidy data frame
adiv_sen_long <- adiv_sen %>%
  tidyr::gather(key = "metric",
                value = "value",
                c("Observed", "Chao1", "Shannon", "Simpson")) %>%
  dplyr::mutate(metric = factor(metric, levels = c("Observed", "Chao1", "Shannon", "Simpson")),
                Normal = metrics_normality[as.character(metric)])

## Function for Normal data: mean SD is main analysis
## Function for Non-Normal data: median IQR is in main analysis

## Plot
adiv_sen_plot <- adiv_sen_long %>%
  ggplot(aes(x = Group, y = value)) +
  geom_jitter(aes(color = Group), size = 2, alpha = 0.6) +
  facet_wrap(~ metric, scales = "free", ncol = 4) +
  stat_compare_means(comparisons = list(c("EHI", "Control")), 
                     method = "wilcox.test", 
                     label = "p.signif", 
                     paired = FALSE, 
                     family = "serif", 
                     bracket.size = 0.6,
                     vjust = -0.2) +
  scale_fill_manual(values = group_cols) +
  scale_color_manual(values = group_cols) +
  labs(
    x = "",
    y = "Alpha Diversity"
  ) +
  theme_classic() +
  theme(text = font,
        axis.title = element_text(size = 24),
        axis.text.y = element_text(size = 16),
        axis.text.x = element_text(size = 20),
        strip.text = element_text(size=26, face = "bold"),
        legend.position = "none") +
  expand_limits(y=0) +
  stat_summary(data = filter(adiv_sen_long, Normal), 
               fun = mean, 
               geom = "point", 
               color = "black", 
               size = 3, 
               show.legend = FALSE) +
  stat_summary(data = filter(adiv_sen_long, Normal), 
               fun.data = mean_sd, 
               geom = "errorbar", 
               color = "black", 
               width = 0.3, 
               linewidth = 0.8, 
               show.legend = FALSE) +
  stat_summary(data = filter(adiv_sen_long, !Normal), 
               fun = median,
               geom = "crossbar", 
               color = "black", 
               width = 0.6, 
               linewidth = 0.4, 
               show.legend = FALSE) +
  stat_summary(data = filter(adiv_sen_long, !Normal), 
               fun.data = median_iqr, 
               geom = "errorbar", 
               color = "black", 
               width = 0.3, 
               linewidth = 0.6, 
               show.legend = FALSE)

adiv_sen_plot
``` 

### Beta diversity - Rarefied and filtered data

The following principal coordinate analysis (PCoA) plot shows the beta diversity following rarefaction and filtering of samples, based on the Bray Curtis distance between samples. 

``` {r Section8_sen_QC_filterdata_beta_diversity, fig.height = 7, fig.width = 9.5}
## beta diversity - PCoA
bdiv_sen_plot <- plot_ordination(phyloseq_dat_sen_filter, 
                                   ordinate(phyloseq_dat_sen_filter, method = "PCoA", distance = "bray", na.rm = TRUE), 
                                   type = "samples", 
                                   color = "Group"
                                   ) + 
  stat_ellipse(aes(group = Group), 
               linetype = 2,
               linewidth = 1) +
  geom_point(size = 5,
             alpha = 0.8) +
  ggtitle("") + 
  scale_fill_manual(values = group_cols) +
  scale_color_manual(values = group_cols) +
  theme_classic() +
  theme(
    text = font,
    axis.title = element_text(size = 20),
    axis.text = element_text(size = 16),
    legend.text = element_text(size = 18),
    legend.title = element_text(size=22, face = "bold"))

## view
bdiv_sen_plot
```

### PERMANOVA

This section will use permutational MANOVA (PERMANOVA) to look for significant effects of the Group variable on the abundance. Dissimilarity is again based on the Bray Curtis difference. 

``` {r Section8_sen_PERMANOVA_analysis}
## Set seed for reproducibility 
set.seed(0)

## Convert the counts across all samples into a community data matrix
plot_dat_sen <- psmelt(phyloseq_dat_sen_filter)
community_mat_sen <- plot_dat_sen %>%
  dplyr::select(Sample, OTU, Abundance) %>%
  spread(OTU, Abundance)
row.names(community_mat_sen) <- community_mat_sen[["Sample"]]
community_mat_sen <- as.matrix(community_mat_sen[, -1]) ## Remove the Sample column

## Transform the data using square root to minimize influence of most abundant groups
community_mat_sen <- sqrt(community_mat_sen)

## Create a dissimilarity matrix
OTU_dist_sen <- vegdist(community_mat_sen, method = "bray")

## Get the matching sample data
ind_dat_sen <- plot_dat_sen %>%
  dplyr::select(Sample, Group) %>%
  distinct()
ind_dat_sen <- ind_dat_sen[match(rownames(community_mat_sen), ind_dat_sen[["Sample"]]), ]

## Run perMANOVA
perm_sen <- adonis2(OTU_dist_sen ~ Group, data = ind_dat_sen, permutations = 9999, method = "bray")
perm_sen
```

## Microbial Abundance

### Normalise

Here we normalise the taxa counts to represent the percentage of all taxas.

``` {r Section8_sen_normalise}
phyloseq_dat_sen_filter_norm <- transform_sample_counts(phyloseq_dat_sen_filter, function (x) 100* ((x+1) / sum(x+1)))
phyloseq_dat_sen_filter_norm_melt <- psmelt(phyloseq_dat_sen_filter_norm)

## melted object of rarefied data
phyloseq_dat_sen_filter_melt <- psmelt(phyloseq_dat_sen_filter)

## melted object on non-rarefied data
phyloseq_dat_sen_norm_melt <- psmelt(phyloseq_dat_sen_deseq) 
```

### Summerise 

``` {r Section8_sen_summarise_taxa}
## rarefied data: print the number and each unique taxas. Each number total will contain one NA. 
length(unique(phyloseq_dat_sen_filter_norm_melt$Phylum))
unique(phyloseq_dat_sen_filter_norm_melt$Phylum)
length(unique(phyloseq_dat_sen_filter_norm_melt$Class))
unique(phyloseq_dat_sen_filter_norm_melt$Class)
length(unique(phyloseq_dat_sen_filter_norm_melt$Order))
unique(phyloseq_dat_sen_filter_norm_melt$Order)
length(unique(phyloseq_dat_sen_filter_norm_melt$Family))
unique(phyloseq_dat_sen_filter_norm_melt$Family)
length(unique(phyloseq_dat_sen_filter_norm_melt$Genus))
unique(phyloseq_dat_sen_filter_norm_melt$Genus)
```

```{r Section8_sen_summarise_taxa_new}
## Function to calculate distinct counts of taxa for each UIN within Group
summarise_taxa_sen <- function(data) {
  data %>%
    ## Filter by a count >= 1
    dplyr::filter(Abundance >= 1) %>%
    ## Group by UIN and Group
    dplyr::group_by(UIN, Group) %>%
    # Summarise unique counts of taxa at each taxonomic level
    dplyr::summarise(
      Phylum = n_distinct(Phylum, na.rm = TRUE),
      Class = n_distinct(Class, na.rm = TRUE),
      Order = n_distinct(Order, na.rm = TRUE),
      Family = n_distinct(Family, na.rm = TRUE),
      Genus = n_distinct(Genus, na.rm = TRUE),
      Species = n_distinct(Species, na.rm = TRUE),
      ASV = n_distinct(OTU, na.rm = TRUE),
      ## drop groupings
      .groups = 'drop'
    )
}

## Return results for unique taxa
summary_taxa_result_sen <- summarise_taxa(phyloseq_dat_sen_filter_melt)

## View the results
head(summary_taxa_result_sen)

## descriptives with S-W test
jmv::descriptives(
    formula = . ~ Group,
    data = summary_taxa_result_sen,
    iqr = TRUE,
    sw = TRUE)

round_half_up <- function(x) {
  floor(x + 0.5)
}

## summarise average & variance of taxa for each taxa level, by group
summary_by_group_sen <- summary_taxa_result_sen %>%
  dplyr::group_by(Group) %>%
  dplyr::summarise(
    ## phyla: median (IQR)
    median_phyla = median(Phylum, na.rm = FALSE),
    iqr_phyla = IQR(Phylum, na.rm = FALSE),
    ## classes: median (IQR)
    median_classes = median(Class, na.rm = FALSE),
    iqr_classes = IQR(Class, na.rm = FALSE),
    ## orders: mean SD
    mean_orders = mean(Order, na.rm = FALSE),
    sd_orders = sd(Order, na.rm = FALSE),
    ## families: mean SD
    mean_families = mean(Family, na.rm = FALSE),
    sd_families = sd(Family, na.rm = FALSE),
    ## genera: median (IQR)
    median_genera = median(Genus, na.rm = FALSE),
    iqr_genera = IQR(Genus, na.rm = FALSE),
    ## ASV: median (IQR)
    median_ASV = median(ASV, na.rm = FALSE),
    iqr_ASV = IQR(ASV, na.rm = FALSE)
  ) %>%
  ## round numbers to 0 decimal places, rounding .5 up
  dplyr::mutate(across(where(is.numeric), round_half_up))

summary_by_group_sen

## Write this into a file
write.table(
  summary_by_group_sen,
  paste0(
    analysis_dir,
    "/tabs/sensitivity_taxa_summary_table_by_group.csv"
  ),
  row.names = FALSE,
  col.names = TRUE,
  sep = ",",
  quote = FALSE
)
```

### Barplot function:

#### Barplot function: Normal text

The following function will plot a barplot showing the normalised abundance across all taxonomic levels with normal text for taxa. 

``` {r Section8_sen_plot_barplot_normal_text}
## Function for non-italicised taxa names in the legend
plot_barplot_sen <-
  function (taxa, plot_dat = phyloseq_dat_sen_filter_norm_melt) {
    # taxa = "Genus"
    # plot_dat = phyloseq_dat_sen_filter_norm_melt
    
    ## Get the total abundance across all samples for each OTU
    tmp <- plot_dat %>%
      dplyr::group_by(OTU) %>%
      dplyr::summarise(sum = sum(Abundance))
    tsums <- tmp[["sum"]]
    names(tsums) <- tmp[["OTU"]]
    tsums <- tsums[order(names(tsums))]
    
    ## Get the OTU taxa
    tmp <- subset(plot_dat, Sample == plot_dat[1, "Sample"])
    taxlist <- tmp[, taxa]
    names(taxlist) <- tmp[, "OTU"]
    
    ## Check the data match
    if (length(tsums) != length(taxlist) ||
        !all(names(tsums) %in% names(taxlist)) ||
        !all(names(taxlist) %in% names(tsums))) {
      stop("Error: tsums and taxlist do not match.\n")
    }
    taxlist <- taxlist[names(tsums)]
    
    ## Keep only the top taxa and add all other taxa into an "other" class
    taxasum <-
      sort(tapply(tsums, taxlist, sum, na.rm = TRUE), decreasing = TRUE)
    taxasum <- taxasum[!names(taxasum) == ""]
    plot_dat[["TaxName"]] <- as.character(plot_dat[[taxa]])
    plot_dat[is.na(plot_dat[["TaxName"]]) |
               plot_dat[["TaxName"]] == "", "TaxName"] <- "Other"
    plot_dat[["TaxName"]] <-
      ifelse(plot_dat[["TaxName"]] %in% names(taxasum)[1:nclass],
             as.character(plot_dat[[taxa]]),
             "Other")
    plot_dat[["TaxName"]] <-
      factor(plot_dat[["TaxName"]], levels = c(names(taxasum)[1:nclass], "Other"))
    
    ## Combine counts at the taxa level
    plot_dat <- plot_dat %>%
      dplyr::group_by(Sample, Group, UIN, TaxName) %>%
      dplyr::summarise(Abundance = sum(Abundance)) %>%
      ## remove "[]", "-", "_" from taxa names
      dplyr::mutate(TaxName = gsub("\\[|\\]|-|_", " ", TaxName)) %>%
      ## remove leading and trialing spaces!
      dplyr::mutate(TaxName = trimws(TaxName)) %>%
      ## remove instances of doubles spaces
      dplyr::mutate(TaxName = gsub("  ", " ", TaxName))
    
    ## Extract unique and edited TaxName values
    unique_taxa <- unique(plot_dat$TaxName)
    
    ## reset factor levels again
    plot_dat <- plot_dat %>%
      dplyr::mutate(TaxName = factor(TaxName, levels = unique_taxa))
    
    ## Write this into a file
    write.table(
      plot_dat,
      paste0(
        analysis_dir,
        "/tabs/sensitivity_individual_abundance_table_",
        taxa,
        ".csv"
      ),
      row.names = FALSE,
      col.names = TRUE,
      sep = ",",
      quote = FALSE
    )
    
    ## Create a colour pallet for the barplot
    set.seed(8)
    cols <-
      sample(custom_col16, length(unique(plot_dat[["TaxName"]])))
    names(cols) <- unique(plot_dat[["TaxName"]])
    cols["Other"] <- "#D3D3D3" ## set other as grey
    
    ## Create the plot with the custom legend formatting
    p <- plot_dat %>%
      ggplot(aes(x = UIN, y = Abundance, fill = TaxName)) +
      geom_bar(stat = "identity", width = 1) +
      labs(x = "",
           y = "Relative Abundance") +
      scale_y_continuous(labels = scales::label_percent(scale = 1),
                         expand = expansion(add = c(0, 1))) +
      scale_fill_manual(values = cols) +
      facet_wrap(. ~ Group, scales = "free_x") +
      theme_classic() +
      theme(
        text = font,
        axis.title = element_text(size = 18),
        axis.text.y = element_text(size = 16),
        axis.text.x = element_blank(),
        legend.text = element_markdown(size = 14),
        ## Format legend text as markdown so names are italicised
        legend.title = element_blank(),
        strip.text.x = element_text(size = 16, face = "bold"),
        strip.text.y = element_text(size = 16, face = "bold")
      )
    
    p
    
    ## Return the figure and the data set
    return(list(p, plot_dat))
  }
```

#### Barplot function: Italicised Text

The following function will plot a barplot showing the normalised abundance across all taxonomic levels with italicised text for taxa. 

```{r Section8_sen_plot_barplot_italised_text}
## Function for italicised taxa names in the legend
plot_barplot_sen2 <-
  function (taxa, plot_dat = phyloseq_dat_sen_filter_norm_melt) {
    # taxa = "Genus"
    # plot_dat = phyloseq_dat_sen_filter_norm_melt
    
    ## Get the total abundance across all samples for each OTU
    tmp <- plot_dat %>%
      dplyr::group_by(OTU) %>%
      dplyr::summarise(sum = sum(Abundance))
    tsums <- tmp[["sum"]]
    names(tsums) <- tmp[["OTU"]]
    tsums <- tsums[order(names(tsums))]
    
    ## Get the OTU taxa
    tmp <- subset(plot_dat, Sample == plot_dat[1, "Sample"])
    taxlist <- tmp[, taxa]
    names(taxlist) <- tmp[, "OTU"]
    
    ## Check the data match
    if (length(tsums) != length(taxlist) ||
        !all(names(tsums) %in% names(taxlist)) ||
        !all(names(taxlist) %in% names(tsums))) {
      stop("Error: tsums and taxlist do not match.\n")
    }
    taxlist <- taxlist[names(tsums)]
    
    ## Keep only the top taxa and add all other taxa into an "other" class
    taxasum <-
      sort(tapply(tsums, taxlist, sum, na.rm = TRUE), decreasing = TRUE)
    taxasum <- taxasum[!names(taxasum) == ""]
    plot_dat[["TaxName"]] <- as.character(plot_dat[[taxa]])
    plot_dat[is.na(plot_dat[["TaxName"]]) |
               plot_dat[["TaxName"]] == "", "TaxName"] <- "Other"
    plot_dat[["TaxName"]] <-
      ifelse(plot_dat[["TaxName"]] %in% names(taxasum)[1:nclass],
             as.character(plot_dat[[taxa]]),
             "Other")
    plot_dat[["TaxName"]] <-
      factor(plot_dat[["TaxName"]], levels = c(names(taxasum)[1:nclass], "Other"))
    
    ## Combine counts at the taxa level
    plot_dat <- plot_dat %>%
      dplyr::group_by(Sample, Group, UIN, TaxName) %>%
      dplyr::summarise(Abundance = sum(Abundance)) %>%
      dplyr::mutate(## remove "[]", "-", "_" from taxa names
        TaxName = gsub("\\[|\\]|-|_", " ", TaxName),
        ## remove leading and trialing spaces!
        TaxName = trimws(TaxName)) %>%
      ## remove instances of doubles spaces
      dplyr::mutate(TaxName = gsub("  ", " ", TaxName)) %>%
      ## Italicising taxa names
      dplyr::mutate(FormattedTaxName = ifelse(
        ## keep uncultured as normal text
        grepl("^uncultured", TaxName),
        gsub("^(uncultured) (.+)$", "\\1 *\\2*", TaxName),
        ## Keep the stated names (e.g. "Other") as normal text (i.e. not italicised) and wrap taxa names with "*"
        ifelse(
          TaxName %in% c(
            "Other",
            "metagenome",
            "gut metagenome",
            "human gut",
            "unidentified"
          ),
          TaxName,
          paste("*", TaxName, "*", sep = "")
        )
      )) %>%
      ## insert line break in long taxa names
      dplyr::mutate(
        FormattedTaxName = ifelse(
          FormattedTaxName == "*Eubacterium hallii group*",
          "*Eubacterium<br>hallii group*",
          FormattedTaxName
        )
      )

    ## Extract unique and edited TaxName values
    unique_taxa <- unique(plot_dat$TaxName)
    unique_taxa2 <- unique(plot_dat$FormattedTaxName)
    
    ## reset factor levels again
    plot_dat <- plot_dat %>%
      dplyr::mutate(
        TaxName = factor(TaxName, levels = unique_taxa),
        FormattedTaxName = factor(FormattedTaxName, levels = unique_taxa2)
      )
    
    ## Write this into a file
    write.table(
      plot_dat,
      paste0(
        analysis_dir,
        "/tabs/sensitivity_individual_abundance_table_",
        taxa,
        ".csv"
      ),
      row.names = FALSE,
      col.names = TRUE,
      sep = ",",
      quote = FALSE
    )
    
    ## Create a colour pallet for the barplot
    set.seed(8)
    cols <-
      sample(custom_col16, length(unique(plot_dat[["FormattedTaxName"]])))
    names(cols) <- unique(plot_dat[["FormattedTaxName"]])
    cols["Other"] <- "#D3D3D3" ## set other as grey
    
    ## Create the plot with the custom legend formatting
    p <- plot_dat %>%
      ggplot(aes(x = UIN, y = Abundance, fill = FormattedTaxName)) +
      geom_bar(stat = "identity", width = 1) +
      labs(x = "",
           y = "Relative Abundance") +
      scale_y_continuous(labels = scales::label_percent(scale = 1),
                         expand = expansion(add = c(0, 1))) +
      scale_fill_manual(values = cols) +
      facet_wrap(. ~ Group, scales = "free_x") +
      theme_classic() +
      theme(
        text = font,
        axis.title = element_text(size = 18),
        axis.text.y = element_text(size = 16),
        axis.text.x = element_blank(),
        legend.text = element_markdown(size = 14),
        ## Format legend text as markdown so names are italicised
        legend.title = element_blank(),
        strip.text.x = element_text(size = 16, face = "bold"),
        strip.text.y = element_text(size = 16, face = "bold")
      )
    
    p
    
    ## Return the figure and the data set
    return(list(p, plot_dat))
  }
```

#### Phylum
``` {r Section8_sen_barplot_top_phylum_normalised_abundance_stacked, fig.height = 8, fig.width = 8}
phylum_dat_sen <- plot_barplot_sen("Phylum")
phylum_sen_p <- phylum_dat_sen[[1]]
phylum_sen_p
```

#### Class
``` {r Section8_sen_barplot_top_class_normalised_abundance_stacked, fig.height = 8, fig.width = 8}
class_dat_sen <- plot_barplot_sen("Class")
class_dat_sen[[1]]
```

#### Order
``` {r Section8_sen_barplot_top_order_normalised_abundance_stacked, fig.height = 8, fig.width = 8}
order_dat_sen <- plot_barplot_sen("Order")
order_dat_sen[[1]]
```

#### Family
``` {r Section8_sen_barplot_top_family_normalised_abundance_stacked, fig.height = 8, fig.width = 8}
family_dat_sen <- plot_barplot_sen2("Family")
family_dat_sen[[1]]
```

#### Genus
``` {r Section8_sen_barplot_top_genus_normalised_abundance_stacked, fig.height = 8, fig.width = 8}
genus_dat_sen <- plot_barplot_sen2("Genus")
genus_sen_p <- genus_dat_sen[[1]]
genus_sen_p
```

### Firmicute/Bacteroidota ratio

This section will calculate the Firmicutes/Bacteroidota (FB) ratio, which is often used as a measure of gut microbiome dysregulation. 

```{r Section8_sen_calculate_FBratio}
fbratio_sen <- phylum_dat_sen[[2]] %>%
  dplyr::group_by(UIN) %>%
  tidyr::spread(TaxName, Abundance) %>%
  dplyr::mutate(FBratio = Firmicutes/Bacteroidota) %>%
  dplyr::select(UIN, Group, Firmicutes, Bacteroidota, FBratio)

head(fbratio_sen)

## Write as csv
write.table(
  fbratio_sen,
  paste0(analysis_dir, "/tabs/sensitivity_FBratio_table.csv"),
  row.names = FALSE,
  col.names = TRUE,
  sep = ",",
  quote = FALSE
)
```

#### Normality

```{r Section8_sen_FBratio_normality}
jmv::descriptives(
    formula = FBratio ~ Group,
    data = fbratio_sen,
    iqr = TRUE,
    sw = TRUE)
```

#### t-test

```{r Section8_sen_FBratio_comparison_stat}
jmv::ttestIS(
    formula = FBratio ~ Group,
    data = fbratio_sen,
    vars = FBratio,
    students = FALSE,
    mann = TRUE,
    effectSize = TRUE)
```

#### Plot

Below is a boxplot showing how the FB Ratio differs between groups.

``` {r Section8_sen_FBratio_plot, fig.height = 6, fig.width = 6}
## plot figure
fbratio_sen_plot <- fbratio_sen %>% 
  ggplot(aes(x = Group, y = FBratio)) +
  geom_boxplot(outlier.color = NA,
               size = 1) +
  geom_jitter(aes(color = Group),
              size = 2,
              alpha = 0.6) +
  labs(x = "", 
       y = "F:B Ratio") +
  stat_compare_means(comparisons = list(c("EHI", "Control")),
                     method = "wilcox.test",
                     label = "p.signif",
                     paired = FALSE,
                     family = "serif",
                     bracket.size = 0.6,
                     vjust = -0.2) +
  scale_y_continuous(breaks = seq(0, 18, 2), 
                     limits = c(0, 18)) +
  scale_color_manual(values = group_cols) +
  theme_classic() +
  theme(
    text = font,
    axis.title   = element_text(size = 24),
    axis.text.y  = element_text(size = 16),
    axis.text.x  = element_text(size = 22),
    legend.position = "none"
    ) 

## view
fbratio_sen_plot
```

## Abundance Comparison All Taxa Function

Below, we will compare all taxa, at each taxonomic level between groups. 

``` {r Section8_sen_abundance_comparison_function}
## Function for non-italicised taxa names
compare_abundance_comparison_sen <-
  function (taxa,
            plot_dat = phyloseq_dat_sen_filter_norm_melt,
            p_type = "individual") {
    # taxa = "Phylum"
    # plot_dat = phyloseq_dat_sen_filter_norm_melt
    # p_type = "individual"
    
    ## Get the total abundance across all samples for each OTU for the given taxa
    tmp <- plot_dat %>%
      dplyr::group_by(OTU) %>%
      dplyr::summarise(sum = sum(Abundance))
    tsums <- tmp[["sum"]]
    names(tsums) <- tmp[["OTU"]]
    tsums <- tsums[order(names(tsums))]
    
    ## Get the OTU taxa
    tmp <- subset(plot_dat, Sample == plot_dat[1, "Sample"])
    taxlist <- tmp[, taxa]
    names(taxlist) <- tmp[, "OTU"]
    
    ## Check the data match
    if (length(tsums) != length(taxlist) ||
        !all(names(tsums) %in% names(taxlist)) ||
        !all(names(taxlist) %in% names(tsums))) {
      stop("Error: tsums and taxlist do not match.\n")
    }
    taxlist <- taxlist[names(tsums)]
    
    taxasum <-
      sort(tapply(tsums, taxlist, sum, na.rm = TRUE), decreasing = TRUE)
    taxasum <- taxasum[!names(taxasum) == ""]
    plot_dat[["TaxName"]] <- as.character(plot_dat[[taxa]])
    plot_dat[["TaxName"]] <- factor(plot_dat[[taxa]])
    
    ## Combine counts at the taxa level
    plot_dat <- plot_dat %>%
      dplyr::group_by(Pair, UIN, Group, Troop, TaxName) %>%
      dplyr::summarise(Abundance = sum(Abundance), .groups = "drop") %>%
      dplyr::filter(TaxName != "Other")
    
    ## create separate wide data frame for all taxa of each individual UIN
    UIN_taxa_wide <- plot_dat
    UIN_taxa_wide <- UIN_taxa_wide %>%
      tidyr::pivot_wider(names_from = TaxName,
                         values_from = Abundance) %>%
      ## arrange by Pair and Group
      dplyr::arrange(Pair, Group)
  
    ## save as .csv file
    write.table(
      UIN_taxa_wide,
      paste0(
        analysis_dir,
        "/tabs/sensitivity_individual_relative_abundance_",
        taxa,
        ".csv"
      ),
      sep = ",",
      row.names = FALSE,
      col.names = TRUE,
      quote = FALSE
    )
    
    ## Combine at the Group level
    plot_dat_SE <- plot_dat %>%
      dplyr::group_by(Group, TaxName) %>%
      dplyr::summarise(
        average = ifelse(
          test_type == "parametric",
          mean(Abundance),
          median(Abundance)
        ),
        variability = ifelse(
          test_type == "parametric",
          sd(Abundance) / sqrt(length(Abundance)),
          IQR(Abundance)
        )
      )
    
    ## Calculate p-value across all groups
    pval <-
      data.frame(TaxName       = levels(plot_dat_SE[["TaxName"]]),
                 pval          = NA)
    for (t in pval[["TaxName"]]) {
      if (test_type == "parametric") {
        pval[pval[["TaxName"]] == t, "pval"] <-
          anova(lm(Abundance ~ Group, data = subset(plot_dat, TaxName == t)))[1, "Pr(>F)"]
      } else if (test_type == "non-parametric") {
        pval[pval[["TaxName"]] == t, "pval"] <-
          kruskal.test(Abundance ~ Group, data = subset(plot_dat, TaxName == t))[["p.value"]]
      }
    }
    
    ## Calculate individual p-values vs control
    for (t in pval[["TaxName"]]) {
      for (g in unique(plot_dat[["Group"]])) {
        if (g == "Control") {
          pval[pval[["TaxName"]] == t, paste0("pval.", g)] <- 1
        } else if (test_type == "parametric") {
          pval[pval[["TaxName"]] == t, paste0("pval.", g)] <-
            t.test(
              subset(plot_dat, TaxName == t &
                       Group == g)[["Abundance"]],
              subset(plot_dat, TaxName == t &
                       Group == control)[["Abundance"]]
            )[["p.value"]]
        } else if (test_type == "non-parametric") {
          pval[pval[["TaxName"]] == t, paste0("pval.", g)] <-
            wilcox.test(
              subset(plot_dat, TaxName == t &
                       Group == g)[["Abundance"]],
              subset(plot_dat, TaxName == t &
                       Group == control)[["Abundance"]]
            )[["p.value"]]
        }
      }
    }
    
    ## Add to data frame
    plot_dat_SE[["pval"]] <- NA
    for (t in pval[["TaxName"]]) {
      for (g in unique(plot_dat[["Group"]])) {
        if (p_type == "combine") {
          if (g == control) {
            plot_dat_SE[plot_dat_SE[["TaxName"]] == t &
                          plot_dat_SE[["Group"]] == g, "pval"] <-
              pval[pval[["TaxName"]] == t, "pval"]
          } else {
            plot_dat_SE[plot_dat_SE[["TaxName"]] == t &
                          plot_dat_SE[["Group"]] == g, "pval"] <- NA
          }
        } else if (p_type == "individual") {
          if (g == control) {
            plot_dat_SE[plot_dat_SE[["TaxName"]] == t &
                          plot_dat_SE[["Group"]] == g, "pval"] <- NA
          } else {
            plot_dat_SE[plot_dat_SE[["TaxName"]] == t &
                          plot_dat_SE[["Group"]] == g, "pval"] <-
              pval[pval[["TaxName"]] == t, paste0("pval.", g)]
          }
        }
      }
    }
    
    ## Correct for multiple testing
    which_nonna <- which(!is.na(plot_dat_SE[["pval"]]))
    plot_dat_SE[["padj"]] <- NA
    plot_dat_SE[which_nonna, "padj"] <-
      p.adjust(as.numeric(plot_dat_SE[which_nonna,][["pval"]]),
               method = "fdr",
               n = nrow(plot_dat_SE[which_nonna,]))
    
    ## Convert to star format
    plot_dat_SE[["padj_star"]] <-
      case_when(
        plot_dat_SE[["padj"]] <= 0.001 ~ "***",
        plot_dat_SE[["padj"]] <= 0.01  ~ "**",
        plot_dat_SE[["padj"]] <= 0.05  ~ "*",
        TRUE ~ ""
      )
    
    ## Save the values
    write.table(
      plot_dat_SE,
      paste0(
        analysis_dir,
        "/tabs/sensitivity_abundance_comparison_table_",
        taxa,
        ".csv"
      ),
      sep = ",",
      row.names = FALSE,
      col.names = TRUE,
      quote = FALSE
    )
    
    ## format table
    plot_dat_format <- plot_dat_SE %>%
      tidyr::pivot_wider(
        names_from = Group,
        values_from = c(average, variability, pval, padj, padj_star)
      ) %>%
      dplyr::mutate(
        average_Control = sprintf("%.1f", average_Control),
        variability_Control = sprintf("%.1f", variability_Control),
        average_EHI = sprintf("%.1f", average_EHI),
        variability_EHI = sprintf("%.1f", variability_EHI),
        Control = paste0(average_Control, " (", variability_Control, ")"),
        EHI = paste0(average_EHI, " (", variability_EHI, ")")
      ) %>%
      dplyr::rename(pval = pval_EHI,
                    padj = padj_EHI,
                    Taxa = TaxName) %>%
      dplyr::mutate(pval = sprintf("%.3f", as.numeric(pval)),
                    padj = sprintf("%.3f", as.numeric(padj))) %>%
      dplyr::select(Taxa, Control, EHI, pval, padj)
    
    ## save formatted table
    write.table(
      plot_dat_format,
      paste0(
        analysis_dir,
        "/tabs/sensitivity_abundance_comparison_table_formatted_",
        taxa,
        ".csv"
      ),
      sep = ",",
      row.names = FALSE,
      col.names = TRUE,
      quote = FALSE
    )
    
    return(plot_dat_format)
  }
```

### Phylum

```{r Section8_sen_abundance_comparison_Phylum}
phylum_comparison_sen <- compare_abundance_comparison_sen("Phylum")
phylum_comparison_sen

## filter by pval
phylum_comparison_sen %>%
  dplyr::filter(pval < 0.05)

## filter by padj
phylum_comparison_sen %>%
  dplyr::filter(padj < 0.05)
```

### Class

```{r Section8_sen_abundance_comparison_Class}
class_comparison_sen <- compare_abundance_comparison_sen("Class")
class_comparison_sen

## filter by pval
class_comparison_sen %>%
  dplyr::filter(pval < 0.05)

## filter by padj
class_comparison_sen %>%
  dplyr::filter(padj < 0.05)
```

### Order

```{r Section8_sen_abundance_comparison_Order}
order_comparison_sen <- compare_abundance_comparison_sen("Order")
order_comparison_sen

## filter by pval
order_comparison_sen %>%
  dplyr::filter(pval < 0.05)

## filter by padj
order_comparison_sen %>%
  dplyr::filter(padj < 0.05)
```

### Family

```{r Section8_sen_abundance_comparison_Family}
family_comparison_sen <- compare_abundance_comparison_sen("Family")
family_comparison_sen

## filter by pval
family_comparison_sen %>%
  dplyr::filter(pval < 0.05)

## filter by padj
family_comparison_sen %>%
  dplyr::filter(padj < 0.05)
```

### Genus

```{r Section8_sen_abundance_comparison_Genus}
genus_comparison_sen <- compare_abundance_comparison_sen("Genus")
genus_comparison_sen

## filter by pval
genus_comparison_sen %>%
  dplyr::filter(pval < 0.05)

## filter by padj
genus_comparison_sen %>%
  dplyr::filter(padj < 0.05)
```


## Plot Abundance Comparison Function: 

### Abundance Comparison Function: Normal Text

The following function will plot the difference between the groups at different taxonomic levels with normal text for taxa:

``` {r Section8_sen_plot_abundance_comparison_normal_text}
## Function for non-italicised taxa names
plot_abundance_comparison_sen <- function (taxa,
                                           plot_dat = phyloseq_dat_sen_filter_norm_melt,
                                           p_type = "individual",
                                           nclass = 15) {
  # taxa = "Genus"
  # plot_dat = phyloseq_dat_sen_filter_norm_melt
  # p_type = "individual"
  
  ## Get the total abundance across all samples for each OTU for the given taxa
  tmp <- plot_dat %>%
    dplyr::group_by(OTU) %>%
    dplyr::summarise(sum = sum(Abundance))
  tsums <- tmp[["sum"]]
  names(tsums) <- tmp[["OTU"]]
  tsums <- tsums[order(names(tsums))]
  
  ## Get the OTU taxa
  tmp <- subset(plot_dat, Sample == plot_dat[1, "Sample"])
  taxlist <- tmp[, taxa]
  names(taxlist) <- tmp[, "OTU"]
  
  ## Check the data match
  if (length(tsums) != length(taxlist) ||
      !all(names(tsums) %in% names(taxlist)) ||
      !all(names(taxlist) %in% names(tsums))) {
    stop("Error: tsums and taxlist do not match.\n")
  }
  taxlist <- taxlist[names(tsums)]
  
  ## Keep only the top taxa and add all other taxa into an "other" class
  taxasum <-
    sort(tapply(tsums, taxlist, sum, na.rm = TRUE), decreasing = TRUE)
  taxasum <- taxasum[!names(taxasum) == ""]
  plot_dat[["TaxName"]] <- as.character(plot_dat[[taxa]])
  plot_dat[is.na(plot_dat[["TaxName"]]) |
             plot_dat[["TaxName"]] == "", "TaxName"] <- "Other"
  plot_dat[["TaxName"]] <-
    ifelse(plot_dat[["TaxName"]] %in% names(taxasum)[1:nclass],
           as.character(plot_dat[[taxa]]),
           "Other")
  plot_dat[["TaxName"]] <-
    factor(plot_dat[["TaxName"]], levels = c(names(taxasum)[1:nclass], "Other"))
  
  ## Combine counts at the taxa level
  plot_dat <- plot_dat %>%
    dplyr::group_by(UIN, Group, TaxName) %>%
    dplyr::summarise(Abundance = sum(Abundance), .groups = "drop") %>%
    dplyr::filter(TaxName != "Other")
  
  ## Combine at the Group level
  plot_dat_SE <- plot_dat %>%
    dplyr::group_by(Group, TaxName) %>%
    dplyr::summarise(
      average = ifelse(
        test_type == "parametric",
        mean(Abundance),
        median(Abundance)
      ),
      variability = ifelse(
        test_type == "parametric",
        sd(Abundance) / sqrt(length(Abundance)),
        IQR(Abundance)
      )
    )
  
  ## Calculate p-value across all groups
  pval <-
    data.frame(TaxName       = head(levels(plot_dat_SE[["TaxName"]]),-1),
               pval          = NA)
  for (t in pval[["TaxName"]]) {
    if (test_type == "parametric") {
      pval[pval[["TaxName"]] == t, "pval"] <-
        anova(lm(Abundance ~ Group, data = subset(plot_dat, TaxName == t)))[1, "Pr(>F)"]
    } else if (test_type == "non-parametric") {
      pval[pval[["TaxName"]] == t, "pval"] <-
        kruskal.test(Abundance ~ Group, data = subset(plot_dat, TaxName == t))[["p.value"]]
    }
  }
  
  ## Calculate individual p-values vs control
  for (t in pval[["TaxName"]]) {
    for (g in unique(plot_dat[["Group"]])) {
      if (g == "Control") {
        pval[pval[["TaxName"]] == t, paste0("pval.", g)] <- 1
      } else if (test_type == "parametric") {
        pval[pval[["TaxName"]] == t, paste0("pval.", g)] <-
          t.test(
            subset(plot_dat, TaxName == t &
                     Group == g)[["Abundance"]],
            subset(plot_dat, TaxName == t &
                     Group == control)[["Abundance"]]
          )[["p.value"]]
      } else if (test_type == "non-parametric") {
        pval[pval[["TaxName"]] == t, paste0("pval.", g)] <-
          wilcox.test(
            subset(plot_dat, TaxName == t &
                     Group == g)[["Abundance"]],
            subset(plot_dat, TaxName == t &
                     Group == control)[["Abundance"]]
          )[["p.value"]]
      }
    }
  }
  
  ## Add to data frame
  plot_dat_SE[["pval"]] <- NA
  for (t in pval[["TaxName"]]) {
    for (g in unique(plot_dat[["Group"]])) {
      if (p_type == "combine") {
        if (g == control) {
          plot_dat_SE[plot_dat_SE[["TaxName"]] == t &
                        plot_dat_SE[["Group"]] == g, "pval"] <-
            pval[pval[["TaxName"]] == t, "pval"]
        } else {
          plot_dat_SE[plot_dat_SE[["TaxName"]] == t &
                        plot_dat_SE[["Group"]] == g, "pval"] <- NA
        }
      } else if (p_type == "individual") {
        if (g == control) {
          plot_dat_SE[plot_dat_SE[["TaxName"]] == t &
                        plot_dat_SE[["Group"]] == g, "pval"] <- NA
        } else {
          plot_dat_SE[plot_dat_SE[["TaxName"]] == t &
                        plot_dat_SE[["Group"]] == g, "pval"] <-
            pval[pval[["TaxName"]] == t, paste0("pval.", g)]
        }
      }
    }
  }
  
  ## Correct for multiple testing
  which_nonna <- which(!is.na(plot_dat_SE[["pval"]]))
  plot_dat_SE[["pval_adj"]] <- NA
  plot_dat_SE[which_nonna, "pval_adj"] <-
    p.adjust(as.numeric(plot_dat_SE[which_nonna,][["pval"]]),
             method = "fdr",
             n = nrow(plot_dat_SE[which_nonna,]))
  
  ## Convert to star format
  plot_dat_SE[["pval_adj_star"]] <-
    case_when(
      plot_dat_SE[["pval_adj"]] <= 0.001 ~ "***",
      plot_dat_SE[["pval_adj"]] <= 0.01  ~ "**",
      plot_dat_SE[["pval_adj"]] <= 0.05  ~ "*",
      TRUE ~ ""
    )
  
  ## Save the values
  write.table(
    plot_dat_SE,
    paste0(
      analysis_dir,
      "/tabs/sensitivity_plot_abundance_comparison_",
      taxa,
      ".csv"
    ),
    sep = ",",
    row.names = FALSE,
    col.names = TRUE,
    quote = FALSE
  )
  
  ## Plot the data

  ## Determine x-axis limits
  x_max <-
    max(plot_dat$Abundance, na.rm = TRUE) * 1.02 ## Extend by 2%
  
  ## box plots
  p2 <- plot_dat %>%
    ## remove "[]", "-", "_" from taxa names
    dplyr::mutate(TaxName = gsub("\\[|\\]|-|_", " ", TaxName)) %>%
    ## remove leading and trialing spaces!
    dplyr::mutate(TaxName = trimws(TaxName)) %>%
    ## remove instances of double spaces
    dplyr::mutate(TaxName = gsub("  ", " ", TaxName)) %>%
    ggplot(aes(
      x = Abundance,
      y = reorder(TaxName, Abundance),
      ## flip group order
      fill = factor(Group, levels = c("EHI", "Control"))
    )) +
    geom_boxplot(
      width = 0.6,
      position = position_dodge(0.8),
      outlier.size = 1,
      outlier.alpha = 0.8
    ) +
    labs(x = "Relative Abundance",
         y = "",
         ## legend label
         fill = "Group") +
    scale_y_discrete(expand = c(0, 0.5)) +
    scale_x_continuous(
      labels = scales::label_percent(scale = 1),
      limits = c(0, x_max),
      expand = c(0, 0)
    ) +
    ## group colour for boxplot
    scale_fill_manual(values = group_cols) +
    ## group colour for jitter
    scale_color_manual(values = group_cols) +
    ## flip legend order so control above EHI
    guides(fill = guide_legend(reverse = TRUE)) +
    theme_classic() +
    theme(
      text = font,
      axis.title.x = element_text(size = 18),
      axis.text.x = element_text(size = 16),
      axis.text.y = element_markdown(size = 16, colour = "black"),
      legend.title = element_text(size = 22),
      legend.text  = element_text(size = 18),
      panel.grid.major.x = element_line(colour = "darkgrey", linewidth = 0.25)
    )
  
  ## boxplot with individual data points
  p3 <- plot_dat %>%
    ## remove "[]", "-", "_" from taxa names
    dplyr::mutate(TaxName = gsub("\\[|\\]|-|_", " ", TaxName)) %>%
    ## remove leading and trialing spaces!
    dplyr::mutate(TaxName = trimws(TaxName)) %>%
    ## remove instances of double spaces
    dplyr::mutate(TaxName = gsub("  ", " ", TaxName)) %>%
    ggplot(aes(
      x = Abundance,
      y = reorder(TaxName, Abundance),
      ## flip group order
      fill = factor(Group, levels = c("EHI", "Control"))
    )) +
    geom_jitter(
      ## colour jitter by group but flip factors to match fill global aesthetic
      aes(colour = factor(Group, levels = c("EHI", "Control"))),
      ## set jitter width and keep dodge consistent
      position = position_jitterdodge(jitter.width = 0.15, dodge.width = 0.8),
      size = 1.2,
      alpha = 0.6,
      show.legend = FALSE,
    ) +
    geom_boxplot(width = 0.6,
                 position = position_dodge(0.8),
                 ## remove outlier dots as we have jitter points
                 outlier.shape = NA) +
    labs(x = "Relative Abundance",
         y = "",
         ## legend label
         fill = "Group") +
    scale_y_discrete(expand = c(0, 0.5)) +
    scale_x_continuous(
      labels = scales::label_percent(scale = 1),
      limits = c(0, x_max),
      expand = c(0, 0)
    ) +
    ## group colour for boxplot
    scale_fill_manual(values = group_cols) +
    ## group colour for jitter
    scale_color_manual(values = group_cols) +
    ## flip legend order so control above EHI
    guides(fill = guide_legend(reverse = TRUE)) +
    theme_classic() +
    theme(
      text = font,
      axis.title.x = element_text(size = 18),
      axis.text.x = element_text(size = 16),
      axis.text.y = element_markdown(size = 16, colour = "black"),
      legend.title = element_text(size = 22),
      legend.text  = element_text(size = 18),
      panel.grid.major.x = element_line(colour = "darkgrey", linewidth = 0.25)
    )
  
  list(
    BoxPlot = p2,
    JitterBoxPlot = p3
    )

} 
```

### Abundance Comparison Function: Italicised Text

The following function will plot the difference between the groups at different taxonomic levels with italicised text for taxa:

```{r Section8_sen_plot_abundance_comparison_italised_text}
## Function for italicised taxa names
plot_abundance_comparison_sen2 <- function (taxa,
                                            plot_dat = phyloseq_dat_sen_filter_norm_melt,
                                            p_type = "individual",
                                            nclass = 15) {
  # taxa = "Genus"
  # plot_dat = phyloseq_dat_sen_filter_norm_melt
  # p_type = "individual"
  
  ## Get the total abundance across all samples for each OTU for the given taxa
  tmp <- plot_dat %>%
    dplyr::group_by(OTU) %>%
    dplyr::summarise(sum = sum(Abundance))
  tsums <- tmp[["sum"]]
  names(tsums) <- tmp[["OTU"]]
  tsums <- tsums[order(names(tsums))]
  
  ## Get the OTU taxa
  tmp <- subset(plot_dat, Sample == plot_dat[1, "Sample"])
  taxlist <- tmp[, taxa]
  names(taxlist) <- tmp[, "OTU"]
  
  ## Check the data match
  if (length(tsums) != length(taxlist) ||
      !all(names(tsums) %in% names(taxlist)) ||
      !all(names(taxlist) %in% names(tsums))) {
    stop("Error: tsums and taxlist do not match.\n")
  }
  taxlist <- taxlist[names(tsums)]
  
  ## Keep only the top taxa and add all other taxa into an "other" class
  taxasum <-
    sort(tapply(tsums, taxlist, sum, na.rm = TRUE), decreasing = TRUE)
  taxasum <- taxasum[!names(taxasum) == ""]
  plot_dat[["TaxName"]] <- as.character(plot_dat[[taxa]])
  plot_dat[is.na(plot_dat[["TaxName"]]) |
             plot_dat[["TaxName"]] == "", "TaxName"] <- "Other"
  plot_dat[["TaxName"]] <-
    ifelse(plot_dat[["TaxName"]] %in% names(taxasum)[1:nclass],
           as.character(plot_dat[[taxa]]),
           "Other")
  plot_dat[["TaxName"]] <-
    factor(plot_dat[["TaxName"]], levels = c(names(taxasum)[1:nclass], "Other"))
  
  ## Combine counts at the taxa level
  plot_dat <- plot_dat %>%
    dplyr::group_by(UIN, Group, TaxName) %>%
    dplyr::summarise(Abundance = sum(Abundance), .groups = "drop") %>%
    dplyr::filter(TaxName != "Other")
  
  ## Combine at the Group level
  plot_dat_SE <- plot_dat %>%
    dplyr::group_by(Group, TaxName) %>%
    dplyr::summarise(
      average = ifelse(
        test_type == "parametric",
        mean(Abundance),
        median(Abundance)
      ),
      variability = ifelse(
        test_type == "parametric",
        sd(Abundance) / sqrt(length(Abundance)),
        IQR(Abundance)
      )
    )
  
  ## Calculate p-value across all groups
  pval <-
    data.frame(TaxName       = head(levels(plot_dat_SE[["TaxName"]]),-1),
               pval          = NA)
  for (t in pval[["TaxName"]]) {
    if (test_type == "parametric") {
      pval[pval[["TaxName"]] == t, "pval"] <-
        anova(lm(Abundance ~ Group, data = subset(plot_dat, TaxName == t)))[1, "Pr(>F)"]
    } else if (test_type == "non-parametric") {
      pval[pval[["TaxName"]] == t, "pval"] <-
        kruskal.test(Abundance ~ Group, data = subset(plot_dat, TaxName == t))[["p.value"]]
    }
  }
  
  ## Calculate individual p-values vs control
  for (t in pval[["TaxName"]]) {
    for (g in unique(plot_dat[["Group"]])) {
      if (g == "Control") {
        pval[pval[["TaxName"]] == t, paste0("pval.", g)] <- 1
      } else if (test_type == "parametric") {
        pval[pval[["TaxName"]] == t, paste0("pval.", g)] <-
          t.test(
            subset(plot_dat, TaxName == t &
                     Group == g)[["Abundance"]],
            subset(plot_dat, TaxName == t &
                     Group == control)[["Abundance"]]
          )[["p.value"]]
      } else if (test_type == "non-parametric") {
        pval[pval[["TaxName"]] == t, paste0("pval.", g)] <-
          wilcox.test(
            subset(plot_dat, TaxName == t &
                     Group == g)[["Abundance"]],
            subset(plot_dat, TaxName == t &
                     Group == control)[["Abundance"]]
          )[["p.value"]]
      }
    }
  }
  
  ## Add to data frame
  plot_dat_SE[["pval"]] <- NA
  for (t in pval[["TaxName"]]) {
    for (g in unique(plot_dat[["Group"]])) {
      if (p_type == "combine") {
        if (g == control) {
          plot_dat_SE[plot_dat_SE[["TaxName"]] == t &
                        plot_dat_SE[["Group"]] == g, "pval"] <-
            pval[pval[["TaxName"]] == t, "pval"]
        } else {
          plot_dat_SE[plot_dat_SE[["TaxName"]] == t &
                        plot_dat_SE[["Group"]] == g, "pval"] <- NA
        }
      } else if (p_type == "individual") {
        if (g == control) {
          plot_dat_SE[plot_dat_SE[["TaxName"]] == t &
                        plot_dat_SE[["Group"]] == g, "pval"] <- NA
        } else {
          plot_dat_SE[plot_dat_SE[["TaxName"]] == t &
                        plot_dat_SE[["Group"]] == g, "pval"] <-
            pval[pval[["TaxName"]] == t, paste0("pval.", g)]
        }
      }
    }
  }
  
  ## Correct for multiple testing
  which_nonna <- which(!is.na(plot_dat_SE[["pval"]]))
  plot_dat_SE[["pval_adj"]] <- NA
  plot_dat_SE[which_nonna, "pval_adj"] <-
    p.adjust(as.numeric(plot_dat_SE[which_nonna,][["pval"]]),
             method = "fdr",
             n = nrow(plot_dat_SE[which_nonna,]))
  
  ## Convert to star format
  plot_dat_SE[["pval_adj_star"]] <-
    case_when(
      plot_dat_SE[["pval_adj"]] <= 0.001 ~ "***",
      plot_dat_SE[["pval_adj"]] <= 0.01  ~ "**",
      plot_dat_SE[["pval_adj"]] <= 0.05  ~ "*",
      TRUE ~ ""
    )
  
  ## Save the values
  write.table(
    plot_dat_SE,
    paste0(
      analysis_dir,
      "/tabs/sensitivity_plot_abundance_comparison_",
      taxa,
      ".csv"
    ),
    sep = ",",
    row.names = FALSE,
    col.names = TRUE,
    quote = FALSE
  )
  
  ## Plot the data

  ## Determine x-axis limits
  x_max <-
    max(plot_dat$Abundance, na.rm = TRUE) * 1.02 ## Extend by 2%
  
  ## box plots
  p2 <- plot_dat %>%
    ## remove "[]", "-", "_" from taxa names
    dplyr::mutate(TaxName = gsub("\\[|\\]|-|_", " ", TaxName)) %>%
    ## remove leading and trialing spaces!
    dplyr::mutate(TaxName = trimws(TaxName)) %>%
    ## remove instances of double spaces
    dplyr::mutate(TaxName = gsub("  ", " ", TaxName)) %>%
    dplyr::mutate(
      FormattedTaxName = ifelse(TaxName != "Other", paste("*", TaxName, "*", sep = ""), TaxName),
      FormattedTaxName = ifelse(
        FormattedTaxName == "*Eubacterium hallii group*",
        "*Eubacterium<br>hallii group*",
        FormattedTaxName
      )
    ) %>%
    ggplot(aes(
      x = Abundance,
      y = reorder(FormattedTaxName, Abundance),
      ## flip group order
      fill = factor(Group, levels = c("EHI", "Control"))
    )) +
    geom_boxplot(
      width = 0.6,
      position = position_dodge(0.8),
      outlier.size = 1,
      outlier.alpha = 0.8
    ) +
    labs(x = "Relative Abundance",
         y = "",
         ## legend label
         fill = "Group") +
    scale_y_discrete(expand = c(0, 0.5)) +
    scale_x_continuous(
      labels = scales::label_percent(scale = 1),
      limits = c(0, x_max),
      expand = c(0, 0)
    ) +
    ## group colour for boxplot
    scale_fill_manual(values = group_cols) +
    ## group colour for jitter
    scale_color_manual(values = group_cols) +
    ## flip legend order so control above EHI
    guides(fill = guide_legend(reverse = TRUE)) +
    theme_classic() +
    theme(
      text = font,
      axis.title.x = element_text(size = 18),
      axis.text.x = element_text(size = 16),
      axis.text.y = element_markdown(size = 16, colour = "black"),
      legend.title = element_text(size = 22),
      legend.text  = element_text(size = 18),
      panel.grid.major.x = element_line(colour = "darkgrey", linewidth = 0.25)
    )
  
  ## boxplot with individual data points
  p3 <- plot_dat %>%
    ## remove "[]", "-", "_" from taxa names
    dplyr::mutate(TaxName = gsub("\\[|\\]|-|_", " ", TaxName)) %>%
    ## remove leading and trialing spaces!
    dplyr::mutate(TaxName = trimws(TaxName)) %>%
    ## remove instances of double spaces
    dplyr::mutate(TaxName = gsub("  ", " ", TaxName)) %>%
    dplyr::mutate(
      FormattedTaxName = ifelse(TaxName != "Other", paste("*", TaxName, "*", sep = ""), TaxName),
      FormattedTaxName = ifelse(
        FormattedTaxName == "*Eubacterium hallii group*",
        "*Eubacterium<br>hallii group*",
        FormattedTaxName
      )
    ) %>%
    ggplot(aes(
      x = Abundance,
      y = reorder(FormattedTaxName, Abundance),
      ## flip group order
      fill = factor(Group, levels = c("EHI", "Control"))
    )) +
    geom_jitter(
      ## colour jitter by group but flip factors to match fill global aesthetic
      aes(colour = factor(Group, levels = c("EHI", "Control"))),
      ## set jitter width and keep dodge consistent
      position = position_jitterdodge(jitter.width = 0.15, dodge.width = 0.8),
      size = 1.2,
      alpha = 0.6,
      show.legend = FALSE,
    ) +
    geom_boxplot(width = 0.6,
                 position = position_dodge(0.8),
                 ## remove outlier dots as we have jitter points
                 outlier.shape = NA) +
    labs(x = "Relative Abundance",
         y = "",
         ## legend label
         fill = "Group") +
    scale_y_discrete(expand = c(0, 0.5)) +
    scale_x_continuous(
      labels = scales::label_percent(scale = 1),
      limits = c(0, x_max),
      expand = c(0, 0)
    ) +
    ## group colour for boxplot
    scale_fill_manual(values = group_cols) +
    ## group colour for jitter
    scale_color_manual(values = group_cols) +
    ## flip legend order so control above EHI
    guides(fill = guide_legend(reverse = TRUE)) +
    theme_classic() +
    theme(
      text = font,
      axis.title.x = element_text(size = 18),
      axis.text.x = element_text(size = 16),
      axis.text.y = element_markdown(size = 16, colour = "black"),
      legend.title = element_text(size = 22),
      legend.text  = element_text(size = 18),
      panel.grid.major.x = element_line(colour = "darkgrey", linewidth = 0.25)
    )
  
  list(
    BoxPlot = p2,
    JitterBoxPlot = p3
    )

}
```

### Phylum
``` {r Section8_sen_plot_abundance_comparison_phylum, fig.height = 5, fig.width = 7}
plot_abundance_comparison_sen("Phylum", nclass = 6)

phylum_abund_group_sen <- plot_abundance_comparison_sen("Phylum", nclass = 6)
phylum_abund_group_sen_p <- phylum_abund_group_sen$JitterBoxPlot
```

### Class
``` {r Section8_sen_plot_abundance_comparison_class, fig.height = 5, fig.width = 7}
plot_abundance_comparison_sen("Class")
```

### Order
``` {r Section8_sen_plot_abundance_comparison_order, fig.height = 5, fig.width = 7}
plot_abundance_comparison_sen("Order")
```

### Family
``` {r Section8_sen_plot_abundance_comparison_family, fig.height = 5, fig.width = 7}
plot_abundance_comparison_sen2("Family")
```

### Genus
``` {r Section8_sen_plot_abundance_comparison_genus, fig.height = 5, fig.width = 7}
plot_abundance_comparison_sen("Genus")

genus_abund_group_sen <- plot_abundance_comparison_sen("Genus")
genus_abund_group_sen_p <- genus_abund_group_sen$JitterBoxPlot
```

## Differential Abundance Analysis

### DESeq2

In this section, we will use the [DESeq2](https://bioconductor.org/packages/release/bioc/html/DESeq2.html) package in R to identify differentially abundant bacterial species at the OTU level between the conditions. 

#### Setup

``` {r Section8_sen_DESeq2_analysis_setup}
## Perform the differential expression analysis
## Run DESeq2 on non-rarefied feature table
deseq_dat_sen <- phyloseq_dat_sen_deseq
## plot the abundance non-rarefied feature table
deseq_plot_sen <- phyloseq_dat_sen_norm_melt 
```

#### Model

``` {r Section8_sen_DESeq2_model_1, cache = TRUE}
## create DESeq2 object from phyloseq object
dds_model1 <- phyloseq_to_deseq2(deseq_dat_sen, design = ~ Group) 

## Pre-filtering
smallestGroupSize <- 3
keep <- rowSums(counts(dds_model1) >= 10) >= smallestGroupSize
dds_model1 <- dds_model1[keep,]

## estimation of size factors was set to use ‘poscounts’, which calculates a modified relative log expression that helps account for features missing in at least one sample
dds_model1 <- DESeq(dds_model1, test = "Wald", fitType = "parametric", sfType = "poscounts")

## Plot deseq2 object
plotDispEsts(dds_model1)
## coefficients DESeq estimated:
resultsNames(dds_model1)

## sum all deviance values and na.rm set as TRUE to ignore NAs
dds_model1_dev <- sum(mcols(dds_model1)$deviance, na.rm = TRUE)
dds_model1_dev

## results
res_model1 <- results(dds_model1, 
                      contrast = c("Group", "EHI", "Control"), 
                      tidy = TRUE, 
                      format = "DataFrame")

rownames(res_model1) <- res_model1$row

res_model1 <- res_model1[,-1]

## summary 
summary(res_model1)

length(res_model1$padj)
```

#### Functions

Below are a number of functions that will generate results and plots for the various Group comparisons that will be made.

```{r Section8_sen_DESeq2_analysis_functions}
## Function to take the results from a specific contrast from DESeq2 and return a significance table
signif_OTU_sen <- function (diff_OTU) {
  diff_OTU <-
    subset(diff_OTU,!is.na(padj) &
             padj <= 0.05 & abs(log2FoldChange) >= 1)
  if (nrow(diff_OTU) == 0)
    return(NULL)
  diff_OTU <-
    cbind(as(diff_OTU, "data.frame"), as(tax_table(deseq_dat_sen)[rownames(diff_OTU),], "matrix"))
  diff_OTU <-
    diff_OTU[order(diff_OTU[["padj"]], decreasing = FALSE), ]
  diff_OTU[["OTU"]] <- rownames(diff_OTU)
  diff_OTU[["OTU_num"]] <- 1:nrow(diff_OTU)
  return(diff_OTU)
}

## Function to plot a specific OTU
plot_OTU_sen <- function (otu) {
  ggplot(subset(deseq_plot_sen, OTU == otu),
         aes(x = Group, y = Abundance, fill = Group)) +
    geom_boxplot() +
    ggtitle(paste(c(otu, as.character(unlist(
      subset(deseq_plot, OTU == otu)[1, c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus")]
    ))), collapse = "; ")) +
    labs(x = "", y = "Relative Abundance") +
    scale_fill_manual(values = group_cols) +
    scale_y_continuous(labels = scales::label_percent(scale = 1)) +
    theme(
      text = font,
      plot.title = element_text(size = 8),
      axis.title   = element_text(size = 18),
      axis.text.y  = element_text(size = 16),
      axis.text.x  = element_text(
        angle = 90,
        hjust = 1,
        size = 24
      ),
      legend.text  = element_text(size = 22),
      legend.title = element_text(size = 22, face = "bold")
    )
}

## Function to process a specific comparison
diff_out_sen <- function (i, j, plot = TRUE) {
  ## Get the significant OTUs
  if (i == j) {
    cat ("Error: ", i, " and ", j, " must be different\n")
    return (NULL)
  }
  ## dds_model1 used for ~Group design
  all_OTU <- results(dds_model1, contrast = c("Group", i, j))
  diff_OTU <- signif_OTU_sen(all_OTU)
  if (is.null(diff_OTU)) {
    cat ("Error: No significant differences identified for ",
         i,
         " vs ",
         j,
         "\n")
    return (list(
      "All"    = all_OTU,
      "Signif" = NULL,
      "Up"     = NULL,
      "Down"   = NULL
    ))
  }
  
  ## Save the results
  write.table(
    diff_OTU,
    paste0(
      analysis_dir,
      "/tabs/Sensitivity_DESeq2_SIGNIF_OTUs_",
      i,
      "_vs_",
      j,
      "_log2foldchange.csv"
    ),
    row.names = FALSE,
    col.names = TRUE,
    sep = ",",
    quote = FALSE
  )
  
  ## Plot the data for these OTUs
  if (plot) {
    outdir <-
      paste0(analysis_dir,
             "/figs/Sensitivity_SIGNIF_OTUs2_",
             i,
             "_vs_",
             j,
             "/")
    if (file.exists(outdir)) {
      unlink(outdir, recursive = TRUE)
      dir.create(outdir)
    }
    for (k in diff_OTU[["OTU_num"]]) {
      otu <- rownames(diff_OTU)[k]
      ggsave(
        filename = paste0(outdir, "OTU_", k, "_", otu, ".png"),
        plot = plot_OTU(otu),
        height = 6,
        width = 8,
        unit = "in",
        dpi = 300
      )
    }
  }
  
  ## Return the results
  return (list(
    "All"    = all_OTU,
    "Signif" = diff_OTU,
    "Up"     = subset(diff_OTU, log2FoldChange > 0),
    "Down"   = subset(diff_OTU, log2FoldChange < 0)
  ))
}

## Function to plot a volcano plot for the results using ggplot2
plot_volcano_sen <- function(x) {
  if (length(x) == 0)
    return(NULL)
  
  # Create a data frame for plotting
  plot_dat <- x[["All"]]
  plot_col <- rep("black", nrow(plot_dat))
  which_down <- which(rownames(plot_dat) %in% rownames(x[["Down"]]))
  which_up <- which(rownames(plot_dat) %in% rownames(x[["Up"]]))
  plot_col[which_down] <- "blue"
  plot_col[which_up] <- "red"
  plot_dat_df <- data.frame(
    log2FoldChange = plot_dat$log2FoldChange,
    padj = plot_dat$padj,
    plot_col = factor(plot_col)
  )
  
  # Create the volcano plot using ggplot2
  p <-
    ggplot(data = plot_dat_df, aes(
      x = log2FoldChange,
      y = -log10(padj),
      color = plot_col
    )) +
    geom_point(size = 3,
               #shape = 1,
               alpha = 0.3) +
    geom_hline(
      yintercept = -log10(0.05),
      linetype = "dashed",
      color = "black"
    ) +
    geom_vline(
      xintercept = log2(2),
      linetype = "dashed",
      color = "black"
    ) +
    geom_vline(
      xintercept = -log2(2),
      linetype = "dashed",
      color = "black"
    ) +
    scale_color_manual(values = c("black", "blue", "red")) +
    labs(x = "log2(Fold Change)", y = "-log10(P Value)") +
    # scale_x_continuous(breaks = seq(-6, 6, 2)) +
    theme_classic() +
    theme(
      text = font,
      legend.position = "none",
      plot.title = element_text(size = 12),
      axis.title = element_text(size = 14),
      axis.text = element_text(size = 16),
      legend.text = element_text(size = 12),
      legend.title = element_text(size = 12, face = "bold")
    )
  
  
  return(p)
}

#### modified below:
plot_volcano2_sen <- function(x) {
  if (length(x) == 0)
    return(NULL)
  
  ## Create a data frame for plotting
  plot_dat <- x[["All"]]
  which_down <- which(rownames(plot_dat) %in% rownames(x[["Down"]]))
  which_up <- which(rownames(plot_dat) %in% rownames(x[["Up"]]))
  plot_col <- rep("black", nrow(plot_dat))
  plot_col[which_down] <- "blue"
  plot_col[which_up] <- "red"
  
  plot_dat_df <- data.frame(
    log2FoldChange = plot_dat$log2FoldChange,
    padj = plot_dat$padj,
    plot_col = factor(plot_col),
    label = rownames(plot_dat)  ## Use rownames for labels to annotate plot
  )
  
  ## Create the volcano plot using ggplot2
  p <-
    ggplot(data = plot_dat_df, aes(
      x = log2FoldChange,
      y = -log10(padj),
      color = plot_col
    )) +
    geom_point(size = 3, alpha = 0.3) +
    geom_hline(
      yintercept = -log10(0.05),
      linetype = "dashed",
      color = "black"
    ) +
    geom_vline(
      xintercept = log2(2),
      linetype = "dashed",
      color = "black"
    ) +
    geom_vline(
      xintercept = -log2(2),
      linetype = "dashed",
      color = "black"
    ) +
    scale_color_manual(values = c("black", "blue", "red")) +
    labs(x = "log2(Fold Change)", y = "-log10(P Value)") +
    theme_classic() +
    theme(
      text = font,
      legend.position = "none",
      plot.title = element_text(size = 12),
      axis.title = element_text(size = 14),
      axis.text = element_text(size = 16),
      legend.text = element_text(size = 12),
      legend.title = element_text(size = 12, face = "bold")
    ) +
    ## Use ggrepel for text repelling, only for OTUs with -log10(padj) > 0.05
    geom_text_repel(
      data = plot_dat_df[which_down | which_up,],
      aes(label = ifelse(
        -log10(padj) > 0.05, as.character(label), ""
      )),
      box.padding = 0.5,
      point.padding = 0.1,
      segment.size = 0.2,
      vjust = 1.5,
      hjust = 0,
      size = 3
    )
  
  return(p)
}

## Function to get the top hits from the DESeq2 analysis
get_top_hits_sen <- function(x, n = 10) {
  if (length(x) == 0)
    return()
  
  # Get the top up and downregulated hits
  positive_df <-
    head(x[["Up"]][order(x[["Up"]][["log2FoldChange"]], decreasing = TRUE),], min(n, nrow(x[["Up"]])))
  negative_df <-
    tail(x[["Down"]][order(x[["Down"]][["log2FoldChange"]], decreasing = TRUE),], min(n, nrow(x[["Down"]])))
  df <- rbind.data.frame(positive_df, negative_df)
  df[["Direction"]] <-
    c(rep("Increase", min(n, nrow(x[["Up"]]))), rep("Decrease", min(n, nrow(x[["Down"]]))))
  
  # Set the sample names from the "Genus" column
  df[["Taxa"]] <-
    factor(paste(otu_or_asv, df[["OTU_num"]], ": ", df[["Genus"]], sep = ""),
           levels = paste(otu_or_asv, df[["OTU_num"]], ": ", df[["Genus"]], sep = ""))
  
  # Create a bar plot using the log2FoldChange values and updated sample names, with colored bars
  df %>%
    ggplot(aes(x = Taxa, y = log2FoldChange, fill = Direction)) +
    geom_bar(stat = 'identity') +
    coord_flip() +
    scale_fill_manual(values = c(
      'Increase' = "#117744",
      'Decrease' = "#DC3220"
    )) +
    ylab("log2 (Fold Change)") + xlab("") +
    # scale_y_continuous(breaks = seq(-6, 6, 2)) +
    theme_classic() +
    theme(
      text = font,
      axis.title   = element_text(size = 14),
      axis.text.y  = element_text(size = 16, face = "bold"),
      axis.text.x  = element_text(size = 14),
      legend.position = "none",
      panel.grid.major.x = element_line(colour = "lightgrey", linewidth = 0.25)
    )
}
```

#### Results
The following results show the differentially abundant taxa between the Patient and Control samples. 

``` {r Section8_sen_DESeq2_analysis_cases_vs_control}
DE_EHIvC_sen <- diff_out("EHI", "Control", plot = FALSE)
cat("Total number of differentially abundant OTUs: ", nrow(DE_EHIvC_sen[["Signif"]]))
cat("Upregulated number of differentially abundant OTUs: ", nrow(DE_EHIvC_sen[["Up"]]))
cat("Downregulated number of differentially abundant OTUs: ", nrow(DE_EHIvC_sen[["Down"]]))
```

#### Volcano

Below is a volcano plot showing the overall differential abundance results for this comparison.

``` {r Section8_sen_DESeq2_analysis_cases_vs_control_volcano, fig.height = 5, fig.width = 7}
volcano_p_sen <- plot_volcano(DE_EHIvC_sen)
volcano_p_sen
```

volcano plot with sig ASVs highlighted

``` {r Section8_sen_DESeq2_analysis_cases_vs_control_volcano2, fig.height = 5, fig.width = 7}
volcano_p_sen <- plot_volcano2(DE_EHIvC_sen)
volcano_p_sen
```

### ALDEx2

#### Setup

```{r Section8_sen_ALDEx2_setup}
## create ASV_table from phyloseq object
ASV_table_sen = otu_table(phyloseq_dat_sen_deseq)

## create groupings variable from phyloseq object
groupings_sen = phyloseq_dat_sen_deseq %>%
  sample_data() %>%
  .$Group %>%
  ## make groupings a character vector! 
  as.character()

ncol(ASV_table_sen)
length(groupings_sen)
```

#### Model

```{r Section8_sen_ALDEx2_model, cache = TRUE}
results_sen <- aldex(reads = ASV_table_sen, 
                     conditions = groupings_sen, 
                     mc.samples = 128, 
                     test="t", 
                     effect=TRUE,
                     include.sample.summary = TRUE, 
                     CI = TRUE,
                     verbose = TRUE, 
                     denom="all")
```

#### Results

```{r Section8_sen_ALDEx2_results}
## Expected p-value of Wilcoxon rank test
results_sen %>%
  dplyr::filter(we.eBH < 0.05)

## Expected Benjamini-Hochberg corrected p-value of Wilcoxon test
results_sen %>%
  dplyr::filter(wi.eBH < 0.05)
```

### ANCOMBC

#### Setup

https://bioconductor.org/packages/devel/bioc/vignettes/ANCOMBC/inst/doc/ANCOM.html

```{r Section8_sen_ANCOMBC_setup}
## object to use
phyloseq_dat_sen_deseq 
## make a copy for ancom analysis
ancom_phyloseq_sen = phyloseq_dat_sen_deseq
## Set Group variable as character string 
sample_data(ancom_phyloseq_sen)$Group <- as.character(sample_data(ancom_phyloseq_sen)$Group)
## Check that levels are now NULL
levels(sample_data(ancom_phyloseq_sen)$Group)
## Make sure names are still "Control" and "EHI"
unique(sample_data(ancom_phyloseq_sen)$Group)
```

#### Model

```{r Section8_sen_ANCOMBC_model}
## use ANCOMBC with phyloseq object
out_sen = ancombc(data = NULL, 
                  assay_name = NULL,
                  tax_level = NULL, 
                  phyloseq = phyloseq_dat_sen_deseq,
                  formula = "Group",
                  p_adj_method = "BH", 
                  prv_cut = 0.10, 
                  lib_cut = min_otus,
                  group = "Group", 
                  struc_zero = TRUE, 
                  neg_lb = TRUE,
                  tol = 1e-5, 
                  max_iter = 100, 
                  conserve = FALSE,
                  alpha = 0.05, 
                  global = TRUE, 
                  n_cl = 1, 
                  verbose = TRUE)
```

#### Results

```{r Section8_sen_ANCOMBC_results}
## Result from the ANCOM-BC log-linear model to determine taxa that are differentially abundant according to the covariate of interest. It contains: 1) log fold changes; 2) standard errors; 3) test statistics; 4) p-values; 5) adjusted p-values; 6) indicators whether the taxon is differentially abundant (TRUE) or not (FALSE).
res_sen = out_sen$res

## Set Colnames
col_names = c("Taxon", "Intercept", "Group")

## 1) log fold changes
res_lfc_sen = res_sen$lfc
colnames(res_lfc_sen) = col_names
#res_lfc
## 2) standard errors
res_se_sen = res_sen$se
colnames(res_se_sen) = col_names
#res_se
## 3) test statistics
res_w_sen = res_sen$W
colnames(res_w_sen) = col_names
#res_w
## 4) p-values
res_p_sen = res_sen$p_val
colnames(res_p_sen) = col_names
#res_p
## 5) adjusted p-values
res_padj_sen = res_sen$q_val
colnames(res_padj_sen) = col_names
#res_padj
## 6) indicators whether the taxon is differentially abundant (TRUE) or not (FALSE).
res_da_sen = res_sen$diff_abn
colnames(res_da_sen) = col_names
#res_da
## check if we have TRUE
unique(res_da_sen$Group)
```

```{r Section8_sen_ANCOM_save_results}
## Create list of ANCOM results
ancom_res_list_sen <- list(
  "Log_Fold_Change" = res_lfc_sen,
  "Standard_Errors" = res_se_sen,
  "Test_Statistics" = res_w_sen,
  "p_value" = res_p_sen,
  "BH_Adjusted_p_value" = res_padj_sen,
  "Dif_Abun_T_or_F" = res_da_sen
)

## Save ANCOM results into a single workbook
openxlsx::write.xlsx(ancom_res_list_sen,
                     file = paste0(analysis_dir, "/tabs/Sensitivity_ANCOM_results.xlsx"))
```

### Compare Methods

```{r Section8_sen_compare}
## DESeq2 results to use
deseq_res_sen = DE_EHIvC_sen$Signif
## set rownames as a new colmun called Taxon
deseq_res_sen = rownames_to_column(deseq_res_sen, var = "Taxon")
## count number of sig ASVs using nrow()
nrow(deseq_res_sen)
## Select only Taxon column
deseq_res_sen = deseq_res_sen %>%
  dplyr::select(Taxon)

## ALDEx2 results to use
aldex2_res_sen = results_sen
## set rownames as a new colmun called Taxon
aldex2_res_sen = rownames_to_column(aldex2_res_sen, var = "Taxon")
## Filter for only p adjust < 0.05
aldex2_res_sen = results %>%
  dplyr::filter(wi.eBH < 0.05)
## count number of sig ASVs using nrow()
nrow(aldex2_res_sen)
## As we have 0 hits, we will create an empty vector for venn diagram
aldex2_hits_sen = character(nrow(aldex2_res_sen))

## ANCOMBC results, filter for only p adjust < 0.05
ancombc_res_sen = res_padj_sen %>%
  dplyr::filter(Group < 0.05)
## count number of sig ASVs using nrow()
nrow(ancombc_res_sen)
## Select only Taxon column
ancombc_res_sen = ancombc_res_sen %>%
  dplyr::select(Taxon)

## Prepare the a list for venn diagram plotting.
da_list_sen <- list(
  DESeq2 = deseq_res_sen$Taxon,
  ALDEx2 = aldex2_hits_sen,  
  "ANCOM-BC" = ancombc_res_sen$Taxon
)
```

#### Venn Diagram

```{r Section8_sen_compare_venn_test, fig.height = 5, fig.width = 5}
## Change font, code from https://stackoverflow.com/questions/69699950/r-geom-venn-how-to-change-font
venn_font <- function(p, font)
{

  grep_grob <- function(gt, lab){
    which(sapply(gt, function(x) grepl(lab, x$name)))
  }
  
  p2 <- ggplot_gtable(ggplot_build(p))
  mygrobs <- p2$grobs
  panel_grob <- mygrobs[[grep_grob(mygrobs, "panel")]]
  venn_grob <- panel_grob$children[[grep_grob(panel_grob$children, "venn")]]
  text_grob <- venn_grob$children[grep_grob(venn_grob$children, "text")]
  text_grob <- do.call(grid::gList, 
                       lapply(text_grob, 
                              function(x) {x$gp$fontfamily <- font; 
                                           x}))
  venn_grob$children[grep_grob(venn_grob$children, "text")] <- text_grob
  panel_grob$children[[grep_grob(panel_grob$children, "venn")]] <- venn_grob
  mygrobs[[grep_grob(mygrobs, "panel")]] <- panel_grob
  p2$grobs <- mygrobs
  grid::grid.newpage()
  grid::grid.draw(p2)
  
}

## Combine all taxa into a single vector and make it unique
venn_all_taxa_sen <- unique(unlist(da_list_sen))

## Create a data frame where each row is a unique taxa
venn_df_sen <- data.frame(Taxon = venn_all_taxa_sen)

## For each method, create a column in taxa_df indicating presence or absence of each taxon
for(method in names(da_list_sen)) {
  venn_df_sen[[method]] <- venn_df_sen$Taxon %in% da_list_sen[[method]]
}

## view
venn_df_sen

## create venn diagram
p_sen <- ggplot(venn_df_sen) +
  geom_venn(aes(A = DESeq2, B = `ALDEx2`, C = `ANCOM-BC`),
            #fill_color = c("#a40000", "#0000ff", "#008000"),
            fill_color = c("#E69F00", "#56B4E9", "green")
            ) +
  coord_fixed() + 
  theme_void() 

## use custom function to set font as serif
venn_font(p_sen, font = "serif")
```

# Session Info

``` {r sessionInfo}
sinfo <- sessionInfo()
sinfo
save.image(paste0(analysis_dir, "/prospective_analysis_manuscript_16S_case_con.Rout"))
```
