---
title: "Biomarkers Analysis (manuscript)"
author: "Alex Gould"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: yes
    toc_depth: 3
    number_sections: yes
    latex_engine: xelatex
  rmarkdown::pdf_document:
    extra_dependencies: float
    toc: yes
html_document:
  toc: yes
toc_float: yes
---

# Setup

First let us set up the R environment:

## Global options

```{r global options, include = TRUE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE,
                      error = FALSE,
                      eval = TRUE,
                      dev = c('pdf', 'png'),
                      width = 60,
                      fig.align = 'center', 
                      fig.height = 10, 
                      fig.width = 10, 
                      fig.dpi = 300,
                      fig.pos = "!H",
                      dpi = 300,
                      out.extra = "",
                      message = FALSE, 
                      cache = FALSE,
                      pdf.options(encoding = "ISOLatin9.enc"),
                      fig.path='figs/rmd/') 
```

## Library 

```{r Library}
## Load the libraries
library(xlsx)
library(rstudioapi)
library(tidyverse)
library(jmv)
library(ggtext)
library(ggpubr)
library(gridtext) 
library(scales)
```

## Parameters

``` {r Parameters}
analysis_dir = "..."
data_dir     = "..."
data_dir2    = "..."
pw_message   = "..."
font         = element_text(family = "serif")
group_cols   = c("Control" = "#0099CC", 
                 "EHI"     = "#FF9900")
```

### Figure parameters 

First, create custom theme to use throughout:

```{r figure_theme}
#create custom theme
theme_ag <- function(){

  theme_classic() %+replace%
    
    theme(
      text = font,
      axis.text.y  = element_text(size = 12, margin = margin(r = 3)),
      axis.title.y = element_text(size = 16, angle = 90, vjust = 2),
      axis.text.x  = element_text(size = 16, margin = margin(r = 3)),
      axis.title.x = element_blank(),
      legend.position = "none"
    )
}
```

Now, make custom axis titles to use throughout:

```{r figure_custom_y_axis}
IFABP_axis = expression("Serum [I-FABP]" ~ "(pg·" * mL^-1 * ")")
CLDN3_axis = expression("Plasma [CLDN-3]" ~ "(ng·" * mL^-1 * ")")
Zon_axis   = expression("Serum [Zonulin]" ~ "(ng·" * mL^-1 * ")")
LBP_axis   = expression("Plasma [LBP]" ~ "(µg·" * mL^-1 * ")")
sCD14_axis  = expression("Plasma [sCD14]" ~ "(µg·" * mL^-1 * ")")
```

Fix for `stat_compare_means()` function to allow Times New Roman font. Thank you: https://stackoverflow.com/questions/77346644/stat-compare-means-unable-to-change-font-family-when-using-comparisons-argu 

```{r stat_compare_means_fix}
fixed_call <- quote(ggsignif::geom_signif(comparisons = comparisons, 
                                          y_position = label.y, 
                                          test = method, test.args = method.args, 
                                          step_increase = step.increase, 
                                          size = bracket.size, textsize = 6, color = color, 
                                          map_signif_level = map_signif_level, 
                                          tip_length = tip.length, data = data,
                                          vjust = vjust, ...))

scm <- as.list(ggpubr::stat_compare_means)
scm[[26]][[2]][[3]][[13]] <- fixed_call
stat_compare_means <- function(...) {
  ggp <- getNamespace('ggpubr')
  .method_info <- ggp$.method_info
  .add_item <- ggp$.add_item
  .is_p.signif_in_mapping <- ggp$.is_p.signif_in_mapping
  .is_empty <- ggp$.is_empty
  do.call(as.function(scm), list(...), quote = FALSE)
}
```

## Load Data
Next we load in the data:

``` {r load_data}
## load data here...
```

### Data Filtering

Check cohort numbers:

```{r cohort_numbers}
## Cohort number
length(unique(data$UIN))

## Number per group (i.e. control/EHI)
length(unique(data$Pair))

table(data$Case)
```

Now remove the three troops that we did not take blood from.

```{r remove_troops}
## data filtering here...
```

Total Ns for analysis:  
SERUM: 65 per group (control/EHI)  
EDTA: 64 per group (control/EHI)  
  
Create two separate data frames from raw data, 1 for serum bloods and 1 for EDTA bloods. 

```{r data_serum}
## data filtering here...
```

```{r data_EDTA}
## data filtering here...
```

# Cases vs Controls

This section will analyse the blood biomarkers of EHI cases vs their pair matched controls. The N is presented in the descriptive section of each biomarker. 

## Intestinal Fatty Acid Binding Protein

### Descriptives

```{r IFABP_descriptives_case}
jmv::descriptives(
    formula = IFABP ~ Case,
    data = data_serum,
    iqr = TRUE,
    sw = TRUE)
```

### Test of Difference

non-parametric data:

```{r IFABP_non-parametric_case}
jmv::ttestIS(
    formula = IFABP ~ Case,
    data = data_serum,
    vars = IFABP,
    students = FALSE,
    mann = TRUE,
    norm = TRUE,
    effectSize = TRUE)
```

### Figure

IFABP figure with log10 y axis. 

``` {r IFABP_case_fig, fig.height = 5, fig.width = 6}
set.seed(123)
IFABP_case_p <- 
  ggplot(data_serum, aes(x = Case, y = IFABP)) +
  geom_jitter(aes(color = Case), 
              width = 0.3,
              height = 0,
              alpha = 0.6,
              size = 3) +
  stat_summary(fun = median,
               geom="crossbar",
               color="black",
               width = 0.6,
               size = 0.6,
               show.legend = FALSE) +
  stat_summary(fun.data = median_iqr,
               geom="errorbar",
               color="black",
               width = 0.3,
               size = 0.8,
               show.legend = FALSE) +
  labs(x = "Group",
       y = IFABP_axis) +
  scale_y_log10(labels = scales::label_number(big.mark = ",")) +
  scale_x_discrete(labels = c("Non-EHI", "EHI")) +
  scale_color_manual(values = group_cols) +
  stat_compare_means(comparisons = list(c("EHI", "Control")),
                     method = "wilcox.test",
                     label = "p.signif",
                     family = "serif",
                     paired = FALSE,
                     bracket.size = 0.6,
                     vjust = -0.2) +
  theme_ag() 

#view figure
IFABP_case_p
```

## Claudin 3

### Descriptives

```{r CLDN3_descriptives_case}
jmv::descriptives(
    formula = CLDN3 ~ Case,
    data = data_EDTA,
    iqr = TRUE,
    sw = TRUE)
```

### Test of Difference

parametric data:

```{r CLDN3_non-parametric_case}
jmv::ttestIS(
    formula = CLDN3 ~ Case,
    data = data_EDTA,
    vars = CLDN3,
    norm = TRUE,
    eqv = TRUE,
    effectSize = TRUE)
```

### Figure

``` {r CLDN3_case_fig, fig.height = 5, fig.width = 6}
set.seed(123)
CLDN3_case_p <- ggplot(data_EDTA, aes(x = Case, y = CLDN3)) +
  geom_jitter(aes(color = Case),
              width = 0.3,
              height = 0,
              alpha = 0.6,
              size = 3) +
  stat_summary(fun = mean, 
               geom = "point", 
               color = "black",
               size = 5,
               show.legend = FALSE) +
  stat_summary(fun.data = mean_sd, 
               geom = "errorbar", 
               color = "black",
               width = 0.3,
               size = 1.0,
               show.legend = FALSE) +
  labs(x = "Group",
       y = CLDN3_axis) +
  scale_y_continuous(breaks = seq(0, 7, 1),
                     limits = c(0, 7),
                     labels = scales::label_number(accuracy = 1)) +
  scale_x_discrete(labels = c("Non-EHI", "EHI")) +
  scale_color_manual(values = group_cols) +
  stat_compare_means(comparisons = list(c("EHI", "Control")),
                     method = "t.test",
                     label = "p.signif",
                     family = "serif",
                     paired = FALSE,
                     bracket.size = 0.6,
                     vjust = -0.2) +
  theme_ag()

#view figure
CLDN3_case_p
```

## Zonulin

### Descriptives

```{r Zonulin_descriptives_case}
jmv::descriptives(
    formula = Zonulin ~ Case,
    data = data_serum,
    iqr = TRUE,
    sw = TRUE)
```

### Test of Difference

non-parametric data:

```{r Zonulin_non-parametric_case}
jmv::ttestIS(
    formula = Zonulin ~ Case,
    data = data,
    vars = Zonulin,
    norm = TRUE,
    eqv = TRUE,
    effectSize = TRUE)
```

### Figure

``` {r Zonulin_case_fig, fig.height = 5, fig.width = 6}
set.seed(123)
Zon_case_p <- ggplot(data_serum, aes(x = Case, y = Zonulin)) +
  geom_jitter(aes(color = Case),
              width = 0.3,
              height = 0,
              alpha = 0.6,
              size = 3) +
  stat_summary(fun = mean, 
               geom = "point", 
               color = "black",
               size = 5,
               show.legend = FALSE) +
  stat_summary(fun.data = mean_sd, 
               geom = "errorbar", 
               color = "black",
               width = 0.3,
               size = 1.0,
               show.legend = FALSE) +
  labs(x = "Group",
       y = Zon_axis) +
  scale_y_continuous(breaks = seq(0, 70, 10),
                     limits = c(0, 70)) +
  scale_x_discrete(labels = c("Non-EHI", "EHI")) +
  scale_color_manual(values = group_cols) +
  stat_compare_means(comparisons = list(c("EHI", "Control")),
                     method = "t.test",
                     label = "p.signif",
                     family = "serif",
                     paired = FALSE,
                     bracket.size = 0.6,
                     vjust = -0.2) +
  theme_ag()

#view figure
Zon_case_p
```

## Lipopolysaccharide Binding Protein

### Descriptives

```{r LBP_descriptives_case}
jmv::descriptives(
    formula = LBP ~ Case,
    data = data_EDTA,
    iqr = TRUE,
    sw = TRUE)
```

### Test of Difference

parametric data:

```{r LBP_parametric_class}
jmv::ttestIS(
    formula = LBP ~ Case,
    data = data_EDTA,
    vars = LBP,
    norm = TRUE,
    eqv = TRUE,
    effectSize = TRUE)
```

### Figure

``` {r LBP_case_fig, fig.height = 5, fig.width = 6}
set.seed(123)
LBP_case_p <- ggplot(data_EDTA, aes(x = Case, y = LBP)) +
  geom_jitter(aes(color = Case), 
              width = 0.3,
              height = 0,
              alpha = 0.6,
              size = 3) + 
  stat_summary(fun = mean, 
               geom = "point", 
               color = "black",
               size = 5,
               show.legend = FALSE) +
  stat_summary(fun.data = mean_sd, 
               geom = "errorbar", 
               color = "black",
               width = 0.3,
               size = 1.0,
               show.legend = FALSE) +
  labs(x = "Group",
       y = LBP_axis) +
  scale_y_continuous(breaks = seq(0.0, 6.0, 1.0),
                     limits = c(0.0, 6.5),
                     labels = scales::label_number(accuracy = 1)) +
  scale_x_discrete(labels = c("Non-EHI", "EHI")) +
  scale_color_manual(values = group_cols) +
  stat_compare_means(comparisons = list(c("EHI", "Control")),
                     method = "t.test",
                     label = "p.signif",
                     family = "serif",
                     paired = FALSE,
                     bracket.size = 0.6,
                     vjust = -0.2) +
  theme_ag()

#view figure
LBP_case_p
```

## Soluble Cluster of Differentiation 14

### Descriptives

```{r sCD14_descriptives_case}
jmv::descriptives(
    formula = sCD14 ~ Case,
    data = data_EDTA,
    iqr = TRUE,
    sw = TRUE)
```

### Test of Difference

non-parametric data:

```{r sCD14_non_parametric_case}
jmv::ttestIS(
    formula = sCD14 ~ Case,
    data = data_EDTA,
    vars = sCD14,
    mann = TRUE,
    students = FALSE,
    norm = TRUE,
    eqv = TRUE,
    effectSize = TRUE)
```

### Figure

``` {r sCD14_case_fig, fig.height = 5, fig.width = 6}
set.seed(123)
sCD14_case_p <- ggplot(data_EDTA, aes(x = Case, y = sCD14)) +
  geom_jitter(aes(color = Case), 
              width = 0.3,
              height = 0,
              alpha = 0.6,
              size = 3) +
  stat_summary(fun = median, 
               geom ="crossbar", 
               color ="black",
               width = 0.6,
               size = 0.6,
               show.legend = FALSE) +
  stat_summary(fun.data = median_iqr,
               geom ="errorbar",
               color ="black",
               width = 0.3,
               size = 0.8,
               show.legend = FALSE) +
  labs(x = "Group",
       y = sCD14_axis) +
  scale_y_continuous(breaks = seq(0, 2, .25),
                     limits = c(0, 2.1),
                     labels = scales::label_number(accuracy = 0.01)) +
  scale_x_discrete(labels = c("Non-EHI", "EHI")) +
  scale_color_manual(values = group_cols) +
  stat_compare_means(comparisons = list(c("EHI", "Control")),
                     method = "wilcox.test",
                     label = "p.signif",
                     family = "serif",
                     paired = FALSE,
                     bracket.size = 0.6,
                     vjust = -0.2) +
  theme_ag()

#view figure
sCD14_case_p
```

# Sensitivity Analysis

## Data

Now create the two data frames for serum and EDTA analysis as done previously. Remove NSAID users and their respective matched-pair UIN.    
Serum:

```{r sen_data_serum}
## Create data frame for serum sensitivity analysis
sen_data_serum <- data_serum

## Create object which contains NSAID users
NSAID_users_serum <- sen_data_serum %>% 
  ## Filter rows where NSAID is "Yes"
  dplyr::filter(NSAID == "Yes")

## Extract unique Pair values
NSAID_pairs_serum <- unique(NSAID_users_serum$Pair)

## Count number of NSAID users
length(unique(NSAID_users_serum$UIN))

## Create serum data frame for sensitivity analysis
sen_data_serum <- sen_data_serum %>%
  ## Remove NSAID pairs from data
  dplyr::filter(!Pair %in% NSAID_pairs_serum)

## Updated numbers:
## Cohort number
length(unique(sen_data_serum$UIN))

## Number per group (i.e. control/EHI)
length(unique(sen_data_serum$Pair))
```

EDTA:

```{r sen_data_EDTA}
## Create data frame for EDTA sensitivity analysis
sen_data_EDTA <- data_EDTA

## Create object which contains NSAID users
NSAID_users_EDTA <- sen_data_EDTA %>% 
  ## Filter rows where NSAID is "Yes"
  dplyr::filter(NSAID == "Yes")

## Extract unique Pair values
NSAID_pairs_EDTA <- unique(NSAID_users_EDTA$Pair)

## Count number of NSAID users
length(unique(NSAID_users_EDTA$UIN))

## Create EDTA data frame for sensitivity analysis
sen_data_EDTA <- sen_data_EDTA %>%
  ## Remove NSAID pairs from data
  dplyr::filter(!Pair %in% NSAID_pairs_EDTA)

## Updated numbers:
## Cohort number
length(unique(sen_data_EDTA$UIN))

## Number per group (i.e. control/EHI)
length(unique(sen_data_EDTA$Pair))
```

As part of sensitivity analysis, this section will analyse the blood biomarkers of EHI cases vs their pair matched controls, where NSAID users and individuals with a GI disorder, and their respective pair are excluded. The N is presented in the descriptive section of each biomarker. 

## Intestinal Fatty Acid Binding Protein

### Descriptives

```{r sen_IFABP_descriptives_case}
jmv::descriptives(
    formula = IFABP ~ Case,
    data = sen_data_serum,
    iqr = TRUE,
    sw = TRUE)
```

### Test of Difference

non-parametric data:

```{r sen_IFABP_non-parametric_case}
jmv::ttestIS(
    formula = IFABP ~ Case,
    data = sen_data_serum,
    vars = IFABP,
    students = FALSE,
    mann = TRUE,
    norm = TRUE,
    effectSize = TRUE)
```

### Figure

``` {r sen_IFABP_case_fig, fig.height = 5, fig.width = 6}
set.seed(123)
sen_IFABP_case_p <- ggplot(sen_data_serum, aes(x = Case, y = IFABP)) +
  geom_jitter(aes(color = Case), 
              width = 0.3,
              height = 0,
              alpha = 0.6,
              size = 3) + 
  stat_summary(fun = median, 
               geom="crossbar", 
               color="black",
               width = 0.6,
               size = 0.6,
               show.legend = FALSE) +
  stat_summary(fun.data = median_iqr,
               geom="errorbar",
               color="black",
               width = 0.3,
               size = 0.8,
               show.legend = FALSE) +
  labs(x = "Group",
       y = IFABP_axis) +
  scale_y_log10(labels = scales::label_number(big.mark = ",")) +
  scale_x_discrete(labels = c("Non-EHI", "EHI")) +
  scale_color_manual(values = group_cols) +
  stat_compare_means(comparisons = list(c("EHI", "Control")),
                     method = "wilcox.test",
                     label = "p.signif",
                     family = "serif",
                     paired = FALSE,
                     bracket.size = 0.6,
                     vjust = -0.2) +
  theme_ag() +
  ## left align y axis text
  theme(axis.text.y = element_text(hjust = 1))

#view figure
sen_IFABP_case_p
```

## Claudin 3

### Descriptives

```{r sen_CLDN3_descriptives_case}
jmv::descriptives(
    formula = CLDN3 ~ Case,
    data = sen_data_EDTA,
    iqr = TRUE,
    sw = TRUE)
```

### Test of Difference

parametric data:

```{r sen_CLDN3_non-parametric_case}
jmv::ttestIS(
    formula = CLDN3 ~ Case,
    data = sen_data_EDTA,
    vars = CLDN3,
    norm = TRUE,
    eqv = TRUE,
    effectSize = TRUE)
```

### Figure

``` {r sen_CLDN3_case_fig, fig.height = 5, fig.width = 6}
set.seed(123)
sen_CLDN3_case_p <- ggplot(sen_data_EDTA, aes(x = Case, y = CLDN3)) +
  geom_jitter(aes(color = Case),
              width = 0.3,
              height = 0,
              alpha = 0.6,
              size = 3) + 
  stat_summary(fun = mean,
               geom = "point",
               color = "black",
               size = 5,
               show.legend = FALSE) +
  stat_summary(fun.data = mean_sd,
               geom = "errorbar",
               color = "black",
               width = 0.3,
               size = 1.0,
               show.legend = FALSE) +
  labs(x = "Group",
       y = CLDN3_axis) +
  scale_y_continuous(breaks = seq(0, 7, 1),
                     limits = c(0, 7),
                     labels = scales::label_number(accuracy = 1)) +
  scale_x_discrete(labels = c("Non-EHI", "EHI")) +
  scale_color_manual(values = group_cols) +
  stat_compare_means(comparisons = list(c("EHI", "Control")),
                     method = "t.test",
                     label = "p.signif",
                     family = "serif",
                     paired = FALSE,
                     bracket.size = 0.6,
                     vjust = -0.2) +
  theme_ag()

#view figure
sen_CLDN3_case_p
```

## Zonulin

### Descriptives

```{r sen_Zonulin_descriptives_case}
jmv::descriptives(
    formula = Zonulin ~ Case,
    data = sen_data_serum,
    iqr = TRUE,
    sw = TRUE)
```

### Test of Difference

non-parametric data:

```{r sen_Zonulin_non-parametric_case}
jmv::ttestIS(
  formula = Zonulin ~ Case,
  data = sen_data_serum,
  vars = Zonulin,
  norm = TRUE,
  eqv = TRUE,
  effectSize = TRUE)
```

### Figure

``` {r sen_Zonulin_case_fig, fig.height = 5, fig.width = 6}
set.seed(123)
sen_Zon_case_p <- ggplot(sen_data_serum, aes(x = Case, y = Zonulin)) +
  geom_jitter(aes(color = Case),
              width = 0.3,
              height = 0,
              alpha = 0.6,
              size = 3) +
  stat_summary(fun = mean,
               geom = "point",
               color = "black",
               size = 5,
               show.legend = FALSE) +
  stat_summary(fun.data = mean_sd,
               geom = "errorbar",
               color = "black",
               width = 0.3,
               size = 1.0,
               show.legend = FALSE) +
  labs(x = "Group",
       y = Zon_axis) +
  scale_y_continuous(breaks = seq(0, 70, 10),
                     limits = c(0, 70)) +
  scale_x_discrete(labels = c("Non-EHI", "EHI")) +
  scale_color_manual(values = group_cols) +
  stat_compare_means(comparisons = list(c("EHI", "Control")),
                     method = "t.test",
                     label = "p.signif",
                     family = "serif",
                     paired = FALSE,
                     bracket.size = 0.6,
                     vjust = -0.2) +
  theme_ag()

#view figure
sen_Zon_case_p
```

## Lipopolysaccharide Binding Protein

### Descriptives

```{r sen_LBP_descriptives_case}
jmv::descriptives(
    formula = LBP ~ Case,
    data = sen_data_EDTA,
    iqr = TRUE,
    sw = TRUE)
```

### Test of Difference

parametric data:

```{r sen_LBP_parametric_class}
jmv::ttestIS(
    formula = LBP ~ Case,
    data = sen_data_EDTA,
    vars = LBP,
    norm = TRUE,
    eqv = TRUE,
    effectSize = TRUE)
```

### Figure

``` {r sen_LBP_case_fig, fig.height = 5, fig.width = 6}
set.seed(123)
sen_LBP_case_p <- ggplot(sen_data_EDTA, aes(x = Case, y = LBP)) +
  geom_jitter(aes(color = Case), 
              width = 0.3,
              height = 0,
              alpha = 0.6,
              size = 3) + 
  stat_summary(fun = mean, 
               geom = "point", 
               color = "black",
               size = 5,
               show.legend = FALSE) +
  stat_summary(fun.data = mean_sd, 
               geom = "errorbar", 
               color = "black",
               width = 0.3,
               size = 1.0,
               show.legend = FALSE) +
  labs(x = "Group",
       y = LBP_axis) +
  scale_y_continuous(breaks = seq(0.0, 6.0, 1.0),
                     limits = c(0.0, 6.5),
                     labels = scales::label_number(accuracy = 1)) +
  scale_x_discrete(labels = c("Non-EHI", "EHI")) +
  scale_color_manual(values = group_cols) +
  stat_compare_means(comparisons = list(c("EHI", "Control")),
                     method = "t.test",
                     label = "p.signif",
                     family = "serif",
                     paired = FALSE,
                     bracket.size = 0.6,
                     vjust = -0.2) +
  theme_ag()

#view figure
sen_LBP_case_p
```

## Soluble Cluster of Differentiation 14

### Descriptives

```{r sen_sCD14_descriptives_case}
jmv::descriptives(
    formula = sCD14 ~ Case,
    data = sen_data_EDTA,
    iqr = TRUE,
    sw = TRUE)
```

### Test of Difference

parametric data:

```{r sen_sCD14_parametric_case}
jmv::ttestIS(
    formula = sCD14 ~ Case,
    data = sen_data_EDTA,
    vars = sCD14,
    norm = TRUE,
    eqv = TRUE,
    effectSize = TRUE)
```

### Figure

``` {r sen_sCD14_case_fig, fig.height = 5, fig.width = 6}
set.seed(123)
sen_sCD14_case_p <- ggplot(sen_data_EDTA, aes(x = Case, y = sCD14)) +
  geom_jitter(aes(color = Case), 
              width = 0.3,
              height = 0,
              alpha = 0.6,
              size = 3) + 
  stat_summary(fun = mean, 
               geom = "point", 
               color = "black",
               size = 5,
               show.legend = FALSE) +
  stat_summary(fun.data = mean_sd, 
               geom = "errorbar", 
               color = "black",
               width = 0.3,
               size = 1.0,
               show.legend = FALSE) +
  labs(x = "Group",
       y = sCD14_axis) +
  scale_y_continuous(breaks = seq(0, 2, .25),
                     limits = c(0, 2.1),
                     labels = scales::label_number(accuracy = 0.01)) +
  scale_x_discrete(labels = c("Non-EHI", "EHI")) +
  scale_color_manual(values = group_cols) +
  stat_compare_means(comparisons = list(c("EHI", "Control")),
                     method = "t.test",
                     label = "p.signif",
                     family = "serif",
                     paired = FALSE,
                     bracket.size = 0.6,
                     vjust = -0.2) +
  theme_ag()

#view figure
sen_sCD14_case_p
```

# Figure Panel

Create a panel of figures.

```{r biomarkers_panel, fig.height = 9, fig.width = 12}
## arrange by biomarker type
## Enterocyte damage & tight junction integrity
row1 <- ggarrange(IFABP_case_p, 
                  Zon_case_p, 
                  CLDN3_case_p,
                  ncol = 3, 
                  labels = LETTERS[1:3], 
                  font.label = list(size = 18, 
                                    face = "bold",
                                    family = "serif"))
## Microbial translocation
row2 <- ggarrange(LBP_case_p, 
                  sCD14_case_p,
                  ncol = 3, 
                  labels = LETTERS[4:5], 
                  font.label = list(size = 18, 
                                    face = "bold",
                                    family = "serif"))
## Combine rows
biomarkers_case_panel <- ggarrange(row1, row2, nrow=2)
## Show panel
biomarkers_case_panel
```

Add annotation: 

```{r biomarkers_annotation_panel, fig.height = 9, fig.width = 13}
## Annotate row1
row1ann <- annotate_figure(row1,
                           left = text_grob("Enterocyte damage &\ntight junction integrity\n", 
                                            rot = 90, 
                                            size = 16, 
                                            face = "bold",
                                            family = "serif"))

## Annotate row2
row2ann <- annotate_figure(row2,
                           left = text_grob("\nMicrobial translocation\n", 
                                            rot = 90, 
                                            size = 16, 
                                            face = "bold",
                                            family = "serif"))
## Combine rows
biomarkers_case_annotation_panel <- ggarrange(row1ann, row2ann, nrow=2)
## Show panel
biomarkers_case_annotation_panel

## Save
ggsave(
  filename = paste0(
    analysis_dir,
    "figs/ggsave_biomarkers_annotation_panel",
    ".png"
  ),
  plot = biomarkers_case_annotation_panel,
  height = 9,
  width = 13,
  unit = "in",
  dpi = 600
)

## Save
ggsave(
  filename = paste0(
    analysis_dir,
    "figs/ggsave_biomarkers_annotation_panel",
    ".jpeg"
  ),
  plot = biomarkers_case_annotation_panel,
  height = 9,
  width = 13,
  unit = "in",
  dpi = 600
)
```

# Session info

``` {r sessionInfo}
sinfo <- sessionInfo()
sinfo
save.image(paste0(analysis_dir, "/prospective_analysis_manuscript_biomarkers.Rout"))
```